### 2020-06-29 tsibbletalk.Rmd
init commit

---
title: "Conversations in time: interactive visualization to explore structured temporal data"
author:
- familyname: Wang
  othernames: Earo
  address: Department of Statistics, The Univeristy of Auckland, New Zealand
  email: earo.wang@auckland.ac.nz
  correspondingauthor: true
- familyname: Cook
  othernames: Dianne
  address: Department of Econometrics and Business Statistics, Monash University, Australia
  email: dicook@monash.edu
abstract: "A brief summary of our ideas"
keywords: ""
blind: false
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
output:
  UOARmdTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    number_sections: yes
    citation_package: biblatex
    dev: "pdf"
---

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
  cache = TRUE, cache.path = 'cache/',
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
```

# Introduction

### 2020-07-19 tsibbletalk.Rmd
bullet points
outline sections

@@
 # Introduction
+
+* An ensemble of graphics
+* Accelerate the exploratory data visualization process
+
+# Background
+
+## Interactive visualization systems with linking views
+
+* {ggobi} and {xggobi}
+* {cranvas} and {cranvastime}
+* [crossfilter.js](http://crossfilter.github.io/crossfilter/)
+* {crosstalk} and html widgets
+
+## Tidy temporal data and workflow
+
+* {tsibble}
+* and tsibble-centered workflow
+
+# Shared temporal data for coordinated views
+
+* Symbolic formula to express structural specifications among keyed units, using `/` and `*` from Wilkinson notation (10.2307/2346786) for nesting and crossing.
+* Nesting variables generate hierarchical tree, hence `plotly_key_tree()`. Overview and navigation made easier.
+* One-to-many linking: marking a single point of interest highlights all other points that share particular data values (connect-type)
+* Lists of key values, and json
+* R6 subclass of `SharedData` from {crosstalk}
+
+# Slicing and dicing time
+
+* Time wrapping: absolute time to relative time, grouped by corresponding lower-resolution time. date-times -> hours, by dates
+* Transform and send newly-transformed data via server, without updating anything else (e.g. layout and graphical elements), to avoid completely redrawing.
+* A shiny module: slider and plot
+
+# Case study: monthly domestic tourist trips in Australia
+
+# Conclusions and discussions
+
+# References {-}

### 2020-07-22 tsibbletalk.Rmd
brief draft on tsibble and its workflow

@@
 author:
-- familyname: Wang
-  othernames: Earo
-  address: Department of Statistics, The Univeristy of Auckland, New Zealand
+- name: Earo Wang
+  affiliation: The Univeristy of Auckland
+  address:
+  - line 1
+  - line 2
   email: earo.wang@auckland.ac.nz
-  correspondingauthor: true
-- familyname: Cook
-  othernames: Dianne
-  address: Department of Econometrics and Business Statistics, Monash University, Australia
+- name: Dianne Cook
+  affiliation: Monash Univeristy
+  address:
+  - line 1
+  - line 2
   email: dicook@monash.edu
-abstract: "A brief summary of our ideas"
-keywords: ""
-blind: false
-toc: false
 bibliography: references.bib
-biblio-style: authoryear-comp
-output:
-  UOARmdTemplates::workingpaper:
-    fig_caption: yes
-    fig_height: 5
-    fig_width: 8
-    includes:
-      in_header: preamble.tex
-    number_sections: yes
-    citation_package: biblatex
-    dev: "pdf"
+abstract: >
+  An abstract of less than 150 words.
+preamble: |
+  % Any extra LaTeX you need in the preamble
+output: rticles::rjournal_article
 ---
@@
 
-# Background
+# Background and motivation
 
@@
 * {cranvas} and {cranvastime}
-* [crossfilter.js](http://crossfilter.github.io/crossfilter/)
+* [crossfilter.js](http://crossfilter.github.io/crossfilter/) & [dc.js](https://dc-js.github.io/dc.js/)
 * {crosstalk} and html widgets
@@
 
-* {tsibble}
-* and tsibble-centered workflow
+The \CRANpkg{tsibble} package extends the `data.frame` and \CRANpkg{tibble} structure to represent temporal data in tidy format [cite].
+A tsibble consists of *index*, *key*, and other measured variables in one data frame.
+The *index* column holds time-based indices.
+The *key* column(s) uniquely identifies a collection of related observational units during a period of time defined by *index*.
+They are "sticky" columns to a `tsibble` over the course of transformation.
+
+The \CRANpkg{feasts} and \CRANpkg{fable} packages, part of the **tidyverts** suite, aim to make time series analysis easier.
+They provide analytical and forecasting tools for the tsibble data structure, generating tsibble-centered workflow.
+Functions, such as `features()` and `model()`, summarise a sequence of indexed values down to a single statistic or model by every observational unit.
+The output is a normal table, where each row corresponds to an observational unit denoted by "key".
+In the context of relational databases, the "key" acts like a foreign key in a reduced form of tsibble, while the index and key together operates like a primary key.
+
+At the early stage of exploratory temporal data analysis, time series plots and scatterplots goes hand by hand. (insert figures below)
 

### 2020-07-22 tsibbletalk.tex
brief draft on tsibble and its workflow

% !TeX root = RJwrapper.tex
\title{Conversations in time: interactive visualization to explore
structured temporal data}
\author{by Earo Wang, Dianne Cook}

\maketitle

\abstract{%
An abstract of less than 150 words.
}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\begin{itemize}
\tightlist
\item
  An ensemble of graphics
\item
  Accelerate the exploratory data visualization process
\end{itemize}

\hypertarget{background-and-motivation}{%
\section{Background and motivation}\label{background-and-motivation}}

\hypertarget{interactive-visualization-systems-with-linking-views}{%
\subsection{Interactive visualization systems with linking
views}\label{interactive-visualization-systems-with-linking-views}}

\begin{itemize}
\tightlist
\item
  \{ggobi\} and \{xggobi\}
\item
  \{cranvas\} and \{cranvastime\}
\item
  \href{http://crossfilter.github.io/crossfilter/}{crossfilter.js} \&
  \href{https://dc-js.github.io/dc.js/}{dc.js}
\item
  \{crosstalk\} and html widgets
\end{itemize}

\hypertarget{tidy-temporal-data-and-workflow}{%
\subsection{Tidy temporal data and
workflow}\label{tidy-temporal-data-and-workflow}}

The \CRANpkg{tsibble} package extends the \texttt{data.frame} and
\CRANpkg{tibble} structure to represent temporal data in tidy format
{[}cite{]}. A tsibble consists of \emph{index}, \emph{key}, and other
measured variables in one data frame. The \emph{index} column holds
time-based indices. The \emph{key} column(s) uniquely identifies a
collection of related observational units during a period of time
defined by \emph{index}. They are ``sticky'' columns to a
\texttt{tsibble} over the course of transformation.

The \CRANpkg{feasts} and \CRANpkg{fable} packages, part of the
\textbf{tidyverts} suite, aim to make time series analysis easier. They
provide analytical and forecasting tools for the tsibble data structure,
generating tsibble-centered workflow. Functions, such as
\texttt{features()} and \texttt{model()}, summarise a sequence of
indexed values down to a single statistic or model by every
observational unit. The output is a normal table, where each row
corresponds to an observational unit denoted by ``key''. In the context
of relational databases, the ``key'' acts like a foreign key in a
reduced form of tsibble, while the index and key together operates like
a primary key.

At the early stage of exploratory temporal data analysis, time series
plots and scatterplots goes hand by hand. (insert figures below)

\hypertarget{shared-temporal-data-for-coordinated-views}{%
\section{Shared temporal data for coordinated
views}\label{shared-temporal-data-for-coordinated-views}}

\begin{itemize}
\tightlist
\item
  Symbolic formula to express structural specifications among keyed
  units, using \texttt{/} and \texttt{*} from Wilkinson notation
  (10.2307/2346786) for nesting and crossing.
\item
  Nesting variables generate hierarchical tree, hence
  \texttt{plotly\_key\_tree()}. Overview and navigation made easier.
\item
  One-to-many linking: marking a single point of interest highlights all
  other points that share particular data values (connect-type)
\item
  Lists of key values, and json
\item
  R6 subclass of \texttt{SharedData} from \{crosstalk\}
\end{itemize}

\hypertarget{slicing-and-dicing-time}{%
\section{Slicing and dicing time}\label{slicing-and-dicing-time}}

\begin{itemize}
\tightlist
\item
  Time wrapping: absolute time to relative time, grouped by
  corresponding lower-resolution time. date-times -\textgreater{} hours,
  by dates
\item
  Transform and send newly-transformed data via server, without updating
  anything else (e.g.~layout and graphical elements), to avoid
  completely redrawing.
\item
  A shiny module: slider and plot
\end{itemize}

\hypertarget{case-study-monthly-domestic-tourist-trips-in-australia}{%
\section{Case study: monthly domestic tourist trips in
Australia}\label{case-study-monthly-domestic-tourist-trips-in-australia}}

\hypertarget{conclusions-and-discussions}{%
\section{Conclusions and
discussions}\label{conclusions-and-discussions}}

\bibliography{references.bib}

\address{%
Earo Wang\\
The Univeristy of Auckland\\
line 1\\ line 2\\
}
\href{mailto:earo.wang@auckland.ac.nz}{\nolinkurl{earo.wang@auckland.ac.nz}}

\address{%
Dianne Cook\\
Monash Univeristy\\
line 1\\ line 2\\
}
\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}

### 2020-07-27 tsibbletalk.Rmd
draft static seasonal plots

@@
   address:
-  - line 1
-  - line 2
+  - Department of Statistics
+  - The Univeristy of Auckland
+  - New Zealand
   email: earo.wang@auckland.ac.nz
@@
   address:
-  - line 1
-  - line 2
+  - Department of Econometrics and Business Statistics
+  - Monash University
+  - Australia
   email: dicook@monash.edu
@@
 
-* Time wrapping: absolute time to relative time, grouped by corresponding lower-resolution time. date-times -> hours, by dates
+The other critical aspect of a tsibble is "index", that provides foundational temporal context.
+A common tool in time series analytical toolkit is seasonal plots that lay time series not on the whole time scale, but on an origin-less relative time unit, such as `gg_season()` in the {feasts} package.
+It helps to examine and emphasise periodic/aperiodic patterns, comparing to time series plots that primarily focus on trends.
+Standard seasonal plots break the overall time into two components: seasonal periods on the x-axis, and grouped by their corresponding lower-resolution time.
+For example, monthly data can be decomposed into months separated by years, and hourly data into hours grouped by days.
+
 * Transform and send newly-transformed data via server, without updating anything else (e.g. layout and graphical elements), to avoid completely redrawing.

### 2020-07-27 tsibbletalk.tex
draft static seasonal plots

@@
 
+The other critical aspect of a tsibble is ``index'', that provides
+foundational temporal context. A common tool in time series analytical
+toolkit is seasonal plots that lay time series not on the whole time
+scale, but on an origin-less relative time unit, such as
+\texttt{gg\_season()} in the \{feasts\} package. It helps to examine and
+emphasise periodic/aperiodic patterns, comparing to time series plots
+that primarily focus on trends. Standard seasonal plots break the
+overall time into two components: seasonal periods on the x-axis, and
+grouped by their corresponding lower-resolution time. For example,
+monthly data can be decomposed into months separated by years, and
+hourly data into hours grouped by days.
+
 \begin{itemize}
 \tightlist
-\item
-  Time wrapping: absolute time to relative time, grouped by
-  corresponding lower-resolution time. date-times -\textgreater{} hours,
-  by dates
 \item
@@
 The Univeristy of Auckland\\
-line 1\\ line 2\\
+Department of Statistics\\ The Univeristy of Auckland\\ New Zealand\\
 }
@@
 Monash Univeristy\\
-line 1\\ line 2\\
+Department of Econometrics and Business Statistics\\ Monash
+University\\ Australia\\
 }

### 2020-07-28 tsibbletalk.Rmd
draft shiny module

@@
 - name: Dianne Cook
-  affiliation: Monash Univeristy
+  affiliation: Monash Univerisity
   address:
@@
 The other critical aspect of a tsibble is "index", that provides foundational temporal context.
-A common tool in time series analytical toolkit is seasonal plots that lay time series not on the whole time scale, but on an origin-less relative time unit, such as `gg_season()` in the {feasts} package.
+A common tool in time series analytical toolkit is seasonal plots that lay time series not on the whole time scale, but on an origin-less relative time unit, for example `gg_season()` in the {feasts} package.
 It helps to examine and emphasise periodic/aperiodic patterns, comparing to time series plots that primarily focus on trends.
@@
 For example, monthly data can be decomposed into months separated by years, and hourly data into hours grouped by days.
+Data collected at lower-level resolutions often exhibits more than one seasonal patterns.
+To discover typical seasonal or non-typical profiles, it is helpful to quickly browse through many possible periods.
+Interactivity ought to be enabled.
 
-* Transform and send newly-transformed data via server, without updating anything else (e.g. layout and graphical elements), to avoid completely redrawing.
-* A shiny module: slider and plot
+The {tsibbletalk} package provides a pair of UI and server functions, as a shiny module, to help with finding interesting time slices in a shiny application.
+The pair, `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece.
+Like all shiny modules, users should supply a unique session id.
+The UI function `tsibbleDiceUI()` shows a slider that controls the number of periods, and a plot specified by users.
+The server function `tsibbleDiceServer()` is the workhorse, transforming data and updating the plot.
+It expects a `ggplot` (converted to `plotly` via `ggplotly()`) or `plotly` object.
+This plot can be line charts, or other graphical elements (such as boxplots).
+But it assumes that tsibble's time index is plotted on the x-axis.
+The other mandatory argument is to specify the number of seasonal periods that requires shifting.
+
+* Transform and send newly-transformed data via server, without updating anything else (e.g. layout and graphical elements), to avoid completely redrawing. Performance.
+* Double dispatch
 

### 2020-07-28 tsibbletalk.tex
draft shiny module

@@
 toolkit is seasonal plots that lay time series not on the whole time
-scale, but on an origin-less relative time unit, such as
+scale, but on an origin-less relative time unit, for example
 \texttt{gg\_season()} in the \{feasts\} package. It helps to examine and
@@
 monthly data can be decomposed into months separated by years, and
-hourly data into hours grouped by days.
+hourly data into hours grouped by days. Data collected at lower-level
+resolutions often exhibits more than one seasonal patterns. To discover
+typical seasonal or non-typical profiles, it is helpful to quickly
+browse through many possible periods. Interactivity ought to be enabled.
+
+The \{tsibbletalk\} package provides a pair of UI and server functions,
+as a shiny module, to help with finding interesting time slices in a
+shiny application. The pair, \texttt{tsibbleDiceUI()} and
+\texttt{tsibbleDiceServer()}, presents a clean interface and forms a
+resusable piece. Like all shiny modules, users should supply a unique
+session id. The UI function \texttt{tsibbleDiceUI()} shows a slider that
+controls the number of periods, and a plot specified by users. The
+server function \texttt{tsibbleDiceServer()} is the workhorse,
+transforming data and updating the plot. It expects a \texttt{ggplot}
+(converted to \texttt{plotly} via \texttt{ggplotly()}) or
+\texttt{plotly} object. This plot can be line charts, or other graphical
+elements (such as boxplots). But it assumes that tsibble's time index is
+plotted on the x-axis. The other mandatory argument is to specify the
+number of seasonal periods that requires shifting.
 
@@
   anything else (e.g.~layout and graphical elements), to avoid
-  completely redrawing.
+  completely redrawing. Performance.
 \item
-  A shiny module: slider and plot
+  Double dispatch
 \end{itemize}
@@
 Dianne Cook\\
-Monash Univeristy\\
+Monash Univerisity\\
 Department of Econometrics and Business Statistics\\ Monash

### 2020-07-28 tsibbletalk.Rmd
draft on plotly.js react method

@@
 
-* Transform and send newly-transformed data via server, without updating anything else (e.g. layout and graphical elements), to avoid completely redrawing. Performance.
-* Double dispatch
+Transformed data generally requires redrawing the plot, and worsen the performance of shiny.
+This shiny module directly manipulate the underlying data with the `plotly.js` react method.
+Only transformed data is sent to the server side, while keeping the rest configuration unchanged (e.g.
+layout and graphical elements), for the performance reason.
+Users will not experience an obvious delay when the slider input is changed.
 

### 2020-07-28 tsibbletalk.tex
draft on plotly.js react method

@@
 
-\begin{itemize}
-\tightlist
-\item
-  Transform and send newly-transformed data via server, without updating
-  anything else (e.g.~layout and graphical elements), to avoid
-  completely redrawing. Performance.
-\item
-  Double dispatch
-\end{itemize}
+Transformed data generally requires redrawing the plot, and worsen the
+performance of shiny. This shiny module directly manipulate the
+underlying data with the \texttt{plotly.js} react method. Only
+transformed data is sent to the server side, while keeping the rest
+configuration unchanged (e.g.~layout and graphical elements), for the
+performance reason. Users will not experience an obvious delay when the
+slider input is changed.
 

### 2020-07-30 tsibbletalk.Rmd
minor tweaks

--- 2020-07-28 tsibbletalk.Rmd
@@
 * Nesting variables generate hierarchical tree, hence `plotly_key_tree()`. Overview and navigation made easier.
-* One-to-many linking: marking a single point of interest highlights all other points that share particular data values (connect-type)
+* One-to-many linking: marking a single point of interest highlights all other points that share particular data values (categorical linking)
 * Lists of key values, and json
-* R6 subclass of `SharedData` from {crosstalk}
+* R6 subclass of mutaframes `SharedData` from {crosstalk} with metadata attached
+* Self-linking: a single data source
 
@@
 The {tsibbletalk} package provides a pair of UI and server functions, as a shiny module, to help with finding interesting time slices in a shiny application.
-The pair, `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece.
+The pair, decoupled to `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece.
 Like all shiny modules, users should supply a unique session id.
@@
 
-* Transform and send newly-transformed data via server, without updating anything else (e.g. layout and graphical elements), to avoid completely redrawing. Performance.
-* Double dispatch
+(Data flows) Transformed data generally requires redrawing the plot, and worsen the performance of shiny.
+The underlying tsibble data is called back and transformed in R.
+Using the `plotly.js` react method, only transformed data is sent to the server side, while keeping the rest configuration unchanged (e.g.
+layout and graphical elements).
+It is performant, and users will not experience notable delay in response to the change in the slider input.
+Dissect time index, and propagate transformed data to shiny server.
 

### 2020-07-30 tsibbletalk.tex
minor tweaks

--- 2020-07-28 tsibbletalk.tex
@@
   One-to-many linking: marking a single point of interest highlights all
-  other points that share particular data values (connect-type)
+  other points that share particular data values (categorical linking)
 \item
@@
 \item
-  R6 subclass of \texttt{SharedData} from \{crosstalk\}
+  R6 subclass of mutaframes \texttt{SharedData} from \{crosstalk\} with
+  metadata attached
+\item
+  Self-linking: a single data source
 \end{itemize}
@@
 as a shiny module, to help with finding interesting time slices in a
-shiny application. The pair, \texttt{tsibbleDiceUI()} and
+shiny application. The pair, decoupled to \texttt{tsibbleDiceUI()} and
 \texttt{tsibbleDiceServer()}, presents a clean interface and forms a
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Transform and send newly-transformed data via server, without updating
-  anything else (e.g.~layout and graphical elements), to avoid
-  completely redrawing. Performance.
-\item
-  Double dispatch
-\end{itemize}
+(Data flows) Transformed data generally requires redrawing the plot, and
+worsen the performance of shiny. The underlying tsibble data is called
+back and transformed in R. Using the \texttt{plotly.js} react method,
+only transformed data is sent to the server side, while keeping the rest
+configuration unchanged (e.g.~layout and graphical elements). It is
+performant, and users will not experience notable delay in response to
+the change in the slider input. Dissect time index, and propagate
+transformed data to shiny server.
 

### 2020-08-02 tsibbletalk.Rmd
draft on as_shared_tsibble()

@@
 
-* Symbolic formula to express structural specifications among keyed units, using `/` and `*` from Wilkinson notation (10.2307/2346786) for nesting and crossing.
+The \pkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble data structure to enable reactivity for temporal data (i.e.
+a `tsibble`).
+The `as_shared_tsibble()` function turns a tsibble to a mutable tsibble `SharedTsibbleData`, as an R6 subclass of `SharedData` from {crosstalk}.
+A *tsibble* data object holds a set of interrelated series, identified by the "key", in a data frame.
+When the "key" includes more than two variables, there can exist intrinsic structures in all combinations of variables: either nesting or crossing.
+(Add an example here)
+
+The `spec` argument in the `as_shared_tsibble()` function takes a symbolic formula to allow users to specify structural combinations, using `/` and `*` adopted from Wilkinson notation (10.2307/2346786) for nesting and crossing respectively.
+
 * Nesting variables generate hierarchical tree, hence `plotly_key_tree()`. Overview and navigation made easier.
@@
 * Lists of key values, and json
-* R6 subclass of mutaframes `SharedData` from {crosstalk} with metadata attached
 * Self-linking: a single data source

### 2020-08-02 tsibbletalk.tex
draft on as_shared_tsibble()

@@
 % !TeX root = RJwrapper.tex
-\title{Conversations in time: interactive visualization to explore
-structured temporal data}
+\title{Conversations in time: interactive visualization to explore structured
+temporal data}
 \author{by Earo Wang, Dianne Cook}
@@
 
+The \pkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk}
+package, introduces a shared tsibble data structure to enable reactivity
+for temporal data (i.e.~a \texttt{tsibble}). The
+\texttt{as\_shared\_tsibble()} function turns a tsibble to a mutable
+tsibble \texttt{SharedTsibbleData}, as an R6 subclass of
+\texttt{SharedData} from \{crosstalk\}. A \emph{tsibble} data object
+holds a set of interrelated series, identified by the ``key'', in a data
+frame. When the ``key'' includes more than two variables, there can
+exist intrinsic structures in all combinations of variables: either
+nesting or crossing. (Add an example here)
+
+The \texttt{spec} argument in the \texttt{as\_shared\_tsibble()}
+function takes a symbolic formula to allow users to specify structural
+combinations, using \texttt{/} and \texttt{*} adopted from Wilkinson
+notation (10.2307/2346786) for nesting and crossing respectively.
+
 \begin{itemize}
 \tightlist
-\item
-  Symbolic formula to express structural specifications among keyed
-  units, using \texttt{/} and \texttt{*} from Wilkinson notation
-  (10.2307/2346786) for nesting and crossing.
 \item
@@
   Lists of key values, and json
-\item
-  R6 subclass of mutaframes \texttt{SharedData} from \{crosstalk\} with
-  metadata attached
 \item
@@
 Monash Univerisity\\
-Department of Econometrics and Business Statistics\\ Monash
-University\\ Australia\\
+Department of Econometrics and Business Statistics\\ Monash University\\ Australia\\
 }

### 2020-08-04 tsibbletalk.Rmd
spelling

@@
 - name: Earo Wang
-  affiliation: The Univeristy of Auckland
+  affiliation: The University of Auckland
   address:
@@
 - name: Dianne Cook
-  affiliation: Monash Univerisity
+  affiliation: Monash University
   address:

### 2020-08-04 tsibbletalk.tex
spelling

@@
 Earo Wang\\
-The Univeristy of Auckland\\
+The University of Auckland\\
 Department of Statistics\\ The Univeristy of Auckland\\ New Zealand\\
@@
 Dianne Cook\\
-Monash Univerisity\\
+Monash University\\
 Department of Econometrics and Business Statistics\\ Monash University\\ Australia\\

### 2020-08-04 tsibbletalk.Rmd
minor tweaks

--- 2020-08-02 tsibbletalk.Rmd
@@
 
-The `spec` argument in the `as_shared_tsibble()` function takes a symbolic formula to allow users to specify structural combinations, using `/` and `*` adopted from Wilkinson notation (10.2307/2346786) for nesting and crossing respectively.
+The `spec` argument in the `as_shared_tsibble()` function takes a symbolic formula to allow users to specify structured combinations, using `/` and `*` adopted from Wilkinson notation (10.2307/2346786) for nesting and crossing expressions respectively.
+By default, the specification assumes crossing structure for all units.
+
+Specifying the structure is particularly useful when there's nesting structure in the data.
 

### 2020-08-04 tsibbletalk.tex
minor tweaks

--- 2020-08-02 tsibbletalk.tex
@@
 The \texttt{spec} argument in the \texttt{as\_shared\_tsibble()}
-function takes a symbolic formula to allow users to specify structural
+function takes a symbolic formula to allow users to specify structured
 combinations, using \texttt{/} and \texttt{*} adopted from Wilkinson
-notation (10.2307/2346786) for nesting and crossing respectively.
+notation (10.2307/2346786) for nesting and crossing expressions
+respectively. By default, the specification assumes crossing structure
+for all units.
+
+Specifying the structure is particularly useful when there's nesting
+structure in the data.
 

### 2020-08-06 tsibbletalk.Rmd
draft on `plotly_key_tree()`

@@
   - Department of Statistics
-  - The Univeristy of Auckland
+  - The Univerisity of Auckland
   - New Zealand
@@
 
-Specifying the structure is particularly useful when there's nesting structure in the data.
+Specifying the structure is particularly useful, when there's a hierarchical structure in the data.
+For such case, a new function `plotly_key_tree()` can help to visualise the "key" structure as a dendrogram, using plotly.
+This type of plot gives an overview of structured information in the data.
+Owing to its interactivity, each node in the tree plot is clickable and linked to other plots, and in turn the navigation at various levels is made easier.
 
-* Nesting variables generate hierarchical tree, hence `plotly_key_tree()`. Overview and navigation made easier.
+When the key structure involves crossing only, one-to-one linking is the default.
+
 * One-to-many linking: marking a single point of interest highlights all other points that share particular data values (categorical linking)
-* Lists of key values, and json
+* Constructing recursive lists. Lists of key values, and json
 * Self-linking: a single data source

### 2020-08-06 tsibbletalk.tex
draft on `plotly_key_tree()`

@@
 
-Specifying the structure is particularly useful when there's nesting
-structure in the data.
+Specifying the structure is particularly useful, when there's a
+hierarchical structure in the data. For such case, a new function
+\texttt{plotly\_key\_tree()} can help to visualise the ``key'' structure
+as a dendrogram, using plotly. This type of plot gives an overview of
+structured information in the data. Owing to its interactivity, each
+node in the tree plot is clickable and linked to other plots, and in
+turn the navigation at various levels is made easier.
+
+When the key structure involves crossing only, one-to-one linking is the
+default.
 
@@
 \item
-  Nesting variables generate hierarchical tree, hence
-  \texttt{plotly\_key\_tree()}. Overview and navigation made easier.
-\item
   One-to-many linking: marking a single point of interest highlights all
@@
 \item
-  Lists of key values, and json
+  Constructing recursive lists. Lists of key values, and json
 \item
@@
 The University of Auckland\\
-Department of Statistics\\ The Univeristy of Auckland\\ New Zealand\\
+Department of Statistics\\ The Univerisity of Auckland\\ New Zealand\\
 }

### 2020-08-07 tsibbletalk.Rmd
rewrote tsibble data

@@
 
-The \CRANpkg{tsibble} package extends the `data.frame` and \CRANpkg{tibble} structure to represent temporal data in tidy format [cite].
-A tsibble consists of *index*, *key*, and other measured variables in one data frame.
-The *index* column holds time-based indices.
-The *key* column(s) uniquely identifies a collection of related observational units during a period of time defined by *index*.
-They are "sticky" columns to a `tsibble` over the course of transformation.
+The \CRANpkg{tsibble} package [@wang2020tsibble] introduces a unified temporal data structure, referred to as a `tsibble`, to represent time series and longitudinal data in a tidy format [@wickham2014tidy].
+That said, a `tsibble` extends the `data.frame` and \CRANpkg{tibble} class with temporally contextual metadata: `index` and `key`.
+The `index` declares a data column that holds time-related indices.
+The `key` identifies a collection of related series or panels observed over the `index`-defined period, which can contain multiple columns.
+Below displays the monthly Australian retail trade turnover data, available in the \CRANpkg{tsibbledata} package.
+The `Month` column holds year-months as `index`.
+The `State` together with `Industry` are the identifiers for these 152 series, highlighted as `key`.
+Note that the column `Series ID` could be an alternative option for setting up `key`, but `State` and `Industry` are informative.
+The `index` and `key` are "sticky" columns to a `tsibble`, forming critical pieces for fluent temporal data analysis later.
+
+```{r, aus-retail}
+library(tsibble)
+library(tsibbledata)
+print(aus_retail, n = 5)
+```
 

### 2020-08-07 tsibbletalk.tex
rewrote tsibble data

@@
 
-The \CRANpkg{tsibble} package extends the \texttt{data.frame} and
-\CRANpkg{tibble} structure to represent temporal data in tidy format
-{[}cite{]}. A tsibble consists of \emph{index}, \emph{key}, and other
-measured variables in one data frame. The \emph{index} column holds
-time-based indices. The \emph{key} column(s) uniquely identifies a
-collection of related observational units during a period of time
-defined by \emph{index}. They are ``sticky'' columns to a
-\texttt{tsibble} over the course of transformation.
+The \CRANpkg{tsibble} package \citep{wang2020tsibble} introduces a
+unified temporal data structure, referred to as a \texttt{tsibble}, to
+represent time series and longitudinal data in a tidy format
+\citep{wickham2014tidy}. That said, a \texttt{tsibble} extends the
+\texttt{data.frame} and \CRANpkg{tibble} class with temporally
+contextual metadata: \texttt{index} and \texttt{key}. The \texttt{index}
+declares a data column that holds time-related indices. The \texttt{key}
+identifies a collection of related series or panels observed over the
+\texttt{index}-defined period, which can contain multiple columns. Below
+displays the monthly Australian retail trade turnover data, available in
+the \CRANpkg{tsibbledata} package. The \texttt{Month} column holds
+year-months as \texttt{index}. The \texttt{State} together with
+\texttt{Industry} are the identifiers for these 152 series, highlighted
+as \texttt{key}. Note that the column \texttt{Series\ ID} could be an
+alternative option for setting up \texttt{key}, but \texttt{State} and
+\texttt{Industry} are informative. The \texttt{index} and \texttt{key}
+are ``sticky'' columns to a \texttt{tsibble}, forming critical pieces
+for fluent temporal data analysis.
+
+\begin{Schunk}
+\begin{Soutput}
+#> # A tsibble: 64,532 x 5 [1M]
+#> # Key:       State, Industry [152]
+#>   State                Industry                    `Series ID`    Month Turnover
+#>   <chr>                <chr>                       <chr>          <mth>    <dbl>
+#> 1 Australian Capital ~ Cafes, restaurants and cat~ A3349849A   1982 Apr      4.4
+#> 2 Australian Capital ~ Cafes, restaurants and cat~ A3349849A   1982 May      3.4
+#> 3 Australian Capital ~ Cafes, restaurants and cat~ A3349849A   1982 Jun      3.6
+#> 4 Australian Capital ~ Cafes, restaurants and cat~ A3349849A   1982 Jul      4  
+#> 5 Australian Capital ~ Cafes, restaurants and cat~ A3349849A   1982 Aug      3.6
+#> # ... with 64,527 more rows
+\end{Soutput}
+\end{Schunk}
 

### 2020-08-10 tsibbletalk.Rmd
reworked on tidyverts intro

@@
 ---
-title: "Conversations in time: interactive visualization to explore structured temporal data"
+title: "Conversations in time: interactive visualisation to explore structured temporal data"
 author:
@@
 )
-knitr::opts_knit$set(root.dir = here::here())
+opts_knit$set(root.dir = here::here())
+read_chunk("scripts/demo.R")
+```
+
+```{r, load-pkgs}
 ```
@@
 
-## Interactive visualization systems with linking views
+## Interactive visualisation systems with linking views
 
@@
 The `State` together with `Industry` are the identifiers for these 152 series, highlighted as `key`.
-Note that the column `Series ID` could be an alternative option for setting up `key`, but `State` and `Industry` are informative.
+Note that the column `Series ID` could be an alternative option for setting up `key`, but `State` and `Industry` are more readable and informative.
 The `index` and `key` are "sticky" columns to a `tsibble`, forming critical pieces for fluent temporal data analysis later.
 
-```{r, aus-retail}
-library(tsibble)
-library(tsibbledata)
-print(aus_retail, n = 5)
+```{r, print-retail}
 ```
 
-The \CRANpkg{feasts} and \CRANpkg{fable} packages, part of the **tidyverts** suite, aim to make time series analysis easier.
-They provide analytical and forecasting tools for the tsibble data structure, generating tsibble-centered workflow.
+In the spirit of tidy data to the \CRANpkg{tidyverse} [@Wickham2019], the **tidyverts** suite features `tsibble` as the foundational data structure, to build a fluid and fluent pipeline for time series analysis.
+Besides \CRANpkg{tsibble}, the \CRANpkg{feasts} and \CRANpkg{fable} packages fill the role of statistical analysis and forecasting in the **tidyverts** ecosystem.
+When time series analysis starts taking off, series of interests defined by `key` are likely to remain unchanged over the course of analysis.
 Functions, such as `features()` and `model()`, summarise a sequence of indexed values down to a single statistic or model by every observational unit.
@@
 In the context of relational databases, the "key" acts like a foreign key in a reduced form of tsibble, while the index and key together operates like a primary key.
+(tsibble-centered workflow)
+
+```{r highlight-retail, fig.height = 3.3, fig.cap = "ToDo"}
+```
 

### 2020-08-10 tsibbletalk.tex
reworked on tidyverts intro

@@
 % !TeX root = RJwrapper.tex
-\title{Conversations in time: interactive visualization to explore structured
+\title{Conversations in time: interactive visualisation to explore structured
 temporal data}
@@
 
-\hypertarget{interactive-visualization-systems-with-linking-views}{%
-\subsection{Interactive visualization systems with linking
-views}\label{interactive-visualization-systems-with-linking-views}}
+\hypertarget{interactive-visualisation-systems-with-linking-views}{%
+\subsection{Interactive visualisation systems with linking
+views}\label{interactive-visualisation-systems-with-linking-views}}
 
@@
 alternative option for setting up \texttt{key}, but \texttt{State} and
-\texttt{Industry} are informative. The \texttt{index} and \texttt{key}
-are ``sticky'' columns to a \texttt{tsibble}, forming critical pieces
-for fluent temporal data analysis.
+\texttt{Industry} are more readable and informative. The \texttt{index}
+and \texttt{key} are ``sticky'' columns to a \texttt{tsibble}, forming
+critical pieces for fluent temporal data analysis later.
 
@@
 
-The \CRANpkg{feasts} and \CRANpkg{fable} packages, part of the
-\textbf{tidyverts} suite, aim to make time series analysis easier. They
-provide analytical and forecasting tools for the tsibble data structure,
-generating tsibble-centered workflow. Functions, such as
-\texttt{features()} and \texttt{model()}, summarise a sequence of
-indexed values down to a single statistic or model by every
-observational unit. The output is a normal table, where each row
-corresponds to an observational unit denoted by ``key''. In the context
-of relational databases, the ``key'' acts like a foreign key in a
-reduced form of tsibble, while the index and key together operates like
-a primary key.
+In the spirit of tidy data to the \CRANpkg{tidyverse}
+\citep{Wickham2019}, the \textbf{tidyverts} suite features
+\texttt{tsibble} as the foundational data structure, to build a fluid
+and fluent pipeline for time series analysis. Besides \CRANpkg{tsibble},
+the \CRANpkg{feasts} and \CRANpkg{fable} packages fill the role of
+statistical analysis and forecasting in the \textbf{tidyverts}
+ecosystem. When time series analysis starts taking off, series of
+interests defined by \texttt{key} are likely to remain unchanged over
+the course of analysis. Functions, such as \texttt{features()} and
+\texttt{model()}, summarise a sequence of indexed values down to a
+single statistic or model by every observational unit. The output is a
+normal table, where each row corresponds to an observational unit
+denoted by ``key''. In the context of relational databases, the ``key''
+acts like a foreign key in a reduced form of tsibble, while the index
+and key together operates like a primary key. (tsibble-centered
+workflow)
+
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/highlight-retail-1}
+
+}
+
+\caption[ToDo]{ToDo}\label{fig:highlight-retail}
+\end{figure}
+\end{Schunk}
 

### 2020-08-10 scripts/demo.R

@@ -1,0 +1,38 @@
+## ---- load-pkgs
+library(feasts)
+library(tsibble)
+library(tsibbledata)
+
+## ---- print-retail
+print(aus_retail, n = 5)
+
+## ---- highlight-retail
+library(tidyverse)
+library(ggrepel)
+library(gghighlight)
+library(patchwork)
+
+p_retail_ts <- aus_retail %>% 
+  as_tibble() %>% # gghighlight issue for group_by() + filter()
+  mutate(group = paste(State, ":", Industry)) %>% 
+  ggplot(aes(x = Month, y = Turnover)) +
+  geom_line(aes(group = group)) +
+  gghighlight(
+    State == "Queensland", Industry == "Department stores",
+    label_params = list(vjust = -2)
+  ) +
+  scale_x_yearmonth(breaks = yearmonth(c("1990 Jan", "2000 Jan", "2010 Jan")))
+
+p_retail_feat <- aus_retail %>% 
+  features(Turnover, feat_stl) %>% 
+  mutate(group = paste(State, ":", Industry)) %>% 
+  ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
+  geom_point(aes(group = group)) +
+  labs(x = "Trend strength", y = "Seasonal strength") +
+  gghighlight(
+    State == "Queensland", Industry == "Department stores",
+    use_direct_label = FALSE
+  ) +
+  geom_label_repel(aes(label = group), vjust = 2, nudge_x = -0.1)
+
+p_retail_ts + p_retail_feat + plot_annotation(tag_levels = "a")

### 2020-08-11 tsibbletalk.Rmd
done with tidy temporal data review

@@
 The `key` identifies a collection of related series or panels observed over the `index`-defined period, which can contain multiple columns.
-Below displays the monthly Australian retail trade turnover data, available in the \CRANpkg{tsibbledata} package.
+Below displays the monthly Australian retail trade turnover data (`aus_retail`), available in the \CRANpkg{tsibbledata} package.
 The `Month` column holds year-months as `index`.
@@
 
-In the spirit of tidy data to the \CRANpkg{tidyverse} [@Wickham2019], the **tidyverts** suite features `tsibble` as the foundational data structure, to build a fluid and fluent pipeline for time series analysis.
+In the spirit of tidy data to the \CRANpkg{tidyverse} [@Wickham2019], the **tidyverts** suite features `tsibble` as the foundational data structure, in order to build a fluid and fluent pipeline for time series analysis.
 Besides \CRANpkg{tsibble}, the \CRANpkg{feasts} and \CRANpkg{fable} packages fill the role of statistical analysis and forecasting in the **tidyverts** ecosystem.
-When time series analysis starts taking off, series of interests defined by `key` are likely to remain unchanged over the course of analysis.
-Functions, such as `features()` and `model()`, summarise a sequence of indexed values down to a single statistic or model by every observational unit.
-The output is a normal table, where each row corresponds to an observational unit denoted by "key".
-In the context of relational databases, the "key" acts like a foreign key in a reduced form of tsibble, while the index and key together operates like a primary key.
-(tsibble-centered workflow)
+When time series analysis starts taking off, series of interest denoted by the `key` variables often remain unchanged over the course of analysis, from trend inspection to forecasting performance.
+Figure \ref{fig:highlight-retail} (a) gives an overview of 152 series for the retail data using an overlaid time series plot, while Figure \ref{fig:highlight-retail} (b) presents a scatterplot, where each series is represented by a dot in the feature space (trend versus seasonal strength).
+The plot making of Figure \ref{fig:highlight-retail} (b) is aided with the `features()` function from \CRANpkg{feasts}, which summarises original data by each series down to various statistical features.
+This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, thus graphically displayed as Figure \ref{fig:highlight-retail} (b).
 
@@
 
-At the early stage of exploratory temporal data analysis, time series plots and scatterplots goes hand by hand. (insert figures below)
+Figure \ref{fig:highlight-retail} highlights not only a series with strongest seasonality, but also a need to querying interesting series on the fly.
+The two plots are static, but can be considered as linked views via the common `key` variables between two tables.
+This motivates enabling interactivity of `tsibble` and `tsibble`-derived objects for exploratory data analysis in a lightning manner.
 
@@
 A *tsibble* data object holds a set of interrelated series, identified by the "key", in a data frame.
-When the "key" includes more than two variables, there can exist intrinsic structures in all combinations of variables: either nesting or crossing.
+In the context of relational databases, the "key" acts like a foreign key in a reduced form of tsibble, while the index and key together operates like a primary key.
+(tsibble-centered workflow) When the "key" includes more than two variables, there can exist intrinsic structures in all combinations of variables: either nesting or crossing.
 (Add an example here)

### 2020-08-11 tsibbletalk.tex
done with tidy temporal data review

@@
 \texttt{index}-defined period, which can contain multiple columns. Below
-displays the monthly Australian retail trade turnover data, available in
-the \CRANpkg{tsibbledata} package. The \texttt{Month} column holds
-year-months as \texttt{index}. The \texttt{State} together with
-\texttt{Industry} are the identifiers for these 152 series, highlighted
-as \texttt{key}. Note that the column \texttt{Series\ ID} could be an
-alternative option for setting up \texttt{key}, but \texttt{State} and
-\texttt{Industry} are more readable and informative. The \texttt{index}
-and \texttt{key} are ``sticky'' columns to a \texttt{tsibble}, forming
-critical pieces for fluent temporal data analysis later.
+displays the monthly Australian retail trade turnover data
+(\texttt{aus\_retail}), available in the \CRANpkg{tsibbledata} package.
+The \texttt{Month} column holds year-months as \texttt{index}. The
+\texttt{State} together with \texttt{Industry} are the identifiers for
+these 152 series, highlighted as \texttt{key}. Note that the column
+\texttt{Series\ ID} could be an alternative option for setting up
+\texttt{key}, but \texttt{State} and \texttt{Industry} are more readable
+and informative. The \texttt{index} and \texttt{key} are ``sticky''
+columns to a \texttt{tsibble}, forming critical pieces for fluent
+temporal data analysis later.
 
@@
 \citep{Wickham2019}, the \textbf{tidyverts} suite features
-\texttt{tsibble} as the foundational data structure, to build a fluid
-and fluent pipeline for time series analysis. Besides \CRANpkg{tsibble},
-the \CRANpkg{feasts} and \CRANpkg{fable} packages fill the role of
-statistical analysis and forecasting in the \textbf{tidyverts}
-ecosystem. When time series analysis starts taking off, series of
-interests defined by \texttt{key} are likely to remain unchanged over
-the course of analysis. Functions, such as \texttt{features()} and
-\texttt{model()}, summarise a sequence of indexed values down to a
-single statistic or model by every observational unit. The output is a
-normal table, where each row corresponds to an observational unit
-denoted by ``key''. In the context of relational databases, the ``key''
-acts like a foreign key in a reduced form of tsibble, while the index
-and key together operates like a primary key. (tsibble-centered
-workflow)
+\texttt{tsibble} as the foundational data structure, in order to build a
+fluid and fluent pipeline for time series analysis. Besides
+\CRANpkg{tsibble}, the \CRANpkg{feasts} and \CRANpkg{fable} packages
+fill the role of statistical analysis and forecasting in the
+\textbf{tidyverts} ecosystem. When time series analysis starts taking
+off, series of interest denoted by the \texttt{key} variables often
+remain unchanged over the course of analysis, from trend inspection to
+forecasting performance. Figure \ref{fig:highlight-retail} (a) gives an
+overview of 152 series for the retail data using an overlaid time series
+plot, while Figure \ref{fig:highlight-retail} (b) presents a
+scatterplot, where each series is represented by a dot in the feature
+space (trend versus seasonal strength). The plot making of Figure
+\ref{fig:highlight-retail} (b) is aided with the \texttt{features()}
+function from \CRANpkg{feasts}, which summarises original data by each
+series down to various statistical features. This function along with
+other \textbf{tidyverts} functions is \texttt{tsibble}-aware, and
+outputs a table in a reduced form where each row corresponds to a
+series, thus graphically displayed as Figure \ref{fig:highlight-retail}
+(b).
 
@@
 
-At the early stage of exploratory temporal data analysis, time series
-plots and scatterplots goes hand by hand. (insert figures below)
+Figure \ref{fig:highlight-retail} highlights not only a series with
+strongest seasonality, but also a need to querying interesting series on
+the fly. The two plots are static, but can be considered as linked views
+via the common \texttt{key} variables between two tables. This motivates
+enabling interactivity of \texttt{tsibble} and \texttt{tsibble}-derived
+objects for exploratory data analysis in a lightning manner.
 
@@
 holds a set of interrelated series, identified by the ``key'', in a data
-frame. When the ``key'' includes more than two variables, there can
-exist intrinsic structures in all combinations of variables: either
-nesting or crossing. (Add an example here)
+frame. In the context of relational databases, the ``key'' acts like a
+foreign key in a reduced form of tsibble, while the index and key
+together operates like a primary key. (tsibble-centered workflow) When
+the ``key'' includes more than two variables, there can exist intrinsic
+structures in all combinations of variables: either nesting or crossing.
+(Add an example here)
 

### 2020-08-12 tsibbletalk.Rmd
minor tweak
added a bit detailed description to figure 1

@@
 
-Figure \ref{fig:highlight-retail} highlights not only a series with strongest seasonality, but also a need to querying interesting series on the fly.
-The two plots are static, but can be considered as linked views via the common `key` variables between two tables.
-This motivates enabling interactivity of `tsibble` and `tsibble`-derived objects for exploratory data analysis in a lightning manner.
+Figure \ref{fig:highlight-retail} highlights not only a series with strongest seasonality, but also a need to querying interesting series on the go.
+Without interactivity, one needs to first filter the interesting series out from the features table, and join back to the original `tsibble` in order to examine its trend in relation to others.
+This procedure can soon grow cumbersome if many series to be discovered.
+Despite that the two plots are static, they can be considered as linked views via the common `key` variables between two tables.
+This motivates enabling interactivity of `tsibble` and `tsibble`-derived objects for rapid exploratory data analysis.
 

### 2020-08-12 tsibbletalk.tex
minor tweak
added a bit detailed description to figure 1

@@
 % !TeX root = RJwrapper.tex
-\title{Conversations in time: interactive visualisation to explore structured
-temporal data}
+\title{Conversations in time: interactive visualisation to explore
+structured temporal data}
 \author{by Earo Wang, Dianne Cook}
@@
 strongest seasonality, but also a need to querying interesting series on
-the fly. The two plots are static, but can be considered as linked views
-via the common \texttt{key} variables between two tables. This motivates
-enabling interactivity of \texttt{tsibble} and \texttt{tsibble}-derived
-objects for exploratory data analysis in a lightning manner.
+the go. Without interactivity, one needs to first filter the interesting
+series out from the features table, and join back to the original
+\texttt{tsibble} in order to examine its trend in relation to others.
+This procedure can soon grow cumbersome if many series to be discovered.
+Despite that the two plots are static, they can be considered as linked
+views via the common \texttt{key} variables between two tables. This
+motivates enabling interactivity of \texttt{tsibble} and
+\texttt{tsibble}-derived objects for rapid exploratory data analysis.
 
@@
 Monash University\\
-Department of Econometrics and Business Statistics\\ Monash University\\ Australia\\
+Department of Econometrics and Business Statistics\\ Monash
+University\\ Australia\\
 }

### 2020-08-13 tsibbletalk.Rmd
reworked on categorical linking

@@
 The `index` declares a data column that holds time-related indices.
-The `key` identifies a collection of related series or panels observed over the `index`-defined period, which can contain multiple columns.
+The `key` identifies a collection of related series or panels observed over the `index`-defined period, which can comprise multiple columns.
 Below displays the monthly Australian retail trade turnover data (`aus_retail`), available in the \CRANpkg{tsibbledata} package.
@@
 
-Figure \ref{fig:highlight-retail} highlights not only a series with strongest seasonality, but also a need to querying interesting series on the go.
+Figure \ref{fig:highlight-retail} highlights not only a series with strongest seasonality, but also a need to querying interesting series on the fly.
 Without interactivity, one needs to first filter the interesting series out from the features table, and join back to the original `tsibble` in order to examine its trend in relation to others.
@@
 
-The \pkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble data structure to enable reactivity for temporal data (i.e.
-a `tsibble`).
-The `as_shared_tsibble()` function turns a tsibble to a mutable tsibble `SharedTsibbleData`, as an R6 subclass of `SharedData` from {crosstalk}.
-A *tsibble* data object holds a set of interrelated series, identified by the "key", in a data frame.
-In the context of relational databases, the "key" acts like a foreign key in a reduced form of tsibble, while the index and key together operates like a primary key.
-(tsibble-centered workflow) When the "key" includes more than two variables, there can exist intrinsic structures in all combinations of variables: either nesting or crossing.
-(Add an example here)
+The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble data structure on top of a `tsibble` to allow for frictionless communication between different plots for temporal data.
+The `as_shared_tsibble()` function provides an entry point in the flow, turning a `tsibble` to a shared data object (i.e.
+`SharedTsibbleData` subclassing of `SharedData` from \CRANpkg{crosstalk}) that powers data transmission between plots.
+The \CRANpkg{tsibbletalk} package aims to streamline interactive graphical analysis with the focus of temporal and structured linking.
 
+As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical linking where marking one or more observations in one category will broadcast to all other observations in this category.
+Given time series plots, click any data point on a line, highlighting the whole line as a result.
+The `as_shared_tsibble()` uses `tsibble`'s `key` data values to facilitate such linking.
 The `spec` argument in the `as_shared_tsibble()` function takes a symbolic formula to allow users to specify structured combinations, using `/` and `*` adopted from Wilkinson notation (10.2307/2346786) for nesting and crossing expressions respectively.
@@
 
-When the key structure involves crossing only, one-to-one linking is the default.
-
-* One-to-many linking: marking a single point of interest highlights all other points that share particular data values (categorical linking)
 * Constructing recursive lists. Lists of key values, and json

### 2020-08-13 tsibbletalk.tex
reworked on categorical linking

--- 2020-08-11 tsibbletalk.tex
@@
 identifies a collection of related series or panels observed over the
-\texttt{index}-defined period, which can contain multiple columns. Below
-displays the monthly Australian retail trade turnover data
+\texttt{index}-defined period, which can comprise multiple columns.
+Below displays the monthly Australian retail trade turnover data
 (\texttt{aus\_retail}), available in the \CRANpkg{tsibbledata} package.
@@
 strongest seasonality, but also a need to querying interesting series on
-the fly. The two plots are static, but can be considered as linked views
-via the common \texttt{key} variables between two tables. This motivates
-enabling interactivity of \texttt{tsibble} and \texttt{tsibble}-derived
-objects for exploratory data analysis in a lightning manner.
+the fly. Without interactivity, one needs to first filter the
+interesting series out from the features table, and join back to the
+original \texttt{tsibble} in order to examine its trend in relation to
+others. This procedure can soon grow cumbersome if many series to be
+discovered. Despite that the two plots are static, they can be
+considered as linked views via the common \texttt{key} variables between
+two tables. This motivates enabling interactivity of \texttt{tsibble}
+and \texttt{tsibble}-derived objects for rapid exploratory data
+analysis.
 
@@
 
-The \pkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk}
-package, introduces a shared tsibble data structure to enable reactivity
-for temporal data (i.e.~a \texttt{tsibble}). The
-\texttt{as\_shared\_tsibble()} function turns a tsibble to a mutable
-tsibble \texttt{SharedTsibbleData}, as an R6 subclass of
-\texttt{SharedData} from \{crosstalk\}. A \emph{tsibble} data object
-holds a set of interrelated series, identified by the ``key'', in a data
-frame. In the context of relational databases, the ``key'' acts like a
-foreign key in a reduced form of tsibble, while the index and key
-together operates like a primary key. (tsibble-centered workflow) When
-the ``key'' includes more than two variables, there can exist intrinsic
-structures in all combinations of variables: either nesting or crossing.
-(Add an example here)
+The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk}
+package, introduces a shared tsibble data structure on top of a
+\texttt{tsibble} to allow for frictionless communication between
+different plots for temporal data. The \texttt{as\_shared\_tsibble()}
+function provides an entry point in the flow, turning a \texttt{tsibble}
+to a shared data object (i.e.~\texttt{SharedTsibbleData} subclassing of
+\texttt{SharedData} from \CRANpkg{crosstalk}) that powers data
+transmission between plots. The \CRANpkg{tsibbletalk} package aims to
+streamline interactive graphical analysis with the focus of temporal and
+structured linking.
 
+As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to
+categorical linking where marking one or more observations in one
+category will broadcast to all other observations in this category.
+Given time series plots, click any data point on a line, highlighting
+the whole line as a result. The \texttt{as\_shared\_tsibble()} uses
+\texttt{tsibble}'s \texttt{key} data values to facilitate such linking.
 The \texttt{spec} argument in the \texttt{as\_shared\_tsibble()}
@@
 
-When the key structure involves crossing only, one-to-one linking is the
-default.
-
 \begin{itemize}
 \tightlist
-\item
-  One-to-many linking: marking a single point of interest highlights all
-  other points that share particular data values (categorical linking)
 \item

### 2020-08-18 tsibbletalk.Rmd
nesting and crossing syntax

@@
 
-# Background and motivation
-
-## Interactive visualisation systems with linking views
+# Related works: interactive visualisation systems with linking views
 
@@
 
-## Tidy temporal data and workflow
+# Motivation: tidy temporal data and workflow
 
@@
 
-# Shared temporal data for coordinated views
+# Shared temporal data for coordinated views with examples
 
 The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble data structure on top of a `tsibble` to allow for frictionless communication between different plots for temporal data.
-The `as_shared_tsibble()` function provides an entry point in the flow, turning a `tsibble` to a shared data object (i.e.
-`SharedTsibbleData` subclassing of `SharedData` from \CRANpkg{crosstalk}) that powers data transmission between plots.
+The `as_shared_tsibble()` function provides an entry point in the integrated flow, turning a `tsibble` to a shared instance (i.e.
+`SharedTsibbleData` subclassing of `SharedData` from \CRANpkg{crosstalk}) that powers data transmission across multiple views.
 The \CRANpkg{tsibbletalk} package aims to streamline interactive graphical analysis with the focus of temporal and structured linking.
@@
 Given time series plots, click any data point on a line, highlighting the whole line as a result.
-The `as_shared_tsibble()` uses `tsibble`'s `key` data values to facilitate such linking.
-The `spec` argument in the `as_shared_tsibble()` function takes a symbolic formula to allow users to specify structured combinations, using `/` and `*` adopted from Wilkinson notation (10.2307/2346786) for nesting and crossing expressions respectively.
-By default, the specification assumes crossing structure for all units.
+The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking, and the `spec` argument takes one step further in constructing hybrid linking, such as hierarchical and categorical linking.
+For example, each series in the `aus_retail` data corresponds to all possible combinations of the `State` and `Industry` variables.
+They are intrinsically crossed with each other.
+If one variable is nested within another, this lends itself to a hierarchical structure, like geographical hierarchy.
+Such collection of inter-related time series are referred to as hierarchical and grouped time series in the literature [@fpp].
+
+To incorporate structured specifications in the `key`, a symbolic formula can be passed to the `spec` argument.
+Adopting Wilkinson notations for factorial models [@Wilkinson1973], the `spec` follows the `/` and `*` operators tradition to declare nesting and crossing variables respectively.
+The `spec` for the `aus_retail` data is therefore specified as `spec = State * Industry`, which is the default for the presence of multiple `key` variables.
+Using `/` indicates if there is a hierarchy in the data.
+`tourism_monthly`
 
@@
 
-# Slicing and dicing time
+# Slicing and dicing time with examples
 
@@
 
-# Case study: monthly domestic tourist trips in Australia
-
 # Conclusions and discussions

### 2020-08-18 tsibbletalk.tex
nesting and crossing syntax

@@
 
-\hypertarget{background-and-motivation}{%
-\section{Background and motivation}\label{background-and-motivation}}
-
-\hypertarget{interactive-visualisation-systems-with-linking-views}{%
-\subsection{Interactive visualisation systems with linking
-views}\label{interactive-visualisation-systems-with-linking-views}}
+\hypertarget{related-works-interactive-visualisation-systems-with-linking-views}{%
+\section{Related works: interactive visualisation systems with linking
+views}\label{related-works-interactive-visualisation-systems-with-linking-views}}
 
@@
 
-\hypertarget{tidy-temporal-data-and-workflow}{%
-\subsection{Tidy temporal data and
-workflow}\label{tidy-temporal-data-and-workflow}}
+\hypertarget{motivation-tidy-temporal-data-and-workflow}{%
+\section{Motivation: tidy temporal data and
+workflow}\label{motivation-tidy-temporal-data-and-workflow}}
 
@@
 
-\hypertarget{shared-temporal-data-for-coordinated-views}{%
-\section{Shared temporal data for coordinated
-views}\label{shared-temporal-data-for-coordinated-views}}
+\hypertarget{shared-temporal-data-for-coordinated-views-with-examples}{%
+\section{Shared temporal data for coordinated views with
+examples}\label{shared-temporal-data-for-coordinated-views-with-examples}}
 
@@
 different plots for temporal data. The \texttt{as\_shared\_tsibble()}
-function provides an entry point in the flow, turning a \texttt{tsibble}
-to a shared data object (i.e.~\texttt{SharedTsibbleData} subclassing of
-\texttt{SharedData} from \CRANpkg{crosstalk}) that powers data
-transmission between plots. The \CRANpkg{tsibbletalk} package aims to
-streamline interactive graphical analysis with the focus of temporal and
-structured linking.
+function provides an entry point in the integrated flow, turning a
+\texttt{tsibble} to a shared instance (i.e.~\texttt{SharedTsibbleData}
+subclassing of \texttt{SharedData} from \CRANpkg{crosstalk}) that powers
+data transmission across multiple views. The \CRANpkg{tsibbletalk}
+package aims to streamline interactive graphical analysis with the focus
+of temporal and structured linking.
 
@@
 the whole line as a result. The \texttt{as\_shared\_tsibble()} uses
-\texttt{tsibble}'s \texttt{key} data values to facilitate such linking.
-The \texttt{spec} argument in the \texttt{as\_shared\_tsibble()}
-function takes a symbolic formula to allow users to specify structured
-combinations, using \texttt{/} and \texttt{*} adopted from Wilkinson
-notation (10.2307/2346786) for nesting and crossing expressions
-respectively. By default, the specification assumes crossing structure
-for all units.
+\texttt{tsibble}'s \texttt{key} variables to achieve these types of
+linking, and the \texttt{spec} argument takes one step further in
+constructing hybrid linking, such as hierarchical and categorical
+linking. For example, each series in the \texttt{aus\_retail} data
+corresponds to all possible combinations of the \texttt{State} and
+\texttt{Industry} variables. They are intrinsically crossed with each
+other. If one variable is nested within another, this lends itself to a
+hierarchical structure, like geographical hierarchy. Such collection of
+inter-related time series are referred to as hierarchical and grouped
+time series in the literature \citep{fpp}.
+
+To incorporate structured specifications in the \texttt{key}, a symbolic
+formula can be passed to the \texttt{spec} argument. Adopting Wilkinson
+notations for factorial models \citep{Wilkinson1973}, the \texttt{spec}
+follows the \texttt{/} and \texttt{*} operators tradition to declare
+nesting and crossing variables respectively. The \texttt{spec} for the
+\texttt{aus\_retail} data is therefore specified as
+\texttt{spec\ =\ State\ *\ Industry}, which is the default for the
+presence of multiple \texttt{key} variables. Using \texttt{/} indicates
+if there is a hierarchy in the data. \texttt{tourism\_monthly}
 
@@
 
-\hypertarget{slicing-and-dicing-time}{%
-\section{Slicing and dicing time}\label{slicing-and-dicing-time}}
+\hypertarget{slicing-and-dicing-time-with-examples}{%
+\section{Slicing and dicing time with
+examples}\label{slicing-and-dicing-time-with-examples}}
 
@@
 
-\hypertarget{case-study-monthly-domestic-tourist-trips-in-australia}{%
-\section{Case study: monthly domestic tourist trips in
-Australia}\label{case-study-monthly-domestic-tourist-trips-in-australia}}
-
 \hypertarget{conclusions-and-discussions}{%

### 2020-08-18 tsibbletalk.Rmd
typo in yaml

@@
   - Department of Statistics
-  - The Univerisity of Auckland
+  - The University of Auckland
   - New Zealand

### 2020-08-18 tsibbletalk.Rmd
sections reorganised

--- 2020-08-18 tsibbletalk.Rmd
@@
 
-# Related works: interactive visualisation systems with linking views
-
-* {ggobi} and {xggobi}
-* {cranvas} and {cranvastime}
-* [crossfilter.js](http://crossfilter.github.io/crossfilter/) & [dc.js](https://dc-js.github.io/dc.js/)
-* {crosstalk} and html widgets
-
-# Motivation: tidy temporal data and workflow
+# Background: tidy temporal data and workflow
 
@@
 When time series analysis starts taking off, series of interest denoted by the `key` variables often remain unchanged over the course of analysis, from trend inspection to forecasting performance.
+
 Figure \ref{fig:highlight-retail} (a) gives an overview of 152 series for the retail data using an overlaid time series plot, while Figure \ref{fig:highlight-retail} (b) presents a scatterplot, where each series is represented by a dot in the feature space (trend versus seasonal strength).
@@
 
-# Shared temporal data for coordinated views with examples
+# Overview of interactivity
+
+* {cranvas} and {cranvastime}
+* [crossfilter.js](http://crossfilter.github.io/crossfilter/) & [dc.js](https://dc-js.github.io/dc.js/)
+* {crosstalk} and html widgets
+* {rJava}, {rbokeh} {loon}
+
+# Interactivity for coordinated views via shared temporal data
 
@@
 
-# Slicing and dicing time with examples
+# Slicing and dicing time
 

### 2020-08-18 tsibbletalk.tex
sections reorganised

@@
 
-\hypertarget{related-works-interactive-visualisation-systems-with-linking-views}{%
-\section{Related works: interactive visualisation systems with linking
-views}\label{related-works-interactive-visualisation-systems-with-linking-views}}
-
-\begin{itemize}
-\tightlist
-\item
-  \{ggobi\} and \{xggobi\}
-\item
-  \{cranvas\} and \{cranvastime\}
-\item
-  \href{http://crossfilter.github.io/crossfilter/}{crossfilter.js} \&
-  \href{https://dc-js.github.io/dc.js/}{dc.js}
-\item
-  \{crosstalk\} and html widgets
-\end{itemize}
-
-\hypertarget{motivation-tidy-temporal-data-and-workflow}{%
-\section{Motivation: tidy temporal data and
-workflow}\label{motivation-tidy-temporal-data-and-workflow}}
+\hypertarget{background-tidy-temporal-data-and-workflow}{%
+\section{Background: tidy temporal data and
+workflow}\label{background-tidy-temporal-data-and-workflow}}
 
@@
 
-\hypertarget{shared-temporal-data-for-coordinated-views-with-examples}{%
-\section{Shared temporal data for coordinated views with
-examples}\label{shared-temporal-data-for-coordinated-views-with-examples}}
+\hypertarget{overview-of-interactivity}{%
+\section{Overview of interactivity}\label{overview-of-interactivity}}
+
+\begin{itemize}
+\tightlist
+\item
+  \{cranvas\} and \{cranvastime\}
+\item
+  \href{http://crossfilter.github.io/crossfilter/}{crossfilter.js} \&
+  \href{https://dc-js.github.io/dc.js/}{dc.js}
+\item
+  \{crosstalk\} and html widgets
+\item
+  \{rJava\}, \{rbokeh\} \{loon\}
+\end{itemize}
+
+\hypertarget{interactivity-for-coordinated-views-via-shared-temporal-data}{%
+\section{Interactivity for coordinated views via shared temporal
+data}\label{interactivity-for-coordinated-views-via-shared-temporal-data}}
 
@@
 
-\hypertarget{slicing-and-dicing-time-with-examples}{%
-\section{Slicing and dicing time with
-examples}\label{slicing-and-dicing-time-with-examples}}
+\hypertarget{slicing-and-dicing-time}{%
+\section{Slicing and dicing time}\label{slicing-and-dicing-time}}
 

### 2020-08-24 tsibbletalk.Rmd
introducing plotly_key_tree()

@@
 read_chunk("scripts/demo.R")
+read_chunk("scripts/tourism.R")
 ```
@@
 Adopting Wilkinson notations for factorial models [@Wilkinson1973], the `spec` follows the `/` and `*` operators tradition to declare nesting and crossing variables respectively.
-The `spec` for the `aus_retail` data is therefore specified as `spec = State * Industry`, which is the default for the presence of multiple `key` variables.
-Using `/` indicates if there is a hierarchy in the data.
-`tourism_monthly`
+The `spec` for the `aus_retail` data is therefore specified as `State * Industry` or `Industry * State`, which is the default for the presence of multiple `key` variables.
+If there is a hierarchy in the data, using `/` is required to indicate the parent-child relation, as strictly one direction `parent/child`.
 
-Specifying the structure is particularly useful, when there's a hierarchical structure in the data.
-For such case, a new function `plotly_key_tree()` can help to visualise the "key" structure as a dendrogram, using plotly.
-This type of plot gives an overview of structured information in the data.
-Owing to its interactivity, each node in the tree plot is clickable and linked to other plots, and in turn the navigation at various levels is made easier.
+The `tourism_monthly` dataset packaged in \CRANpkg{tsibbletalk}, contains monthly domestic overnight trips across Australia, to give an illustrator of nesting and crossing.
+The `key` is comprised of three identifying variables: `State`, `Region`, and `Purpose` (of trip), in particular `State` nesting of `Region`, together crossed with `Purpose`.
+This specification can be translated as follows:
 
-* Constructing recursive lists. Lists of key values, and json
-* Self-linking: a single data source
+```{r tourism-shared, echo = TRUE}
+```
+
+```{r load-other-pkgs}
+```
+
+This dataset contains a three-level hierarchy: the root node is implicitly Australia, and geographically disaggregated to states and lower-level tourism regions.
+A new function `plotly_key_tree()` has been implemented to address the need of hierarchical discovery arising from data.
+It visually interprets hierarchies in the shared tsibble's `spec` as a tree view, made with \CRANpkg{plotly}.
+From a static point of view, a tree visualisation of hierarchical data facilitates to grasp the overview structure.
+From the interactive perspective, selecting a parent node along the tree subsequently triggers its children and their corresponding series, or vice versa.
+It makes navigation simpler than conditional panels, and also enables persistent brushing for comparisons across groups.
+
+```{r tourism-linking}
+include_graphics("img/tourism-linking.png")
+```
+
+```{r plotly-key-tree, eval = FALSE}
+```
 

### 2020-08-24 tsibbletalk.tex
introducing plotly_key_tree()

@@
 remain unchanged over the course of analysis, from trend inspection to
-forecasting performance. Figure \ref{fig:highlight-retail} (a) gives an
-overview of 152 series for the retail data using an overlaid time series
-plot, while Figure \ref{fig:highlight-retail} (b) presents a
-scatterplot, where each series is represented by a dot in the feature
-space (trend versus seasonal strength). The plot making of Figure
-\ref{fig:highlight-retail} (b) is aided with the \texttt{features()}
-function from \CRANpkg{feasts}, which summarises original data by each
-series down to various statistical features. This function along with
-other \textbf{tidyverts} functions is \texttt{tsibble}-aware, and
-outputs a table in a reduced form where each row corresponds to a
-series, thus graphically displayed as Figure \ref{fig:highlight-retail}
-(b).
+forecasting performance.
+
+Figure \ref{fig:highlight-retail} (a) gives an overview of 152 series
+for the retail data using an overlaid time series plot, while Figure
+\ref{fig:highlight-retail} (b) presents a scatterplot, where each series
+is represented by a dot in the feature space (trend versus seasonal
+strength). The plot making of Figure \ref{fig:highlight-retail} (b) is
+aided with the \texttt{features()} function from \CRANpkg{feasts}, which
+summarises original data by each series down to various statistical
+features. This function along with other \textbf{tidyverts} functions is
+\texttt{tsibble}-aware, and outputs a table in a reduced form where each
+row corresponds to a series, thus graphically displayed as Figure
+\ref{fig:highlight-retail} (b).
 
@@
 \texttt{aus\_retail} data is therefore specified as
-\texttt{spec\ =\ State\ *\ Industry}, which is the default for the
-presence of multiple \texttt{key} variables. Using \texttt{/} indicates
-if there is a hierarchy in the data. \texttt{tourism\_monthly}
+\texttt{State\ *\ Industry} or \texttt{Industry\ *\ State}, which is the
+default for the presence of multiple \texttt{key} variables. If there is
+a hierarchy in the data, using \texttt{/} is required to indicate the
+parent-child relation, as strictly one direction \texttt{parent/child}.
 
-Specifying the structure is particularly useful, when there's a
-hierarchical structure in the data. For such case, a new function
-\texttt{plotly\_key\_tree()} can help to visualise the ``key'' structure
-as a dendrogram, using plotly. This type of plot gives an overview of
-structured information in the data. Owing to its interactivity, each
-node in the tree plot is clickable and linked to other plots, and in
-turn the navigation at various levels is made easier.
+The \texttt{tourism\_monthly} dataset packaged in \CRANpkg{tsibbletalk},
+contains monthly domestic overnight trips across Australia, to give an
+illustrator of nesting and crossing. The \texttt{key} is comprised of
+three identifying variables: \texttt{State}, \texttt{Region}, and
+\texttt{Purpose} (of trip), in particular \texttt{State} nesting of
+\texttt{Region}, together crossed with \texttt{Purpose}. This
+specification can be translated as follows:
 
-\begin{itemize}
-\tightlist
-\item
-  Constructing recursive lists. Lists of key values, and json
-\item
-  Self-linking: a single data source
-\end{itemize}
+\begin{Schunk}
+\begin{Sinput}
+library(tsibbletalk)
+tourism_shared <- tourism_monthly %>% 
+  as_shared_tsibble(spec = (State / Region) * Purpose)
+\end{Sinput}
+\end{Schunk}
+
+This dataset contains a three-level hierarchy: the root node is
+implicitly Australia, and geographically disaggregated to states and
+lower-level tourism regions. A new function \texttt{plotly\_key\_tree()}
+has been implemented to address the need of hierarchical discovery
+arising from data. It visually interprets hierarchies in the shared
+tsibble's \texttt{spec} as a tree view, made with \CRANpkg{plotly}. From
+a static point of view, a tree visualisation of hierarchical data
+facilitates to grasp the overview structure. From the interactive
+perspective, selecting a parent node along the tree subsequently
+triggers its children and their corresponding series, or vice versa. It
+makes navigation simpler than conditional panels, and also enables
+persistent brushing for comparisons across groups.
+
+\begin{Schunk}
+
+\begin{center}\includegraphics[width=\textwidth]{img/tourism-linking} \end{center}
+
+\end{Schunk}
 
@@
 The University of Auckland\\
-Department of Statistics\\ The Univerisity of Auckland\\ New Zealand\\
+Department of Statistics\\ The University of Auckland\\ New Zealand\\
 }

### 2020-08-24 scripts/tourism.R

@@ -1,0 +1,32 @@
+## ---- tourism-shared
+library(tsibbletalk)
+tourism_shared <- tourism_monthly %>% 
+  as_shared_tsibble(spec = (State / Region) * Purpose)
+
+## ---- load-other-pkgs
+library(feasts)
+library(plotly)
+library(ggplot2)
+library(tsibble)
+
+## ---- plotly-key-tree
+tourism_feat <- tourism_shared %>%
+  features(Trips, feat_stl)
+
+p0 <- plotly_key_tree(tourism_shared, height = 1100, width = 800)
+p1 <- tourism_shared %>%
+  ggplot(aes(x = Month, y = Trips)) +
+  geom_line(aes(group = Region), alpha = .5, size = .4) +
+  facet_wrap(~ Purpose, scales = "free_y") +
+  scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
+p2 <- tourism_feat %>%
+  ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
+  geom_point(aes(group = Region))
+
+subplot(p0,
+  subplot(
+    ggplotly(p1, tooltip = "Region", width = 1100),
+    ggplotly(p2, tooltip = "Region", width = 1100),
+    nrows = 2),
+  widths = c(.4, .6)) %>%
+  highlight(dynamic = TRUE)

### 2020-08-25 tsibbletalk.Rmd
more on figure 2

@@
 This dataset contains a three-level hierarchy: the root node is implicitly Australia, and geographically disaggregated to states and lower-level tourism regions.
-A new function `plotly_key_tree()` has been implemented to address the need of hierarchical discovery arising from data.
+A new handy function `plotly_key_tree()` has been implemented to address the need of hierarchical discovery arising from the data.
 It visually interprets hierarchies in the shared tsibble's `spec` as a tree view, made with \CRANpkg{plotly}.
-From a static point of view, a tree visualisation of hierarchical data facilitates to grasp the overview structure.
-From the interactive perspective, selecting a parent node along the tree subsequently triggers its children and their corresponding series, or vice versa.
-It makes navigation simpler than conditional panels, and also enables persistent brushing for comparisons across groups.
+The following code line produces the linked tree diagram and fills the left panel of Figure \ref{fig:tourism-linking-fig}.
 
-```{r tourism-linking}
+```{r plotly-key-tree, echo = TRUE}
+```
+
+The visual of tree hierarchy untangles a group of related series and snapshots the data organisation from a bird's eye view.
+It is accompanied with other two plots to compose an ensemble of coordinated graphics for multi-facetted exploration, shown as Figure \ref{fig:tourism-linking-fig}.
+The time series and scatter plots of Figure \ref{fig:tourism-linking-fig} are the interactive realisation of Figure \ref{fig:highlight-retail}.
+Since all plots are derived from one shared tsibble data source, they are self-linking multiple views.
+For example, selecting the leaf node "Melbourne" fires up four respective series of the `Purpose` category in other subplots, characterising "categorical linking".
+Press the "Shift" key to switch on persistent selection and simultaneously select the parent node "Western Australia" to trigger its children.
+The domestic tourism performance in Melbourne can be quickly compared against the whole state of Western Australia.
+It makes navigation simpler than conditional panels, and also enables persistent selection for comparisons across groups.
+
+```{r tourism-linking-fig, fig.cap = "ToDo"}
 include_graphics("img/tourism-linking.png")
@@
 
-```{r plotly-key-tree, eval = FALSE}
+```{r tourism-linking, eval = FALSE}
 ```

### 2020-08-25 tsibbletalk.tex
more on figure 2

@@
 implicitly Australia, and geographically disaggregated to states and
-lower-level tourism regions. A new function \texttt{plotly\_key\_tree()}
-has been implemented to address the need of hierarchical discovery
-arising from data. It visually interprets hierarchies in the shared
-tsibble's \texttt{spec} as a tree view, made with \CRANpkg{plotly}. From
-a static point of view, a tree visualisation of hierarchical data
-facilitates to grasp the overview structure. From the interactive
-perspective, selecting a parent node along the tree subsequently
-triggers its children and their corresponding series, or vice versa. It
-makes navigation simpler than conditional panels, and also enables
-persistent brushing for comparisons across groups.
+lower-level tourism regions. A new handy function
+\texttt{plotly\_key\_tree()} has been implemented to address the need of
+hierarchical discovery arising from the data. It visually interprets
+hierarchies in the shared tsibble's \texttt{spec} as a tree view, made
+with \CRANpkg{plotly}. The following code line produces the linked tree
+diagram and fills the left panel of Figure
+\ref{fig:tourism-linking-fig}.
 
 \begin{Schunk}
+\begin{Sinput}
+p_l <- plotly_key_tree(tourism_shared, height = 1100, width = 800)
+\end{Sinput}
+\end{Schunk}
 
-\begin{center}\includegraphics[width=\textwidth]{img/tourism-linking} \end{center}
+The visual of tree hierarchy untangles a group of related series and
+snapshots the data organisation from a bird's eye view. It is
+accompanied with other two plots to compose an ensemble of coordinated
+graphics for multi-facetted exploration, shown as Figure
+\ref{fig:tourism-linking-fig}. The time series and scatter plots of
+Figure \ref{fig:tourism-linking-fig} are the interactive realisation of
+Figure \ref{fig:highlight-retail}. Since all plots are derived from one
+shared tsibble data source, they are self-linking multiple views. For
+example, selecting the leaf node ``Melbourne'' fires up four respective
+series of the \texttt{Purpose} category in other subplots,
+characterising ``categorical linking''. Press the ``Shift'' key to
+switch on persistent selection and simultaneously select the parent node
+``Western Australia'' to trigger its children. The domestic tourism
+performance in Melbourne can be quickly compared against the whole state
+of Western Australia. It makes navigation simpler than conditional
+panels, and also enables persistent selection for comparisons across
+groups.
 
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{img/tourism-linking}
+
+}
+
+\caption[ToDo]{ToDo}\label{fig:tourism-linking-fig}
+\end{figure}
 \end{Schunk}

### 2020-08-25 scripts/tourism.R

@@ -10,23 +10,24 @@ library(ggplot2)
 library(tsibble)
 
 ## ---- plotly-key-tree
+p_l <- plotly_key_tree(tourism_shared, height = 1100, width = 800)
+
+## ---- tourism-linking
 tourism_feat <- tourism_shared %>%
   features(Trips, feat_stl)
-
-p0 <- plotly_key_tree(tourism_shared, height = 1100, width = 800)
-p1 <- tourism_shared %>%
+p_tr <- tourism_shared %>%
   ggplot(aes(x = Month, y = Trips)) +
   geom_line(aes(group = Region), alpha = .5, size = .4) +
   facet_wrap(~ Purpose, scales = "free_y") +
   scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
-p2 <- tourism_feat %>%
+p_br <- tourism_feat %>%
   ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
   geom_point(aes(group = Region))
 
-subplot(p0,
+subplot(p_l,
   subplot(
-    ggplotly(p1, tooltip = "Region", width = 1100),
-    ggplotly(p2, tooltip = "Region", width = 1100),
+    ggplotly(p_tr, tooltip = "Region", width = 1100),
+    ggplotly(p_br, tooltip = "Region", width = 1100),
     nrows = 2),
   widths = c(.4, .6)) %>%
   highlight(dynamic = TRUE)

### 2020-08-26 tsibbletalk.Rmd
added narratives to code chunks

@@
 A new handy function `plotly_key_tree()` has been implemented to address the need of hierarchical discovery arising from the data.
-It visually interprets hierarchies in the shared tsibble's `spec` as a tree view, made with \CRANpkg{plotly}.
+It interprets hierarchies in the shared tsibble's `spec` as a tree view, built with \CRANpkg{plotly}.
 The following code line produces the linked tree diagram and fills the left panel of Figure \ref{fig:tourism-linking-fig}.
+The visual of tree hierarchy untangles a group of related series and snapshots the data organisation from a bird's eye view.
 
@@
 
-The visual of tree hierarchy untangles a group of related series and snapshots the data organisation from a bird's eye view.
-It is accompanied with other two plots to compose an ensemble of coordinated graphics for multi-facetted exploration, shown as Figure \ref{fig:tourism-linking-fig}.
-The time series and scatter plots of Figure \ref{fig:tourism-linking-fig} are the interactive realisation of Figure \ref{fig:highlight-retail}.
-Since all plots are derived from one shared tsibble data source, they are self-linking multiple views.
-For example, selecting the leaf node "Melbourne" fires up four respective series of the `Purpose` category in other subplots, characterising "categorical linking".
-Press the "Shift" key to switch on persistent selection and simultaneously select the parent node "Western Australia" to trigger its children.
+The tree plot provides backbones of the data, and much flesh yet to be attached.
+Small multiples of time series lines are composed and placed at the top right of Figure \ref{fig:tourism-linking-fig} to unpack the temporal trend across regions by purposes of trips.
+The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code.
+
+```{r tourism-series, echo = TRUE, eval = FALSE}
+```
+
+To tease apart these overlaid time series, they are funnelled through the `features()` S3 method to extract some key characteristics, including the measurements of trend and seasonality.
+A scatterplot is populated from these statistics for each series.
+
+```{r tourism-scatter, echo = TRUE, eval = FALSE}
+```
+
+Lastly, three graphics are composed as an ensemble of coordinated views for multi-facetted exploration, shown as Figure \ref{fig:tourism-linking-fig} (the interactive realisation of Figure \ref{fig:highlight-retail}).
+Routine functions bring about new interaction with temporal data on the client side (i.e.
+without Shiny).
+
+```{r tourism-multi, echo = TRUE, eval = FALSE}
+```
+
+Since all plots are stemmed from one shared tsibble data source, they are self-linking views.
+Nodes, lines, and points are clickable.
+For example, selecting the leaf node "Melbourne" fires up four respective series of the `Purpose` category in other subplots, characterised by "categorical linking".
+Press the "Shift" key to switch on persistent selection, and simultaneously select the parent node "Western Australia" to trigger its children.
 The domestic tourism performance in Melbourne can be quickly compared against the whole state of Western Australia.
@@
 include_graphics("img/tourism-linking.png")
-```
-
-```{r tourism-linking, eval = FALSE}
 ```

### 2020-08-26 tsibbletalk.tex
added narratives to code chunks

@@
 \texttt{plotly\_key\_tree()} has been implemented to address the need of
-hierarchical discovery arising from the data. It visually interprets
-hierarchies in the shared tsibble's \texttt{spec} as a tree view, made
-with \CRANpkg{plotly}. The following code line produces the linked tree
+hierarchical discovery arising from the data. It interprets hierarchies
+in the shared tsibble's \texttt{spec} as a tree view, built with
+\CRANpkg{plotly}. The following code line produces the linked tree
 diagram and fills the left panel of Figure
-\ref{fig:tourism-linking-fig}.
+\ref{fig:tourism-linking-fig}. The visual of tree hierarchy untangles a
+group of related series and snapshots the data organisation from a
+bird's eye view.
 
@@
 
-The visual of tree hierarchy untangles a group of related series and
-snapshots the data organisation from a bird's eye view. It is
-accompanied with other two plots to compose an ensemble of coordinated
-graphics for multi-facetted exploration, shown as Figure
-\ref{fig:tourism-linking-fig}. The time series and scatter plots of
-Figure \ref{fig:tourism-linking-fig} are the interactive realisation of
-Figure \ref{fig:highlight-retail}. Since all plots are derived from one
-shared tsibble data source, they are self-linking multiple views. For
+The tree plot provides backbones of the data, and much flesh yet to be
+attached. Small multiples of time series lines are composed and placed
+at the top right of Figure \ref{fig:tourism-linking-fig} to unpack the
+temporal trend across regions by purposes of trips. The shared tsibble
+data can be directly piped into \CRANpkg{ggplot2} code.
+
+\begin{Schunk}
+\begin{Sinput}
+p_tr <- tourism_shared %>%
+  ggplot(aes(x = Month, y = Trips)) +
+  geom_line(aes(group = Region), alpha = .5, size = .4) +
+  facet_wrap(~ Purpose, scales = "free_y") +
+  scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
+\end{Sinput}
+\end{Schunk}
+
+To tease apart these overlaid time series, they are funnelled through
+the \texttt{features()} S3 method to extract some key characteristics,
+including the measurements of trend and seasonality. A scatterplot is
+populated from these statistics for each series.
+
+\begin{Schunk}
+\begin{Sinput}
+tourism_feat <- tourism_shared %>%
+  features(Trips, feat_stl)
+p_br <- tourism_feat %>%
+  ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
+  geom_point(aes(group = Region))
+\end{Sinput}
+\end{Schunk}
+
+Lastly, three graphics are composed as an ensemble of coordinated views
+for multi-facetted exploration, shown as Figure
+\ref{fig:tourism-linking-fig} (the interactive realisation of Figure
+\ref{fig:highlight-retail}). Routine functions bring about new
+interaction with temporal data on the client side (i.e.~without Shiny).
+
+\begin{Schunk}
+\begin{Sinput}
+subplot(p_l,
+  subplot(
+    ggplotly(p_tr, tooltip = "Region", width = 1100),
+    ggplotly(p_br, tooltip = "Region", width = 1100),
+    nrows = 2),
+  widths = c(.4, .6)) %>%
+  highlight(dynamic = TRUE)
+\end{Sinput}
+\end{Schunk}
+
+Since all plots are stemmed from one shared tsibble data source, they
+are self-linking views. Nodes, lines, and points are clickable. For
 example, selecting the leaf node ``Melbourne'' fires up four respective
-series of the \texttt{Purpose} category in other subplots,
-characterising ``categorical linking''. Press the ``Shift'' key to
-switch on persistent selection and simultaneously select the parent node
+series of the \texttt{Purpose} category in other subplots, characterised
+by ``categorical linking''. Press the ``Shift'' key to switch on
+persistent selection, and simultaneously select the parent node
 ``Western Australia'' to trigger its children. The domestic tourism

### 2020-08-26 scripts/tourism.R

@@ -12,18 +12,21 @@ library(tsibble)
 ## ---- plotly-key-tree
 p_l <- plotly_key_tree(tourism_shared, height = 1100, width = 800)
 
-## ---- tourism-linking
-tourism_feat <- tourism_shared %>%
-  features(Trips, feat_stl)
+## ---- tourism-series
 p_tr <- tourism_shared %>%
   ggplot(aes(x = Month, y = Trips)) +
   geom_line(aes(group = Region), alpha = .5, size = .4) +
   facet_wrap(~ Purpose, scales = "free_y") +
   scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
+
+## ---- tourism-scatter
+tourism_feat <- tourism_shared %>%
+  features(Trips, feat_stl)
 p_br <- tourism_feat %>%
   ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
   geom_point(aes(group = Region))
 
+## ---- tourism-multi
 subplot(p_l,
   subplot(
     ggplotly(p_tr, tooltip = "Region", width = 1100),

### 2020-08-28 tsibbletalk.Rmd
done with shared tsibble

@@
   email: dicook@monash.edu
-bibliography: references.bib
+bibliography: tsibbletalk.bib
 abstract: >
@@
 
+```{r tourism-linking-fig, fig.cap = "ToDo"}
+include_graphics("img/tourism-linking.png")
+```
+
 This dataset contains a three-level hierarchy: the root node is implicitly Australia, and geographically disaggregated to states and lower-level tourism regions.
@@
 Lastly, three graphics are composed as an ensemble of coordinated views for multi-facetted exploration, shown as Figure \ref{fig:tourism-linking-fig} (the interactive realisation of Figure \ref{fig:highlight-retail}).
-Routine functions bring about new interaction with temporal data on the client side (i.e.
-without Shiny).
+Routine functions bring about new interaction with temporal data on the client side.
 
@@
 Since all plots are stemmed from one shared tsibble data source, they are self-linking views.
-Nodes, lines, and points are clickable.
-For example, selecting the leaf node "Melbourne" fires up four respective series of the `Purpose` category in other subplots, characterised by "categorical linking".
-Press the "Shift" key to switch on persistent selection, and simultaneously select the parent node "Western Australia" to trigger its children.
-The domestic tourism performance in Melbourne can be quickly compared against the whole state of Western Australia.
-It makes navigation simpler than conditional panels, and also enables persistent selection for comparisons across groups.
+Nodes, lines, and points are hoverable and clickable.
+Given the `spec`, clicking either one element in any plot highlights all points that match the `Region` category, briefly "categorical linking".
+In Figure \ref{fig:tourism-linking-fig}, when hovering and selecting the circle associated with "Sydney" in the scatter plot, all data records with shared values of "Sydney" listen and react to this interaction via self updating in red.
+In order for comparison with other regions or states, press the "Shift" key to enable persistent selection, and simultaneously select the parent node on the tree, saying "Western Australia", to include all the children by switching to the blue colour.
+The domestic tourism sees Sydney as one of the most popular destinations in realm of business and friends visiting over years.
+Despite of relatively weaker performance in Western Australia, Australia's North West region sees the strongest upward trend, bypassing Sydney in some years.
 
-```{r tourism-linking-fig, fig.cap = "ToDo"}
-include_graphics("img/tourism-linking.png")
-```
+In summary, shared tsibble data nicely bridges between the \CRANpkg{crosstalk} and **tidyverts** ecosystems.
+The `as_shared_tsibble()` provides a symbolic user interface for effortless construction of a hybrid of hierarchical and categorical linkings.
+And the `plotly_key_tree()` in turn decodes the specification to plot a tree for data overview and navigation, accompanied with more detailed plots.
 

### 2020-08-28 tsibbletalk.tex
done with shared tsibble

@@
 structured temporal data}
-\author{by Earo Wang, Dianne Cook}
+\author{by Earo Wang and Dianne Cook}
 
@@
 
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{img/tourism-linking}
+
+}
+
+\caption[ToDo]{ToDo}\label{fig:tourism-linking-fig}
+\end{figure}
+\end{Schunk}
+
 This dataset contains a three-level hierarchy: the root node is
@@
   ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
-  geom_point(aes(group = Region))
+  geom_point(aes(group = Region), alpha = .8, size = 2)
 \end{Sinput}
@@
 \ref{fig:highlight-retail}). Routine functions bring about new
-interaction with temporal data on the client side (i.e.~without Shiny).
+interaction with temporal data on the client side.
 
@@
 Since all plots are stemmed from one shared tsibble data source, they
-are self-linking views. Nodes, lines, and points are clickable. For
-example, selecting the leaf node ``Melbourne'' fires up four respective
-series of the \texttt{Purpose} category in other subplots, characterised
-by ``categorical linking''. Press the ``Shift'' key to switch on
-persistent selection, and simultaneously select the parent node
-``Western Australia'' to trigger its children. The domestic tourism
-performance in Melbourne can be quickly compared against the whole state
-of Western Australia. It makes navigation simpler than conditional
-panels, and also enables persistent selection for comparisons across
-groups.
+are self-linking views. Nodes, lines, and points are hoverable and
+clickable. Given the \texttt{spec}, clicking either one element in any
+plot highlights all points that match the \texttt{Region} category,
+briefly ``categorical linking''. In Figure
+\ref{fig:tourism-linking-fig}, when hovering and selecting the circle
+associated with ``Sydney'' in the scatter plot, all data records with
+shared values of ``Sydney'' listen and react to this interaction via
+self updating in red. In order for comparison with other regions or
+states, press the ``Shift'' key to enable persistent selection, and
+simultaneously select the parent node on the tree, saying ``Western
+Australia'', to include all the children by switching to the blue
+colour. The domestic tourism sees Sydney as one of the most popular
+destinations in realm of business and friends visiting over years.
+Despite of relatively weaker performance in Western Australia,
+Australia's North West region sees the strongest upward trend, bypassing
+Sydney in some years.
 
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{img/tourism-linking}
-
-}
-
-\caption[ToDo]{ToDo}\label{fig:tourism-linking-fig}
-\end{figure}
-\end{Schunk}
+In summary, shared tsibble data nicely bridges between the
+\CRANpkg{crosstalk} and \textbf{tidyverts} ecosystems. The
+\texttt{as\_shared\_tsibble()} provides a symbolic user interface for
+effortless construction of a hybrid of hierarchical and categorical
+linkings. And the \texttt{plotly\_key\_tree()} in turn decodes the
+specification to plot a tree for data overview and navigation,
+accompanied with more detailed plots.
 
@@
 
-\bibliography{references.bib}
+\bibliography{tsibbletalk.bib}
 
@@
 Earo Wang\\
-The University of Auckland\\
+The University of Auckland\\%
 Department of Statistics\\ The University of Auckland\\ New Zealand\\
+%
+%
+%
+\\\href{mailto:earo.wang@auckland.ac.nz}{\nolinkurl{earo.wang@auckland.ac.nz}}
 }
-\href{mailto:earo.wang@auckland.ac.nz}{\nolinkurl{earo.wang@auckland.ac.nz}}
 
@@
 Dianne Cook\\
-Monash University\\
+Monash University\\%
 Department of Econometrics and Business Statistics\\ Monash
 University\\ Australia\\
+%
+%
+%
+\\\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
 }
-\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}

### 2020-08-28 scripts/tourism.R

@@ -24,7 +24,7 @@ tourism_feat <- tourism_shared %>%
   features(Trips, feat_stl)
 p_br <- tourism_feat %>%
   ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
-  geom_point(aes(group = Region))
+  geom_point(aes(group = Region), alpha = .8, size = 2)
 
 ## ---- tourism-multi
 subplot(p_l,

### 2020-08-28 tsibbletalk.R

@@ -1,0 +1,102 @@
+## ----setup, echo = FALSE, cache = FALSE, include = FALSE----------------------
+options("knitr.graphics.auto_pdf" = TRUE)
+library(knitr)
+opts_chunk$set(
+  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
+  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
+  cache = TRUE, cache.path = 'cache/',
+  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
+)
+opts_knit$set(root.dir = here::here())
+read_chunk("scripts/demo.R")
+read_chunk("scripts/tourism.R")
+
+
+## ---- load-pkgs---------------------------------------------------------------
+library(feasts)
+library(tsibble)
+library(tsibbledata)
+
+
+## ---- print-retail------------------------------------------------------------
+print(aus_retail, n = 5)
+
+
+## ----highlight-retail, fig.height = 3.3, fig.cap = "ToDo"---------------------
+library(tidyverse)
+library(ggrepel)
+library(gghighlight)
+library(patchwork)
+
+p_retail_ts <- aus_retail %>% 
+  as_tibble() %>% # gghighlight issue for group_by() + filter()
+  mutate(group = paste(State, ":", Industry)) %>% 
+  ggplot(aes(x = Month, y = Turnover)) +
+  geom_line(aes(group = group)) +
+  gghighlight(
+    State == "Queensland", Industry == "Department stores",
+    label_params = list(vjust = -2)
+  ) +
+  scale_x_yearmonth(breaks = yearmonth(c("1990 Jan", "2000 Jan", "2010 Jan")))
+
+p_retail_feat <- aus_retail %>% 
+  features(Turnover, feat_stl) %>% 
+  mutate(group = paste(State, ":", Industry)) %>% 
+  ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
+  geom_point(aes(group = group)) +
+  labs(x = "Trend strength", y = "Seasonal strength") +
+  gghighlight(
+    State == "Queensland", Industry == "Department stores",
+    use_direct_label = FALSE
+  ) +
+  geom_label_repel(aes(label = group), vjust = 2, nudge_x = -0.1)
+
+p_retail_ts + p_retail_feat + plot_annotation(tag_levels = "a")
+
+
+## ----tourism-shared, echo = TRUE----------------------------------------------
[…]

### 2020-08-31 tsibbletalk.Rmd
added code and imgs for slicing section

@@
 preamble: |
-  % Any extra LaTeX you need in the preamble
+  \usepackage{subfig}
 output: rticles::rjournal_article
@@
 
-Figure \ref{fig:highlight-retail} (a) gives an overview of 152 series for the retail data using an overlaid time series plot, while Figure \ref{fig:highlight-retail} (b) presents a scatterplot, where each series is represented by a dot in the feature space (trend versus seasonal strength).
-The plot making of Figure \ref{fig:highlight-retail} (b) is aided with the `features()` function from \CRANpkg{feasts}, which summarises original data by each series down to various statistical features.
-This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, thus graphically displayed as Figure \ref{fig:highlight-retail} (b).
+Figure \ref{fig:highlight-retail-1} gives an overview of 152 series for the retail data using an overlaid time series plot, while Figure \ref{fig:highlight-retail-2} presents a scatterplot, where each series is represented by a dot in the feature space (trend versus seasonal strength).
+The plot making of Figure \ref{fig:highlight-retail-2} is aided with the `features()` function from \CRANpkg{feasts}, which summarises original data by each series down to various statistical features.
+This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, thus graphically displayed as Figure \ref{fig:highlight-retail-2}.
 
-```{r highlight-retail, fig.height = 3.3, fig.cap = "ToDo"}
+```{r highlight-retail, fig.height = 6, out.width = ".49\\linewidth", fig.show = "hold", fig.cap = "ToDo", fig.subcap = c("1", "2"), fig.ncol = 2}
 ```
@@
 
+```{r wrap-ped, fig.show = "hold", fig.cap = "ToDo", fig.subcap = c("(1)", "(2)", "(3)"), fig.ncol = 1}
+include_graphics("img/wrap-0.png")
+include_graphics("img/wrap-1.png")
+include_graphics("img/wrap-7.png")
+```
+
 The other critical aspect of a tsibble is "index", that provides foundational temporal context.

### 2020-08-31 tsibbletalk.tex
added code and imgs for slicing section

@@
 % !TeX root = RJwrapper.tex
-\title{Conversations in time: interactive visualisation to explore
-structured temporal data}
+\title{Conversations in time: interactive visualisation to explore structured
+temporal data}
 \author{by Earo Wang and Dianne Cook}
@@
 
-Figure \ref{fig:highlight-retail} (a) gives an overview of 152 series
-for the retail data using an overlaid time series plot, while Figure
-\ref{fig:highlight-retail} (b) presents a scatterplot, where each series
+Figure \ref{fig:highlight-retail-1} gives an overview of 152 series for
+the retail data using an overlaid time series plot, while Figure
+\ref{fig:highlight-retail-2} presents a scatterplot, where each series
 is represented by a dot in the feature space (trend versus seasonal
-strength). The plot making of Figure \ref{fig:highlight-retail} (b) is
+strength). The plot making of Figure \ref{fig:highlight-retail-2} is
 aided with the \texttt{features()} function from \CRANpkg{feasts}, which
@@
 row corresponds to a series, thus graphically displayed as Figure
-\ref{fig:highlight-retail} (b).
+\ref{fig:highlight-retail-2}.
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figure/highlight-retail-1}
+{\centering \subfloat[1\label{fig:highlight-retail-1}]{\includegraphics[width=.49\linewidth]{figure/highlight-retail-1} }\subfloat[2\label{fig:highlight-retail-2}]{\includegraphics[width=.49\linewidth]{figure/highlight-retail-2} }
 
@@
 
+\begin{Schunk}
+\begin{figure}
+
+{\centering \subfloat[(1)\label{fig:wrap-ped-1}]{\includegraphics[width=\textwidth]{img/wrap-0} }\newline\subfloat[(2)\label{fig:wrap-ped-2}]{\includegraphics[width=\textwidth]{img/wrap-1} }\newline\subfloat[(3)\label{fig:wrap-ped-3}]{\includegraphics[width=\textwidth]{img/wrap-7} }
+
+}
+
+\caption[ToDo]{ToDo}\label{fig:wrap-ped}
+\end{figure}
+\end{Schunk}
+
 The other critical aspect of a tsibble is ``index'', that provides
@@
 Monash University\\%
-Department of Econometrics and Business Statistics\\ Monash
-University\\ Australia\\
+Department of Econometrics and Business Statistics\\ Monash University\\ Australia\\
 %

### 2020-08-31 scripts/demo.R

@@ -10,9 +10,8 @@ print(aus_retail, n = 5)
 library(tidyverse)
 library(ggrepel)
 library(gghighlight)
-library(patchwork)
 
-p_retail_ts <- aus_retail %>% 
+aus_retail %>% 
   as_tibble() %>% # gghighlight issue for group_by() + filter()
   mutate(group = paste(State, ":", Industry)) %>% 
   ggplot(aes(x = Month, y = Turnover)) +
@@ -23,7 +22,7 @@ p_retail_ts <- aus_retail %>%
   ) +
   scale_x_yearmonth(breaks = yearmonth(c("1990 Jan", "2000 Jan", "2010 Jan")))
 
-p_retail_feat <- aus_retail %>% 
+aus_retail %>% 
   features(Turnover, feat_stl) %>% 
   mutate(group = paste(State, ":", Industry)) %>% 
   ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
@@ -34,5 +33,3 @@ p_retail_feat <- aus_retail %>%
     use_direct_label = FALSE
   ) +
   geom_label_repel(aes(label = group), vjust = 2, nudge_x = -0.1)
-
-p_retail_ts + p_retail_feat + plot_annotation(tag_levels = "a")

### 2020-08-31 scripts/pedestrian.R

@@ -1,0 +1,33 @@
+library(rwalkr)
+library(tsibble)
+library(tsibbletalk)
+library(shiny)
+library(plotly)
+library(tidyverse)
+library(lubridate)
+
+sensors <- c("Birrarung Marr", "Bourke Street Mall (North)",
+  "QV Market-Elizabeth St (West)", "Southern Cross Station")
+
+ped20 <- melb_walk_fast(year = 2020, sensor = sensors) %>% 
+  filter(Date < ymd("2020-06-01")) %>% 
+  mutate(Lockdown = ifelse(Date > ymd("2020-03-16"), "Yes", "No")) %>% 
+  as_tsibble(index = Date_Time, key = Sensor)
+
+p_line <- ped20 %>%
+  ggplot(aes(x = Date_Time, y = Count, colour = Lockdown)) +
+  geom_line(size = .3) +
+  facet_wrap(~ Sensor, scales = "free_y") +
+  labs(x = "Date Time") +
+  scale_colour_brewer(palette = "Dark2") +
+  theme(legend.position = "none")
+
+ui <- fluidPage(
+  tsibbleDiceUI("dice")
+)
+
+server <- function(input, output, session) {
+  tsibbleDiceServer("dice", ggplotly(p_line, height = 700), period = "1 day")
+}
+
+shinyApp(ui, server)

### 2020-09-03 tsibbletalk.Rmd
openning paragraph for slicing time

@@
 
+The shared tsibble data leverages the `key` attribute to converse with many coordinated views, with or without \CRANpkg{shiny}.
+On the other hand, a second critical attribute--`index`--lays the foundational temporal context that augments the conversation.
+When temporal data are plotted and stretched against the entire span like Figure \ref{fig:highlight-retail-1}, it puts the emphasis on the trend.
+When the data are wrapped over origin-less time units, such as one or two quarters, it makes seasonal patterns more pronounced.
+The \CRANpkg{tsibbletalk} package provides a shiny module
+
 ```{r wrap-ped, fig.show = "hold", fig.cap = "ToDo", fig.subcap = c("(1)", "(2)", "(3)"), fig.ncol = 1}
@@
 
-The other critical aspect of a tsibble is "index", that provides foundational temporal context.
-A common tool in time series analytical toolkit is seasonal plots that lay time series not on the whole time scale, but on an origin-less relative time unit, for example `gg_season()` in the {feasts} package.
-It helps to examine and emphasise periodic/aperiodic patterns, comparing to time series plots that primarily focus on trends.
 Standard seasonal plots break the overall time into two components: seasonal periods on the x-axis, and grouped by their corresponding lower-resolution time.

### 2020-09-03 tsibbletalk.tex
openning paragraph for slicing time

@@
 
+The shared tsibble data leverages the \texttt{key} attribute to converse
+with many coordinated views, with or without \CRANpkg{shiny}. On the
+other hand, a second critical attribute--\texttt{index}--lays the
+foundational temporal context that augments the conversation. When
+temporal data are plotted and stretched against the entire span like
+Figure \ref{fig:highlight-retail-1}, it puts the emphasis on the trend.
+When the data are wrapped over origin-less time units, such as one or
+two quarters, it makes seasonal patterns more pronounced. The
+\CRANpkg{tsibbletalk} package provides a shiny module
+
 \begin{Schunk}
@@
 
-The other critical aspect of a tsibble is ``index'', that provides
-foundational temporal context. A common tool in time series analytical
-toolkit is seasonal plots that lay time series not on the whole time
-scale, but on an origin-less relative time unit, for example
-\texttt{gg\_season()} in the \{feasts\} package. It helps to examine and
-emphasise periodic/aperiodic patterns, comparing to time series plots
-that primarily focus on trends. Standard seasonal plots break the
-overall time into two components: seasonal periods on the x-axis, and
-grouped by their corresponding lower-resolution time. For example,
-monthly data can be decomposed into months separated by years, and
-hourly data into hours grouped by days. Data collected at lower-level
-resolutions often exhibits more than one seasonal patterns. To discover
-typical seasonal or non-typical profiles, it is helpful to quickly
-browse through many possible periods. Interactivity ought to be enabled.
+Standard seasonal plots break the overall time into two components:
+seasonal periods on the x-axis, and grouped by their corresponding
+lower-resolution time. For example, monthly data can be decomposed into
+months separated by years, and hourly data into hours grouped by days.
+Data collected at lower-level resolutions often exhibits more than one
+seasonal patterns. To discover typical seasonal or non-typical profiles,
+it is helpful to quickly browse through many possible periods.
+Interactivity ought to be enabled.
 

### 2020-09-05 tsibbletalk.Rmd
figure 3 descriptions

@@
 read_chunk("scripts/tourism.R")
+read_chunk("scripts/pedestrian.R")
 ```
@@
 On the other hand, a second critical attribute--`index`--lays the foundational temporal context that augments the conversation.
-When temporal data are plotted and stretched against the entire span like Figure \ref{fig:highlight-retail-1}, it puts the emphasis on the trend.
-When the data are wrapped over origin-less time units, such as one or two quarters, it makes seasonal patterns more pronounced.
-The \CRANpkg{tsibbletalk} package provides a shiny module
+When temporal data are plotted and stretched against the entire span like Figure \ref{fig:highlight-retail-1}, it puts emphasis on the trend perception.
+Yet to digest periodic/aperiodic patterns, data should be wrapped over relative time units that are origin-less, such as one quarter or one day.
 
@@
 
-Standard seasonal plots break the overall time into two components: seasonal periods on the x-axis, and grouped by their corresponding lower-resolution time.
-For example, monthly data can be decomposed into months separated by years, and hourly data into hours grouped by days.
-Data collected at lower-level resolutions often exhibits more than one seasonal patterns.
-To discover typical seasonal or non-typical profiles, it is helpful to quickly browse through many possible periods.
-Interactivity ought to be enabled.
+The city of Melbourne has sensors installed to count hourly tallies of pedestrians in order to capture downtown daily rhythms (cite).
+Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four locations, with the depiction of three pronounced slices in time.
+Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on their absolute time scales, facetted by four sensors.
+On March 16, Melbourne went to the stage three lockdown due to COVID-19, seeing a significant decline in traffic volume across the city.
+These lines are then folded into daily and weekly sections, shown as Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} respectively.
+Seasonal variations have been popped out to viewers, complementing the not-just-magnitude-drop data story.
+The pre-lockdown period is coloured with dark green and lockdown with orange.
 
-The {tsibbletalk} package provides a pair of UI and server functions, as a shiny module, to help with finding interesting time slices in a shiny application.
+Figure \ref{fig:wrap-ped} suggests that there could be more than one interesting time slices.
+The \CRANpkg{tsibbletalk} package provides a pair of UI and server functions, as a shiny module, to help with finding interesting seasonal slices in a shiny application.
 The pair, decoupled to `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece.
@@
 The other mandatory argument is to specify the number of seasonal periods that requires shifting.
+
+```{r load-ped, eval = FALSE}
+```
+
+```{r ped-slice, echo = TRUE, eval = FALSE}
+```
+
+Standard seasonal plots break the overall time into two components: seasonal periods on the x-axis, and grouped by their corresponding lower-resolution time.
+For example, monthly data can be decomposed into months separated by years, and hourly data into hours grouped by days.
+Data collected at lower-level resolutions often exhibits more than one seasonal patterns.
+To discover typical seasonal or non-typical profiles, it is helpful to quickly browse through many possible periods.
+Interactivity ought to be enabled.
 

### 2020-09-05 tsibbletalk.tex
figure 3 descriptions

@@
 % !TeX root = RJwrapper.tex
-\title{Conversations in time: interactive visualisation to explore structured
-temporal data}
+\title{Conversations in time: interactive visualisation to explore
+structured temporal data}
 \author{by Earo Wang and Dianne Cook}
@@
 temporal data are plotted and stretched against the entire span like
-Figure \ref{fig:highlight-retail-1}, it puts the emphasis on the trend.
-When the data are wrapped over origin-less time units, such as one or
-two quarters, it makes seasonal patterns more pronounced. The
-\CRANpkg{tsibbletalk} package provides a shiny module
+Figure \ref{fig:highlight-retail-1}, it puts emphasis on the trend
+perception. Yet to digest periodic/aperiodic patterns, data should be
+wrapped over relative time units that are origin-less, such as one
+quarter or one day.
 
@@
 
+The city of Melbourne has sensors installed to count hourly tallies of
+pedestrians in order to capture downtown daily rhythms (cite). Figure
+\ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at
+four locations, with the depiction of three pronounced slices in time.
+Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on
+their absolute time scales, facetted by four sensors. On March 16,
+Melbourne went to the stage three lockdown due to COVID-19, seeing a
+significant decline in traffic volume across the city. These lines are
+then folded into daily and weekly sections, shown as Figure
+\ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} respectively. Seasonal
+variations have been popped out to viewers, complementing the
+not-just-magnitude-drop data story. The pre-lockdown period is coloured
+with dark green and lockdown with orange.
+
+Figure \ref{fig:wrap-ped} suggests that there could be more than one
+interesting time slices. The \CRANpkg{tsibbletalk} package provides a
+pair of UI and server functions, as a shiny module, to help with finding
+interesting seasonal slices in a shiny application. The pair, decoupled
+to \texttt{tsibbleDiceUI()} and \texttt{tsibbleDiceServer()}, presents a
+clean interface and forms a resusable piece. Like all shiny modules,
+users should supply a unique session id. The UI function
+\texttt{tsibbleDiceUI()} shows a slider that controls the number of
+periods, and a plot specified by users. The server function
+\texttt{tsibbleDiceServer()} is the workhorse, transforming data and
+updating the plot. It expects a \texttt{ggplot} (converted to
+\texttt{plotly} via \texttt{ggplotly()}) or \texttt{plotly} object. This
+plot can be line charts, or other graphical elements (such as boxplots).
+But it assumes that tsibble's time index is plotted on the x-axis. The
+other mandatory argument is to specify the number of seasonal periods
+that requires shifting.
+
+\begin{Schunk}
+\begin{Sinput}
+p_line <- pedestrian20 %>%
+  ggplot(aes(x = Date_Time, y = Count, colour = Lockdown)) +
+  geom_line(size = .3) +
+  facet_wrap(~ Sensor, scales = "free_y") +
+  labs(x = "Date Time") +
+  scale_colour_brewer(palette = "Dark2") +
+  theme(legend.position = "none")
+
+ui <- fluidPage(tsibbleDiceUI("dice"))
+server <- function(input, output, session) {
+  tsibbleDiceServer("dice", ggplotly(p_line, height = 700), period = "1 day")
+}
+shinyApp(ui, server)
+\end{Sinput}
+\end{Schunk}
+
 Standard seasonal plots break the overall time into two components:
@@
 Interactivity ought to be enabled.
-
-The \{tsibbletalk\} package provides a pair of UI and server functions,
-as a shiny module, to help with finding interesting time slices in a
-shiny application. The pair, decoupled to \texttt{tsibbleDiceUI()} and
-\texttt{tsibbleDiceServer()}, presents a clean interface and forms a
-resusable piece. Like all shiny modules, users should supply a unique
-session id. The UI function \texttt{tsibbleDiceUI()} shows a slider that
-controls the number of periods, and a plot specified by users. The
-server function \texttt{tsibbleDiceServer()} is the workhorse,
-transforming data and updating the plot. It expects a \texttt{ggplot}
-(converted to \texttt{plotly} via \texttt{ggplotly()}) or
-\texttt{plotly} object. This plot can be line charts, or other graphical
-elements (such as boxplots). But it assumes that tsibble's time index is
-plotted on the x-axis. The other mandatory argument is to specify the
-number of seasonal periods that requires shifting.
 
@@
 Monash University\\%
-Department of Econometrics and Business Statistics\\ Monash University\\ Australia\\
+Department of Econometrics and Business Statistics\\ Monash
+University\\ Australia\\
 %

### 2020-09-05 scripts/pedestrian.R

@@ -1,3 +1,4 @@
+## ---- load-ped
 library(rwalkr)
 library(tsibble)
 library(tsibbletalk)
@@ -9,12 +10,13 @@ library(lubridate)
 sensors <- c("Birrarung Marr", "Bourke Street Mall (North)",
   "QV Market-Elizabeth St (West)", "Southern Cross Station")
 
-ped20 <- melb_walk_fast(year = 2020, sensor = sensors) %>% 
+pedestrian20 <- melb_walk_fast(year = 2020, sensor = sensors) %>% 
   filter(Date < ymd("2020-06-01")) %>% 
   mutate(Lockdown = ifelse(Date > ymd("2020-03-16"), "Yes", "No")) %>% 
   as_tsibble(index = Date_Time, key = Sensor)
 
-p_line <- ped20 %>%
+## ---- ped-slice
+p_line <- pedestrian20 %>%
   ggplot(aes(x = Date_Time, y = Count, colour = Lockdown)) +
   geom_line(size = .3) +
   facet_wrap(~ Sensor, scales = "free_y") +
@@ -22,12 +24,8 @@ p_line <- ped20 %>%
   scale_colour_brewer(palette = "Dark2") +
   theme(legend.position = "none")
 
-ui <- fluidPage(
-  tsibbleDiceUI("dice")
-)
-
+ui <- fluidPage(tsibbleDiceUI("dice"))
 server <- function(input, output, session) {
   tsibbleDiceServer("dice", ggplotly(p_line, height = 700), period = "1 day")
 }
-
 shinyApp(ui, server)

### 2020-09-06 tsibbletalk.Rmd
explained wrapping and shiny module functions

@@
 These lines are then folded into daily and weekly sections, shown as Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} respectively.
-Seasonal variations have been popped out to viewers, complementing the not-just-magnitude-drop data story.
+Seasonal variations have been popped out to viewers, complementing the not-just-magnitude-drop story.
 The pre-lockdown period is coloured with dark green and lockdown with orange.
 
-Figure \ref{fig:wrap-ped} suggests that there could be more than one interesting time slices.
-The \CRANpkg{tsibbletalk} package provides a pair of UI and server functions, as a shiny module, to help with finding interesting seasonal slices in a shiny application.
-The pair, decoupled to `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece.
-Like all shiny modules, users should supply a unique session id.
-The UI function `tsibbleDiceUI()` shows a slider that controls the number of periods, and a plot specified by users.
-The server function `tsibbleDiceServer()` is the workhorse, transforming data and updating the plot.
-It expects a `ggplot` (converted to `plotly` via `ggplotly()`) or `plotly` object.
-This plot can be line charts, or other graphical elements (such as boxplots).
-But it assumes that tsibble's time index is plotted on the x-axis.
-The other mandatory argument is to specify the number of seasonal periods that requires shifting.
+The wrapping procedure involves slicing time indices into seasonal periods of interest and their corresponding time dices.
+For example, hourly pedestrian data can be decomposed into 24-hour blocks grouped by all respective days, like Figure \ref{fig:wrap-ped-2}.
+Figure \ref{fig:wrap-ped} suggests that there could be more than one eye-catching slices out of many possible combinations, and thus repeated wrappings can be unwieldy.
+To visually locate an interesting slice, the \CRANpkg{tsibbletalk} package implements a shiny module, a pair of UI and server functions, to automate this wrapping procedure.
+
+This shiny module, decoupled to `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece in a shiny application.
+Like all shiny modules, the first argument in both functions requires a user-supplied session id that must be unique.
+The UI function `tsibbleDiceUI()` simply shows a slider that animates or controls the number of periods to be diced.
+The workhorse is certainly the server function `tsibbleDiceServer()`, encapsulating the core algorithm that transforms data and sends messages to update the plot accordingly.
+It expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots).
+As the function name suggests, a (shared) tsibble is needed to start the engine.
+The `period` argument is to specify the minimum number of seasonal blocks, for example data shifted by "1 day", "2 days", or "1 week", etc.
+The following code chunk generates Figure \ref{fig:wrap-ped}.
 
@@
 
-Standard seasonal plots break the overall time into two components: seasonal periods on the x-axis, and grouped by their corresponding lower-resolution time.
-For example, monthly data can be decomposed into months separated by years, and hourly data into hours grouped by days.
-Data collected at lower-level resolutions often exhibits more than one seasonal patterns.
-To discover typical seasonal or non-typical profiles, it is helpful to quickly browse through many possible periods.
-Interactivity ought to be enabled.
+The initial state is
 

### 2020-09-06 tsibbletalk.tex
explained wrapping and shiny module functions

@@
 variations have been popped out to viewers, complementing the
-not-just-magnitude-drop data story. The pre-lockdown period is coloured
-with dark green and lockdown with orange.
+not-just-magnitude-drop story. The pre-lockdown period is coloured with
+dark green and lockdown with orange.
 
-Figure \ref{fig:wrap-ped} suggests that there could be more than one
-interesting time slices. The \CRANpkg{tsibbletalk} package provides a
-pair of UI and server functions, as a shiny module, to help with finding
-interesting seasonal slices in a shiny application. The pair, decoupled
-to \texttt{tsibbleDiceUI()} and \texttt{tsibbleDiceServer()}, presents a
-clean interface and forms a resusable piece. Like all shiny modules,
-users should supply a unique session id. The UI function
-\texttt{tsibbleDiceUI()} shows a slider that controls the number of
-periods, and a plot specified by users. The server function
-\texttt{tsibbleDiceServer()} is the workhorse, transforming data and
-updating the plot. It expects a \texttt{ggplot} (converted to
-\texttt{plotly} via \texttt{ggplotly()}) or \texttt{plotly} object. This
-plot can be line charts, or other graphical elements (such as boxplots).
-But it assumes that tsibble's time index is plotted on the x-axis. The
-other mandatory argument is to specify the number of seasonal periods
-that requires shifting.
+The wrapping procedure involves slicing time indices into seasonal
+periods of interest and their corresponding time dices. For example,
+hourly pedestrian data can be decomposed into 24-hour blocks grouped by
+all respective days, like Figure \ref{fig:wrap-ped-2}. Figure
+\ref{fig:wrap-ped} suggests that there could be more than one
+eye-catching slices out of many possible combinations, and thus repeated
+wrappings can be unwieldy. To visually locate an interesting slice, the
+\CRANpkg{tsibbletalk} package implements a shiny module, a pair of UI
+and server functions, to automate this wrapping procedure.
+
+This shiny module, decoupled to \texttt{tsibbleDiceUI()} and
+\texttt{tsibbleDiceServer()}, presents a clean interface and forms a
+resusable piece in a shiny application. Like all shiny modules, the
+first argument in both functions requires a user-supplied session id
+that must be unique. The UI function \texttt{tsibbleDiceUI()} simply
+shows a slider that animates or controls the number of periods to be
+diced. The workhorse is certainly the server function
+\texttt{tsibbleDiceServer()}, encapsulating the core algorithm that
+transforms data and sends messages to update the plot accordingly. It
+expects a \texttt{ggplot} or \texttt{plotly} object, where one can plot
+data using either lines or other graphical elements (such as boxplots).
+As the function name suggests, a (shared) tsibble is needed to start the
+engine. The \texttt{period} argument is to specify the minimum number of
+seasonal blocks, for example data shifted by ``1 day'', ``2 days'', or
+``1 week'', etc. The following code chunk generates Figure
+\ref{fig:wrap-ped}.
 
@@
 
-ui <- fluidPage(tsibbleDiceUI("dice"))
+ui <- fluidPage(
+  tsibbleDiceUI("dice")
+)
 server <- function(input, output, session) {
@@
 
-Standard seasonal plots break the overall time into two components:
-seasonal periods on the x-axis, and grouped by their corresponding
-lower-resolution time. For example, monthly data can be decomposed into
-months separated by years, and hourly data into hours grouped by days.
-Data collected at lower-level resolutions often exhibits more than one
-seasonal patterns. To discover typical seasonal or non-typical profiles,
-it is helpful to quickly browse through many possible periods.
-Interactivity ought to be enabled.
+The initial state is
 

### 2020-09-06 scripts/pedestrian.R

@@ -24,7 +24,9 @@ p_line <- pedestrian20 %>%
   scale_colour_brewer(palette = "Dark2") +
   theme(legend.position = "none")
 
-ui <- fluidPage(tsibbleDiceUI("dice"))
+ui <- fluidPage(
+  tsibbleDiceUI("dice")
+)
 server <- function(input, output, session) {
   tsibbleDiceServer("dice", ggplotly(p_line, height = 700), period = "1 day")
 }

### 2020-09-07 tsibbletalk.Rmd
done with slicing and dicing

@@
 It expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots).
-As the function name suggests, a (shared) tsibble is needed to start the engine.
-The `period` argument is to specify the minimum number of seasonal blocks, for example data shifted by "1 day", "2 days", or "1 week", etc.
+As the function name suggests, a (shared) tsibble is needed to start the engine, and thereby the time `index` can be retrieved for dissection.
+The `period` argument is to specify the minimum number of seasonal blocks as desired, for example data shifted by "1 day", "2 days", or "1 week", etc.
 The following code chunk generates Figure \ref{fig:wrap-ped}.
@@
 
-The initial state is
+Upon running the shiny application, Figure \ref{fig:wrap-ped-1} corresponds to the initial state, with the slider incremented by 1-day unit.
+The "play" button near the end of slider begins animating the slicing and dicing process by walking through all 24 hours by 153 days.
+Alternatively, users can drag the handler to poke around certain slices themselves.
 
-(Data flows) Transformed data generally requires redrawing the plot, and worsen the performance of shiny.
-The underlying tsibble data is called back and transformed in R.
-Using the `plotly.js` react method, only transformed data is sent to the server side, while keeping the rest configuration unchanged (e.g.
-layout and graphical elements).
-It is performant, and users will not experience notable delay in response to the change in the slider input.
-Dissect time index, and propagate transformed data to shiny server.
+In response to the slider input, the plot will be updated and loaded with newly transformed data.
+Keeping the application as performant as possible is the top priority.
+Without completely redrawing the plot, the `plotly.js` react method is invoked internally.
+The underlying tsibble data is being called back and processed in R.
+Only transformed data gets fed back to the shiny server, along with reseting the x-axis ranges and breaks.
+The rest plot configurations, such as marks, y-axes, and layouts, are properly cached.
+
+The new shiny module exploits the temporal aspect for a tsibble object. It allows users to slide through relative periods to explore seasonal insights, with slick user experience.
 
@@
 
-# References {-}
+# References

### 2020-09-07 tsibbletalk.tex
done with slicing and dicing

@@
 As the function name suggests, a (shared) tsibble is needed to start the
-engine. The \texttt{period} argument is to specify the minimum number of
-seasonal blocks, for example data shifted by ``1 day'', ``2 days'', or
-``1 week'', etc. The following code chunk generates Figure
-\ref{fig:wrap-ped}.
+engine, and thereby the time \texttt{index} can be retrieved for
+dissection. The \texttt{period} argument is to specify the minimum
+number of seasonal blocks as desired, for example data shifted by ``1
+day'', ``2 days'', or ``1 week'', etc. The following code chunk
+generates Figure \ref{fig:wrap-ped}.
 
@@
 
-The initial state is
+Upon running the shiny application, Figure \ref{fig:wrap-ped-1}
+corresponds to the initial state, with the slider incremented by 1-day
+unit. The ``play'' button near the end of slider begins animating the
+slicing and dicing process by walking through all 24 hours by 153 days.
+Alternatively, users can drag the handler to poke around certain slices
+themselves.
 
-(Data flows) Transformed data generally requires redrawing the plot, and
-worsen the performance of shiny. The underlying tsibble data is called
-back and transformed in R. Using the \texttt{plotly.js} react method,
-only transformed data is sent to the server side, while keeping the rest
-configuration unchanged (e.g.~layout and graphical elements). It is
-performant, and users will not experience notable delay in response to
-the change in the slider input. Dissect time index, and propagate
-transformed data to shiny server.
+In response to the slider input, the plot will be updated and loaded
+with newly transformed data. Keeping the application as performant as
+possible is the top priority. Without completely redrawing the plot, the
+\texttt{plotly.js} react method is invoked internally. The underlying
+tsibble data is being called back and processed in R. Only transformed
+data gets fed back to the shiny server, along with reseting the x-axis
+ranges and breaks. The rest plot configurations, such as marks, y-axes,
+and layouts, are properly cached.
+
+The new shiny module exploits the temporal aspect for a tsibble object.
+It allows users to slide through relative periods to explore seasonal
+insights, with slick user experience.
 

### 2020-09-09 tsibbletalk.Rmd
minor tweaks

@@
 
+* [crossfilter.js](http://crossfilter.github.io/crossfilter/) & [dc.js](https://dc-js.github.io/dc.js/)
 * {cranvas} and {cranvastime}
-* [crossfilter.js](http://crossfilter.github.io/crossfilter/) & [dc.js](https://dc-js.github.io/dc.js/)
 * {crosstalk} and html widgets
@@
 Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four locations, with the depiction of three pronounced slices in time.
-Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on their absolute time scales, facetted by four sensors.
+Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on their absolute time line, facetted by four sensors.
 On March 16, Melbourne went to the stage three lockdown due to COVID-19, seeing a significant decline in traffic volume across the city.
@@
 In response to the slider input, the plot will be updated and loaded with newly transformed data.
-Keeping the application as performant as possible is the top priority.
+At its core, keeping the application as performant as possible is the top priority.
 Without completely redrawing the plot, the `plotly.js` react method is invoked internally.

### 2020-09-13 tsibbletalk.Rmd
review interactivity

@@
 * An ensemble of graphics
-* Accelerate the exploratory data visualization process
+* Accelerate the exploratory data visualisation process
 
@@
 The \CRANpkg{tsibble} package [@wang2020tsibble] introduces a unified temporal data structure, referred to as a `tsibble`, to represent time series and longitudinal data in a tidy format [@wickham2014tidy].
-That said, a `tsibble` extends the `data.frame` and \CRANpkg{tibble} class with temporally contextual metadata: `index` and `key`.
+That said, a `tsibble` extends the `data.frame` and \CRANpkg{tibble} classes with temporally contextual metadata: `index` and `key`.
 The `index` declares a data column that holds time-related indices.
@@
 
-* [crossfilter.js](http://crossfilter.github.io/crossfilter/) & [dc.js](https://dc-js.github.io/dc.js/)
-* {cranvas} and {cranvastime}
-* {crosstalk} and html widgets
-* {rJava}, {rbokeh} {loon}
+The R community has a long-standing contribution to interactive visualisation systems in abundance.
+These systems can be roughly divided into two classes: ones with web technology and the others without.
+
+R \CRANpkg{shiny} and \CRANpkg{htmlwidgets} lay a fundamental infrastructure in marrying R with HTML elements and Javascript for interactivity, although they do not emphasise on interactive graphics.
+The \CRANpkg{htmlwidgets} package opens a door to embedding Javascript libraries into R such that users are able to write only R code to generate web-based plots.
+Many Javascript charting libraries have been ported to R as HTML widgets, including \CRANpkg{plotly}, \CRANpkg{rbokeh}, and \CRANpkg{leaflet} for maps.
+To enable interactions between different widgets, it can be achieved with \CRANpkg{shiny} or \CRANpkg{crosstalk}.
+The \CRANpkg{crosstalk} extends \CRANpkg{htmlwidgets} with shared R6 instances to support linked brushing and filtering across widgets.
+
+If not using web technology, two examples are provided here: \CRANpkg{loon} based on Tcl/Tk and \CRANpkg{cranvas} based on Qt.
+They offer a wide array of pre-defined interactions, such as selecting and zooming, to manipulate plots via mouse, keyboard strokes, and menus.
+The \CRANpkg{cranvastime} is an add-on of \CRANpkg{cranvas}, which develops a set of novel interactions for temporal data, for example wrapping and mirroring.
 
@@
 
-The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble data structure on top of a `tsibble` to allow for frictionless communication between different plots for temporal data.
+The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble instance on top of a `tsibble` to allow for frictionless communication between different plots for temporal data.
 The `as_shared_tsibble()` function provides an entry point in the integrated flow, turning a `tsibble` to a shared instance (i.e.
@@
 
-In summary, shared tsibble data nicely bridges between the \CRANpkg{crosstalk} and **tidyverts** ecosystems.
+In summary, shared tsibble data nicely bridges between the \CRANpkg{crosstalk} and **tidyverts** ecosystems for temporal data over their common term "key".
 The `as_shared_tsibble()` provides a symbolic user interface for effortless construction of a hybrid of hierarchical and categorical linkings.
@@
 This shiny module, decoupled to `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece in a shiny application.
+A shiny module provides the vehicle in modularising shiny applications for both users and developers.
 Like all shiny modules, the first argument in both functions requires a user-supplied session id that must be unique.
 The UI function `tsibbleDiceUI()` simply shows a slider that animates or controls the number of periods to be diced.
-The workhorse is certainly the server function `tsibbleDiceServer()`, encapsulating the core algorithm that transforms data and sends messages to update the plot accordingly.
-It expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots).
+The workhorse is certainly the server function `tsibbleDiceServer()`, encapsulating the algorithm that transforms data and sends messages to update the plot accordingly.
+The `plot` argument expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots).
 As the function name suggests, a (shared) tsibble is needed to start the engine, and thereby the time `index` can be retrieved for dissection.
-The `period` argument is to specify the minimum number of seasonal blocks as desired, for example data shifted by "1 day", "2 days", or "1 week", etc.
-The following code chunk generates Figure \ref{fig:wrap-ped}.
+The `period` option semantically takes a desired number of seasonal periods to be shifted, for example data shifted by "1 day", "2 days", or "1 week", etc.
+In other words, the `period` defines the grind level.
+As for date-times (represented by `POSIXt`), it ranges from from fine "day" to coarse "year".
+The following code snippet generates Figure \ref{fig:wrap-ped}.
 

### 2020-09-13 tsibbletalk.tex
review interactivity

@@
 \item
-  Accelerate the exploratory data visualization process
+  Accelerate the exploratory data visualisation process
 \end{itemize}
@@
 \citep{wickham2014tidy}. That said, a \texttt{tsibble} extends the
-\texttt{data.frame} and \CRANpkg{tibble} class with temporally
+\texttt{data.frame} and \CRANpkg{tibble} classes with temporally
 contextual metadata: \texttt{index} and \texttt{key}. The \texttt{index}
@@
 
-\begin{itemize}
-\tightlist
-\item
-  \{cranvas\} and \{cranvastime\}
-\item
-  \href{http://crossfilter.github.io/crossfilter/}{crossfilter.js} \&
-  \href{https://dc-js.github.io/dc.js/}{dc.js}
-\item
-  \{crosstalk\} and html widgets
-\item
-  \{rJava\}, \{rbokeh\} \{loon\}
-\end{itemize}
+The R community has a long-standing contribution to interactive
+visualisation systems in abundance. These systems can be roughly divided
+into two classes: ones with web technology and the others without.
+
+R \CRANpkg{shiny} and \CRANpkg{htmlwidgets} lay a fundamental
+infrastructure in marrying R with HTML elements and Javascript for
+interactivity, although they do not emphasise on interactive graphics.
+The \CRANpkg{htmlwidgets} package opens a door to embedding Javascript
+libraries into R such that users are able to write only R code to
+generate web-based plots. Many Javascript charting libraries have been
+ported to R as HTML widgets, including \CRANpkg{plotly},
+\CRANpkg{rbokeh}, and \CRANpkg{leaflet} for maps. To enable interactions
+between different widgets, it can be achieved with \CRANpkg{shiny} or
+\CRANpkg{crosstalk}. The \CRANpkg{crosstalk} extends
+\CRANpkg{htmlwidgets} with shared R6 instances to support linked
+brushing and filtering across widgets.
+
+If not using web technology, two examples are provided here:
+\CRANpkg{loon} based on Tcl/Tk and \CRANpkg{cranvas} based on Qt. They
+offer a wide array of pre-defined interactions, such as selecting and
+zooming, to manipulate plots via mouse, keyboard strokes, and menus. The
+\CRANpkg{cranvastime} is an add-on of \CRANpkg{cranvas}, which develops
+a set of novel interactions for temporal data, for example wrapping and
+mirroring.
 
@@
 The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk}
-package, introduces a shared tsibble data structure on top of a
+package, introduces a shared tsibble instance on top of a
 \texttt{tsibble} to allow for frictionless communication between
@@
 In summary, shared tsibble data nicely bridges between the
-\CRANpkg{crosstalk} and \textbf{tidyverts} ecosystems. The
-\texttt{as\_shared\_tsibble()} provides a symbolic user interface for
-effortless construction of a hybrid of hierarchical and categorical
-linkings. And the \texttt{plotly\_key\_tree()} in turn decodes the
-specification to plot a tree for data overview and navigation,
-accompanied with more detailed plots.
+\CRANpkg{crosstalk} and \textbf{tidyverts} ecosystems for temporal data
+over their common term ``key''. The \texttt{as\_shared\_tsibble()}
+provides a symbolic user interface for effortless construction of a
+hybrid of hierarchical and categorical linkings. And the
+\texttt{plotly\_key\_tree()} in turn decodes the specification to plot a
+tree for data overview and navigation, accompanied with more detailed
+plots.
 
@@
 Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on
-their absolute time scales, facetted by four sensors. On March 16,
+their absolute time line, facetted by four sensors. On March 16,
 Melbourne went to the stage three lockdown due to COVID-19, seeing a
@@
 \texttt{tsibbleDiceServer()}, presents a clean interface and forms a
-resusable piece in a shiny application. Like all shiny modules, the
-first argument in both functions requires a user-supplied session id
-that must be unique. The UI function \texttt{tsibbleDiceUI()} simply
-shows a slider that animates or controls the number of periods to be
-diced. The workhorse is certainly the server function
-\texttt{tsibbleDiceServer()}, encapsulating the core algorithm that
-transforms data and sends messages to update the plot accordingly. It
-expects a \texttt{ggplot} or \texttt{plotly} object, where one can plot
-data using either lines or other graphical elements (such as boxplots).
-As the function name suggests, a (shared) tsibble is needed to start the
-engine, and thereby the time \texttt{index} can be retrieved for
-dissection. The \texttt{period} argument is to specify the minimum
-number of seasonal blocks as desired, for example data shifted by ``1
-day'', ``2 days'', or ``1 week'', etc. The following code chunk
+resusable piece in a shiny application. A shiny module provides the
+vehicle in modularising shiny applications for both users and
+developers. Like all shiny modules, the first argument in both functions
+requires a user-supplied session id that must be unique. The UI function
+\texttt{tsibbleDiceUI()} simply shows a slider that animates or controls
+the number of periods to be diced. The workhorse is certainly the server
+function \texttt{tsibbleDiceServer()}, encapsulating the algorithm that
+transforms data and sends messages to update the plot accordingly. The
+\texttt{plot} argument expects a \texttt{ggplot} or \texttt{plotly}
+object, where one can plot data using either lines or other graphical
+elements (such as boxplots). As the function name suggests, a (shared)
+tsibble is needed to start the engine, and thereby the time
+\texttt{index} can be retrieved for dissection. The \texttt{period}
+option semantically takes a desired number of seasonal periods to be
+shifted, for example data shifted by ``1 day'', ``2 days'', or ``1
+week'', etc. In other words, the \texttt{period} defines the grind
+level. As for date-times (represented by \texttt{POSIXt}), it ranges
+from from fine ``day'' to coarse ``year''. The following code snippet
 generates Figure \ref{fig:wrap-ped}.
@@
 In response to the slider input, the plot will be updated and loaded
-with newly transformed data. Keeping the application as performant as
-possible is the top priority. Without completely redrawing the plot, the
-\texttt{plotly.js} react method is invoked internally. The underlying
-tsibble data is being called back and processed in R. Only transformed
-data gets fed back to the shiny server, along with reseting the x-axis
-ranges and breaks. The rest plot configurations, such as marks, y-axes,
-and layouts, are properly cached.
+with newly transformed data. At its core, keeping the application as
+performant as possible is the top priority. Without completely redrawing
+the plot, the \texttt{plotly.js} react method is invoked internally. The
+underlying tsibble data is being called back and processed in R. Only
+transformed data gets fed back to the shiny server, along with reseting
+the x-axis ranges and breaks. The rest plot configurations, such as
+marks, y-axes, and layouts, are properly cached.
 

### 2020-09-13 tsibbletalk.Rmd
more bibs

@@
   affiliation: The University of Auckland
-  address:
-  - Department of Statistics
-  - The University of Auckland
-  - New Zealand
+  address: Auckland, New Zealand
   email: earo.wang@auckland.ac.nz
@@
   affiliation: Monash University
-  address:
-  - Department of Econometrics and Business Statistics
-  - Monash University
-  - Australia
+  address: Melbourne, Australia
   email: dicook@monash.edu
@@
 The `key` identifies a collection of related series or panels observed over the `index`-defined period, which can comprise multiple columns.
-Below displays the monthly Australian retail trade turnover data (`aus_retail`), available in the \CRANpkg{tsibbledata} package.
+Below displays the monthly Australian retail trade turnover data (`aus_retail`), available in the \CRANpkg{tsibbledata} package [@R-tsibbledata].
 The `Month` column holds year-months as `index`.
@@
 In the spirit of tidy data to the \CRANpkg{tidyverse} [@Wickham2019], the **tidyverts** suite features `tsibble` as the foundational data structure, in order to build a fluid and fluent pipeline for time series analysis.
-Besides \CRANpkg{tsibble}, the \CRANpkg{feasts} and \CRANpkg{fable} packages fill the role of statistical analysis and forecasting in the **tidyverts** ecosystem.
+Besides \CRANpkg{tsibble}, the \CRANpkg{feasts} [@R-feasts] and \CRANpkg{fable} [@R-fable] packages fill the role of statistical analysis and forecasting in the **tidyverts** ecosystem.
 When time series analysis starts taking off, series of interest denoted by the `key` variables often remain unchanged over the course of analysis, from trend inspection to forecasting performance.
@@
 
-R \CRANpkg{shiny} and \CRANpkg{htmlwidgets} lay a fundamental infrastructure in marrying R with HTML elements and Javascript for interactivity, although they do not emphasise on interactive graphics.
+R \CRANpkg{shiny} [@R-shiny] and \CRANpkg{htmlwidgets} [@R-htmlwidgets] lay a fundamental infrastructure in marrying R with HTML elements and Javascript for interactivity, although they do not emphasise on interactive graphics.
 The \CRANpkg{htmlwidgets} package opens a door to embedding Javascript libraries into R such that users are able to write only R code to generate web-based plots.
-Many Javascript charting libraries have been ported to R as HTML widgets, including \CRANpkg{plotly}, \CRANpkg{rbokeh}, and \CRANpkg{leaflet} for maps.
-To enable interactions between different widgets, it can be achieved with \CRANpkg{shiny} or \CRANpkg{crosstalk}.
-The \CRANpkg{crosstalk} extends \CRANpkg{htmlwidgets} with shared R6 instances to support linked brushing and filtering across widgets.
+Many Javascript charting libraries have been ported to R as HTML widgets, including \CRANpkg{plotly} [@plotly2020], \CRANpkg{rbokeh} [@R-rbokeh], and \CRANpkg{leaflet} [@R-leaflet] for maps.
+To enable interactions between different widgets, it can be achieved with \CRANpkg{shiny} or \CRANpkg{crosstalk} [@R-crosstalk].
+The \CRANpkg{crosstalk} extends \CRANpkg{htmlwidgets} with shared R6 instances to support linked brushing and filtering across widgets, without relying on \CRANpkg{shiny}.
 
-If not using web technology, two examples are provided here: \CRANpkg{loon} based on Tcl/Tk and \CRANpkg{cranvas} based on Qt.
+If not using web technology, two examples are provided here: \CRANpkg{loon} [@R-loon] based on Tcl/Tk and \pkg{cranvas} [@xie_reactive_2014] based on Qt.
 They offer a wide array of pre-defined interactions, such as selecting and zooming, to manipulate plots via mouse, keyboard strokes, and menus.
-The \CRANpkg{cranvastime} is an add-on of \CRANpkg{cranvas}, which develops a set of novel interactions for temporal data, for example wrapping and mirroring.
+The \pkg{cranvastime} package [@cheng_enabling_2016] is an add-on of \pkg{cranvas}, which develops a set of novel interactions for temporal data, for example wrapping and mirroring.
 
@@
 
-The `tourism_monthly` dataset packaged in \CRANpkg{tsibbletalk}, contains monthly domestic overnight trips across Australia, to give an illustrator of nesting and crossing.
+The `tourism_monthly` dataset [@tourism] packaged in \CRANpkg{tsibbletalk}, contains monthly domestic overnight trips across Australia, to give an illustrator of nesting and crossing.
 The `key` is comprised of three identifying variables: `State`, `Region`, and `Purpose` (of trip), in particular `State` nesting of `Region`, together crossed with `Purpose`.
@@
 
-The city of Melbourne has sensors installed to count hourly tallies of pedestrians in order to capture downtown daily rhythms (cite).
+The city of Melbourne has sensors installed to count hourly tallies of pedestrians in order to capture downtown daily rhythms [@ped].
 Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four locations, with the depiction of three pronounced slices in time.
@@
 
-This shiny module, decoupled to `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece in a shiny application.
+This shiny module, decoupled to `tsibbleWrapUI()` and `tsibbleWrapServer()`, presents a clean interface and forms a resusable piece in a shiny application.
 A shiny module provides the vehicle in modularising shiny applications for both users and developers.
 Like all shiny modules, the first argument in both functions requires a user-supplied session id that must be unique.
-The UI function `tsibbleDiceUI()` simply shows a slider that animates or controls the number of periods to be diced.
-The workhorse is certainly the server function `tsibbleDiceServer()`, encapsulating the algorithm that transforms data and sends messages to update the plot accordingly.
+The UI function `tsibbleWrapUI()` simply shows a slider that animates or controls the number of periods to be diced.
+The workhorse is certainly the server function `tsibbleWrapServer()`, encapsulating the algorithm that transforms data and sends messages to update the plot accordingly.
 The `plot` argument expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots).
@@
 # References
+
+```{r pkg-bib, eval = FALSE}
+pkgs <- c("tsibbledata", "feasts", "fable", "shiny", "htmlwidgets", "plotly",
+  "rbokeh", "leaflet", "crosstalk", "loon")
+write_bib(pkgs, "rpkgs.bib")
+```

### 2020-09-13 tsibbletalk.tex
more bibs

@@
 Below displays the monthly Australian retail trade turnover data
-(\texttt{aus\_retail}), available in the \CRANpkg{tsibbledata} package.
-The \texttt{Month} column holds year-months as \texttt{index}. The
-\texttt{State} together with \texttt{Industry} are the identifiers for
-these 152 series, highlighted as \texttt{key}. Note that the column
-\texttt{Series\ ID} could be an alternative option for setting up
-\texttt{key}, but \texttt{State} and \texttt{Industry} are more readable
-and informative. The \texttt{index} and \texttt{key} are ``sticky''
-columns to a \texttt{tsibble}, forming critical pieces for fluent
-temporal data analysis later.
+(\texttt{aus\_retail}), available in the \CRANpkg{tsibbledata} package
+\citep{R-tsibbledata}. The \texttt{Month} column holds year-months as
+\texttt{index}. The \texttt{State} together with \texttt{Industry} are
+the identifiers for these 152 series, highlighted as \texttt{key}. Note
+that the column \texttt{Series\ ID} could be an alternative option for
+setting up \texttt{key}, but \texttt{State} and \texttt{Industry} are
+more readable and informative. The \texttt{index} and \texttt{key} are
+``sticky'' columns to a \texttt{tsibble}, forming critical pieces for
+fluent temporal data analysis later.
 
@@
 fluid and fluent pipeline for time series analysis. Besides
-\CRANpkg{tsibble}, the \CRANpkg{feasts} and \CRANpkg{fable} packages
-fill the role of statistical analysis and forecasting in the
-\textbf{tidyverts} ecosystem. When time series analysis starts taking
-off, series of interest denoted by the \texttt{key} variables often
-remain unchanged over the course of analysis, from trend inspection to
-forecasting performance.
+\CRANpkg{tsibble}, the \CRANpkg{feasts} \citep{R-feasts} and
+\CRANpkg{fable} \citep{R-fable} packages fill the role of statistical
+analysis and forecasting in the \textbf{tidyverts} ecosystem. When time
+series analysis starts taking off, series of interest denoted by the
+\texttt{key} variables often remain unchanged over the course of
+analysis, from trend inspection to forecasting performance.
 
@@
 
-R \CRANpkg{shiny} and \CRANpkg{htmlwidgets} lay a fundamental
-infrastructure in marrying R with HTML elements and Javascript for
-interactivity, although they do not emphasise on interactive graphics.
-The \CRANpkg{htmlwidgets} package opens a door to embedding Javascript
-libraries into R such that users are able to write only R code to
-generate web-based plots. Many Javascript charting libraries have been
-ported to R as HTML widgets, including \CRANpkg{plotly},
-\CRANpkg{rbokeh}, and \CRANpkg{leaflet} for maps. To enable interactions
-between different widgets, it can be achieved with \CRANpkg{shiny} or
-\CRANpkg{crosstalk}. The \CRANpkg{crosstalk} extends
-\CRANpkg{htmlwidgets} with shared R6 instances to support linked
-brushing and filtering across widgets.
+R \CRANpkg{shiny} \citep{R-shiny} and \CRANpkg{htmlwidgets}
+\citep{R-htmlwidgets} lay a fundamental infrastructure in marrying R
+with HTML elements and Javascript for interactivity, although they do
+not emphasise on interactive graphics. The \CRANpkg{htmlwidgets} package
+opens a door to embedding Javascript libraries into R such that users
+are able to write only R code to generate web-based plots. Many
+Javascript charting libraries have been ported to R as HTML widgets,
+including \CRANpkg{plotly} \citep{plotly2020}, \CRANpkg{rbokeh}
+\citep{R-rbokeh}, and \CRANpkg{leaflet} \citep{R-leaflet} for maps. To
+enable interactions between different widgets, it can be achieved with
+\CRANpkg{shiny} or \CRANpkg{crosstalk} \citep{R-crosstalk}. The
+\CRANpkg{crosstalk} extends \CRANpkg{htmlwidgets} with shared R6
+instances to support linked brushing and filtering across widgets,
+without relying on \CRANpkg{shiny}.
 
 If not using web technology, two examples are provided here:
-\CRANpkg{loon} based on Tcl/Tk and \CRANpkg{cranvas} based on Qt. They
-offer a wide array of pre-defined interactions, such as selecting and
-zooming, to manipulate plots via mouse, keyboard strokes, and menus. The
-\CRANpkg{cranvastime} is an add-on of \CRANpkg{cranvas}, which develops
-a set of novel interactions for temporal data, for example wrapping and
-mirroring.
+\CRANpkg{loon} \citep{R-loon} based on Tcl/Tk and \pkg{cranvas}
+\citep{xie_reactive_2014} based on Qt. They offer a wide array of
+pre-defined interactions, such as selecting and zooming, to manipulate
+plots via mouse, keyboard strokes, and menus. The \pkg{cranvastime}
+package \citep{cheng_enabling_2016} is an add-on of \pkg{cranvas}, which
+develops a set of novel interactions for temporal data, for example
+wrapping and mirroring.
 
@@
 
-The \texttt{tourism\_monthly} dataset packaged in \CRANpkg{tsibbletalk},
-contains monthly domestic overnight trips across Australia, to give an
-illustrator of nesting and crossing. The \texttt{key} is comprised of
-three identifying variables: \texttt{State}, \texttt{Region}, and
-\texttt{Purpose} (of trip), in particular \texttt{State} nesting of
-\texttt{Region}, together crossed with \texttt{Purpose}. This
-specification can be translated as follows:
+The \texttt{tourism\_monthly} dataset \citep{tourism} packaged in
+\CRANpkg{tsibbletalk}, contains monthly domestic overnight trips across
+Australia, to give an illustrator of nesting and crossing. The
+\texttt{key} is comprised of three identifying variables:
+\texttt{State}, \texttt{Region}, and \texttt{Purpose} (of trip), in
+particular \texttt{State} nesting of \texttt{Region}, together crossed
+with \texttt{Purpose}. This specification can be translated as follows:
 
@@
 The city of Melbourne has sensors installed to count hourly tallies of
-pedestrians in order to capture downtown daily rhythms (cite). Figure
-\ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at
-four locations, with the depiction of three pronounced slices in time.
-Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on
-their absolute time line, facetted by four sensors. On March 16,
+pedestrians in order to capture downtown daily rhythms \citep{ped}.
+Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot
+traffic at four locations, with the depiction of three pronounced slices
+in time. Figure \ref{fig:wrap-ped-1} unfolds all counts from January to
+May on their absolute time line, facetted by four sensors. On March 16,
 Melbourne went to the stage three lockdown due to COVID-19, seeing a
@@
 
-This shiny module, decoupled to \texttt{tsibbleDiceUI()} and
-\texttt{tsibbleDiceServer()}, presents a clean interface and forms a
+This shiny module, decoupled to \texttt{tsibbleWrapUI()} and
+\texttt{tsibbleWrapServer()}, presents a clean interface and forms a
 resusable piece in a shiny application. A shiny module provides the
@@
 requires a user-supplied session id that must be unique. The UI function
-\texttt{tsibbleDiceUI()} simply shows a slider that animates or controls
+\texttt{tsibbleWrapUI()} simply shows a slider that animates or controls
 the number of periods to be diced. The workhorse is certainly the server
-function \texttt{tsibbleDiceServer()}, encapsulating the algorithm that
+function \texttt{tsibbleWrapServer()}, encapsulating the algorithm that
 transforms data and sends messages to update the plot accordingly. The
@@
 The University of Auckland\\%
-Department of Statistics\\ The University of Auckland\\ New Zealand\\
+Auckland, New Zealand\\
 %
@@
 Monash University\\%
-Department of Econometrics and Business Statistics\\ Monash
-University\\ Australia\\
+Melbourne, Australia\\
 %

### 2020-09-13 tsibbletalk.Rmd
minor tweaks

@@
 
-R \CRANpkg{shiny} [@R-shiny] and \CRANpkg{htmlwidgets} [@R-htmlwidgets] lay a fundamental infrastructure in marrying R with HTML elements and Javascript for interactivity, although they do not emphasise on interactive graphics.
+R \CRANpkg{shiny} [@R-shiny] and \CRANpkg{htmlwidgets} [@R-htmlwidgets] lay a fundamental infrastructure in marrying R with HTML elements and Javascript for interactivity, although they seldom emphasise on interactive graphics.
 The \CRANpkg{htmlwidgets} package opens a door to embedding Javascript libraries into R such that users are able to write only R code to generate web-based plots.
@@
 In other words, the `period` defines the grind level.
-As for date-times (represented by `POSIXt`), it ranges from from fine "day" to coarse "year".
+For date-times (represented by `POSIXt`), the adjustment ranges from from fine "day" to coarse "year".
 The following code snippet generates Figure \ref{fig:wrap-ped}.

### 2020-09-13 tsibbletalk.Rmd
abstract

@@
 abstract: >
-  An abstract of less than 150 words.
+  The \pkg{tsibbletalk} package can be used to interactively explore temporal data
+  of nesting and crossing structures, and discover interesting periodic/aperiodic
+  slices. This package implements a shared `tsibble` object that allows for
+  linked brushing for coordinated views, and a shiny module that aids in wrapping
+  time lines for seasonal patterns. The basic usage of functions is demonstrated
+  in the article, wit data examples.
 preamble: |

### 2020-09-13 tsibbletalk.tex
abstract

@@
 \abstract{%
-An abstract of less than 150 words.
+The \pkg{tsibbletalk} package can be used to interactively explore
+temporal data of nesting and crossing structures, and discover
+interesting periodic/aperiodic slices. This package implements a shared
+\texttt{tsibble} object that allows for linked brushing for coordinated
+views, and a shiny module that aids in wrapping time lines for seasonal
+patterns. The basic usage of functions is demonstrated in the article,
+wit data examples.
 }
@@
 \citep{R-htmlwidgets} lay a fundamental infrastructure in marrying R
-with HTML elements and Javascript for interactivity, although they do
-not emphasise on interactive graphics. The \CRANpkg{htmlwidgets} package
-opens a door to embedding Javascript libraries into R such that users
-are able to write only R code to generate web-based plots. Many
+with HTML elements and Javascript for interactivity, although they
+seldom emphasise on interactive graphics. The \CRANpkg{htmlwidgets}
+package opens a door to embedding Javascript libraries into R such that
+users are able to write only R code to generate web-based plots. Many
 Javascript charting libraries have been ported to R as HTML widgets,
@@
 week'', etc. In other words, the \texttt{period} defines the grind
-level. As for date-times (represented by \texttt{POSIXt}), it ranges
-from from fine ``day'' to coarse ``year''. The following code snippet
-generates Figure \ref{fig:wrap-ped}.
+level. For date-times (represented by \texttt{POSIXt}), the adjustment
+ranges from from fine ``day'' to coarse ``year''. The following code
+snippet generates Figure \ref{fig:wrap-ped}.
 
@@
 ui <- fluidPage(
-  tsibbleDiceUI("dice")
+  tsibbleWrapUI("dice")
 )
 server <- function(input, output, session) {
-  tsibbleDiceServer("dice", ggplotly(p_line, height = 700), period = "1 day")
+  tsibbleWrapServer("dice", ggplotly(p_line, height = 700), period = "1 day")
 }

### 2020-09-13 scripts/pedestrian.R

@@ -25,9 +25,9 @@ p_line <- pedestrian20 %>%
   theme(legend.position = "none")
 
 ui <- fluidPage(
-  tsibbleDiceUI("dice")
+  tsibbleWrapUI("dice")
 )
 server <- function(input, output, session) {
-  tsibbleDiceServer("dice", ggplotly(p_line, height = 700), period = "1 day")
+  tsibbleWrapServer("dice", ggplotly(p_line, height = 700), period = "1 day")
 }
 shinyApp(ui, server)

### 2020-09-13 tsibbletalk.Rmd
fig caps

@@
   time lines for seasonal patterns. The basic usage of functions is demonstrated
-  in the article, wit data examples.
+  in the article, with data examples.
 preamble: |
@@
 
-```{r highlight-retail, fig.height = 6, out.width = ".49\\linewidth", fig.show = "hold", fig.cap = "ToDo", fig.subcap = c("1", "2"), fig.ncol = 2}
+```{r highlight-retail, fig.height = 6, out.width = ".49\\linewidth", fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted.
+(a) An overlaid time series plot.
+(b) A scatter plot drawn from their time series features, where each dot represents a time series from (a).", fig.subcap = c("", ""), fig.ncol = 2}
 ```
@@
 Given time series plots, click any data point on a line, highlighting the whole line as a result.
-The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking, and the `spec` argument takes one step further in constructing hybrid linking, such as hierarchical and categorical linking.
+The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking, and the `spec` argument takes one step further in constructing hybrid linking, for example hierarchical and categorical linking.
 For example, each series in the `aus_retail` data corresponds to all possible combinations of the `State` and `Industry` variables.
@@
 
-```{r tourism-linking-fig, fig.cap = "ToDo"}
+```{r tourism-linking-fig, fig.cap = "Coordinated views generated from the \\code{tourism\\_shared} with persistent linked brushing."}
 include_graphics("img/tourism-linking.png")
@@
 It interprets hierarchies in the shared tsibble's `spec` as a tree view, built with \CRANpkg{plotly}.
-The following code line produces the linked tree diagram and fills the left panel of Figure \ref{fig:tourism-linking-fig}.
+The following code line produces the linked tree diagram and spans the left panel of Figure \ref{fig:tourism-linking-fig}.
 The visual of tree hierarchy untangles a group of related series and snapshots the data organisation from a bird's eye view.
@@
 
-```{r wrap-ped, fig.show = "hold", fig.cap = "ToDo", fig.subcap = c("(1)", "(2)", "(3)"), fig.ncol = 1}
+```{r wrap-ped, fig.show = "hold", fig.cap = "The shiny module for wrapping the \\code{pedestrian20} data.", fig.subcap = c("Initial overview state", "1-day state", "7-day state"), fig.ncol = 1}
 include_graphics("img/wrap-0.png")

### 2020-09-13 tsibbletalk.tex
fig caps

@@
 patterns. The basic usage of functions is demonstrated in the article,
-wit data examples.
+with data examples.
 }
@@
 
-{\centering \subfloat[1\label{fig:highlight-retail-1}]{\includegraphics[width=.49\linewidth]{figure/highlight-retail-1} }\subfloat[2\label{fig:highlight-retail-2}]{\includegraphics[width=.49\linewidth]{figure/highlight-retail-2} }
+{\centering \subfloat[\label{fig:highlight-retail-1}]{\includegraphics[width=.49\linewidth]{figure/highlight-retail-1} }\subfloat[\label{fig:highlight-retail-2}]{\includegraphics[width=.49\linewidth]{figure/highlight-retail-2} }
 
@@
 
-\caption[ToDo]{ToDo}\label{fig:highlight-retail}
+\caption[Plots for the \code{aus\_retail} data, with the series of strongest seasonal strength highlighted]{Plots for the \code{aus\_retail} data, with the series of strongest seasonal strength highlighted.
+(a) An overlaid time series plot.
+(b) A scatter plot drawn from their time series features, where each dot represents a time series from (a).}\label{fig:highlight-retail}
 \end{figure}
@@
 linking, and the \texttt{spec} argument takes one step further in
-constructing hybrid linking, such as hierarchical and categorical
+constructing hybrid linking, for example hierarchical and categorical
 linking. For example, each series in the \texttt{aus\_retail} data
@@
 
-\caption[ToDo]{ToDo}\label{fig:tourism-linking-fig}
+\caption[Coordinated views generated from the \code{tourism\_shared} with persistent linked brushing]{Coordinated views generated from the \code{tourism\_shared} with persistent linked brushing.}\label{fig:tourism-linking-fig}
 \end{figure}
@@
 \CRANpkg{plotly}. The following code line produces the linked tree
-diagram and fills the left panel of Figure
+diagram and spans the left panel of Figure
 \ref{fig:tourism-linking-fig}. The visual of tree hierarchy untangles a
@@
 
-{\centering \subfloat[(1)\label{fig:wrap-ped-1}]{\includegraphics[width=\textwidth]{img/wrap-0} }\newline\subfloat[(2)\label{fig:wrap-ped-2}]{\includegraphics[width=\textwidth]{img/wrap-1} }\newline\subfloat[(3)\label{fig:wrap-ped-3}]{\includegraphics[width=\textwidth]{img/wrap-7} }
+{\centering \subfloat[Initial overview state\label{fig:wrap-ped-1}]{\includegraphics[width=\textwidth]{img/wrap-0} }\newline\subfloat[1-day state\label{fig:wrap-ped-2}]{\includegraphics[width=\textwidth]{img/wrap-1} }\newline\subfloat[7-day state\label{fig:wrap-ped-3}]{\includegraphics[width=\textwidth]{img/wrap-7} }
 
@@
 
-\caption[ToDo]{ToDo}\label{fig:wrap-ped}
+\caption[The shiny module for wrapping the \code{pedestrian20} data]{The shiny module for wrapping the \code{pedestrian20} data.}\label{fig:wrap-ped}
 \end{figure}

### 2020-09-14 tsibbletalk.Rmd
started intro and summary

@@
   affiliation: The University of Auckland
-  address: Auckland, New Zealand
+  address: Department of Statistics
   email: earo.wang@auckland.ac.nz
@@
   affiliation: Monash University
-  address: Melbourne, Australia
+  address: Department of Econometrics and Business Statistics
   email: dicook@monash.edu
@@
   linked brushing for coordinated views, and a shiny module that aids in wrapping
-  time lines for seasonal patterns. The basic usage of functions is demonstrated
+  time lines for seasonal patterns. Basic usage of functions is demonstrated
   in the article, with data examples.
@@
 
-* An ensemble of graphics
-* Accelerate the exploratory data visualisation process
+A group of cohesive observational units measured over time typically form one temporal data set.
+Ensemble graphics [@unwin_ensemble_2018] that bundle multi-views for a data set together, are essential for exploring and mining temporal data in multi-facets.
+Adding interactivity, such as linked brushing via common observations and wrapping by time, can accelerate this process.
+
+There are a plenty of R packages, but few
+
+This article demonstrates basic usage of the \CRANpkg{tsibbletalk} package with temporal data applications for linked views and time wrapping, proceeded by brief reviews of temporal data structure and interactive graphics.
 
@@
 
-R \CRANpkg{shiny} [@R-shiny] and \CRANpkg{htmlwidgets} [@R-htmlwidgets] lay a fundamental infrastructure in marrying R with HTML elements and Javascript for interactivity, although they seldom emphasise on interactive graphics.
-The \CRANpkg{htmlwidgets} package opens a door to embedding Javascript libraries into R such that users are able to write only R code to generate web-based plots.
-Many Javascript charting libraries have been ported to R as HTML widgets, including \CRANpkg{plotly} [@plotly2020], \CRANpkg{rbokeh} [@R-rbokeh], and \CRANpkg{leaflet} [@R-leaflet] for maps.
+R \CRANpkg{shiny} [@R-shiny] and \CRANpkg{htmlwidgets} [@R-htmlwidgets] lay a fundamental infrastructure in marrying R with HTML elements and JavaScript for interactivity, although they seldom emphasise on interactive graphics.
+The \CRANpkg{htmlwidgets} package opens a door to embedding JavaScript libraries into R such that users are able to write only R code to generate web-based plots.
+Many JavaScript charting libraries have been ported to R as HTML widgets, including \CRANpkg{plotly} [@plotly2020], \CRANpkg{rbokeh} [@R-rbokeh], and \CRANpkg{leaflet} [@R-leaflet] for maps.
 To enable interactions between different widgets, it can be achieved with \CRANpkg{shiny} or \CRANpkg{crosstalk} [@R-crosstalk].
@@
 In other words, the `period` defines the grind level.
-For date-times (represented by `POSIXt`), the adjustment ranges from from fine "day" to coarse "year".
+For date-times (represented by `POSIXt`), the granularity ranges from fine "day" to coarse "year".
 The following code snippet generates Figure \ref{fig:wrap-ped}.
@@
 At its core, keeping the application as performant as possible is the top priority.
-Without completely redrawing the plot, the `plotly.js` react method is invoked internally.
+Without completely redrawing the plot, the `plotlyProxy()` react method is invoked internally for talking to shiny.
 The underlying tsibble data is being called back and processed in R.
@@
 
-# Conclusions and discussions
+# Summary
+
+Close to the heart of the \CRANpkg{tsibbletalk} package is blending the best bits from \CRANpkg{tsibble} and \CRANpkg{crosstalk}.
 

### 2020-09-14 tsibbletalk.tex
started intro and summary

@@
 views, and a shiny module that aids in wrapping time lines for seasonal
-patterns. The basic usage of functions is demonstrated in the article,
-with data examples.
+patterns. Basic usage of functions is demonstrated in the article, with
+data examples.
 }
@@
 
-\begin{itemize}
-\tightlist
-\item
-  An ensemble of graphics
-\item
-  Accelerate the exploratory data visualisation process
-\end{itemize}
+A group of cohesive observational units measured over time typically
+form one temporal data set. Ensemble graphics
+\citep{unwin_ensemble_2018} that bundle multi-views for a data set
+together, are essential for exploring and mining temporal data in
+multi-facets. Adding interactivity, such as linked brushing via common
+observations and wrapping by time, can accelerate this process.
+
+There are a plenty of R packages, but few
+
+This article demonstrates basic usage of the \CRANpkg{tsibbletalk}
+package with temporal data applications for linked views and time
+wrapping, proceeded by brief reviews of temporal data structure and
+interactive graphics.
 
@@
 \citep{R-htmlwidgets} lay a fundamental infrastructure in marrying R
-with HTML elements and Javascript for interactivity, although they
+with HTML elements and JavaScript for interactivity, although they
 seldom emphasise on interactive graphics. The \CRANpkg{htmlwidgets}
-package opens a door to embedding Javascript libraries into R such that
+package opens a door to embedding JavaScript libraries into R such that
 users are able to write only R code to generate web-based plots. Many
-Javascript charting libraries have been ported to R as HTML widgets,
+JavaScript charting libraries have been ported to R as HTML widgets,
 including \CRANpkg{plotly} \citep{plotly2020}, \CRANpkg{rbokeh}
@@
 week'', etc. In other words, the \texttt{period} defines the grind
-level. For date-times (represented by \texttt{POSIXt}), the adjustment
-ranges from from fine ``day'' to coarse ``year''. The following code
-snippet generates Figure \ref{fig:wrap-ped}.
+level. For date-times (represented by \texttt{POSIXt}), the granularity
+ranges from fine ``day'' to coarse ``year''. The following code snippet
+generates Figure \ref{fig:wrap-ped}.
 
@@
 performant as possible is the top priority. Without completely redrawing
-the plot, the \texttt{plotly.js} react method is invoked internally. The
-underlying tsibble data is being called back and processed in R. Only
-transformed data gets fed back to the shiny server, along with reseting
-the x-axis ranges and breaks. The rest plot configurations, such as
-marks, y-axes, and layouts, are properly cached.
+the plot, the \texttt{plotlyProxy()} react method is invoked internally
+for talking to shiny. The underlying tsibble data is being called back
+and processed in R. Only transformed data gets fed back to the shiny
+server, along with reseting the x-axis ranges and breaks. The rest plot
+configurations, such as marks, y-axes, and layouts, are properly cached.
 
@@
 
-\hypertarget{conclusions-and-discussions}{%
-\section{Conclusions and
-discussions}\label{conclusions-and-discussions}}
+\hypertarget{summary}{%
+\section{Summary}\label{summary}}
+
+Close to the heart of the \CRANpkg{tsibbletalk} package is blending the
+best bits from \CRANpkg{tsibble} and \CRANpkg{crosstalk}.
 
@@
 The University of Auckland\\%
-Auckland, New Zealand\\
+Department of Statistics\\
 %
@@
 Monash University\\%
-Melbourne, Australia\\
+Department of Econometrics and Business Statistics\\
 %

### 2020-09-15 tsibbletalk.Rmd
minor tweak

@@
 A group of cohesive observational units measured over time typically form one temporal data set.
-Ensemble graphics [@unwin_ensemble_2018] that bundle multi-views for a data set together, are essential for exploring and mining temporal data in multi-facets.
+Ensemble graphics [@unwin_ensemble_2018] that bundle multi-views for a data set together, are essential for exploring and mining many different aspects of temporal data.
 Adding interactivity, such as linked brushing via common observations and wrapping by time, can accelerate this process.
@@
 
-Close to the heart of the \CRANpkg{tsibbletalk} package is blending the best bits from \CRANpkg{tsibble} and \CRANpkg{crosstalk}.
+Close to the heart of the \CRANpkg{tsibbletalk} package is blending the best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, and \CRANpkg{plotly}.
 

### 2020-09-16 tsibbletalk.Rmd
a bit more on conclusion

@@
 
-Close to the heart of the \CRANpkg{tsibbletalk} package is blending the best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, and \CRANpkg{plotly}.
+Close to the heart of the \CRANpkg{tsibbletalk} package is blending the best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly}, and \CRANpkg{shiny}.
+
+The `as_shared_tsibble()` turns a `tsibble` object to a shared data class, with an option to express any nesting and crossing structures from the `key` attribute.
+If nesting is found in the data, the `plotly_key_tree()` creates an interactive hierarchical tree to help with data overview.
+This sets the stage for hierarchical and categorical linking between multi-views from one shared tsibble.
+
+A new shiny module, `tsibbleWrapUI()` and `tsibbleWrapServer()`, provides a lens for exploring temporal aspects for a `tsibble`, in particular seasonal variations.
 

### 2020-09-16 tsibbletalk.tex
a bit more on conclusion

@@
 \citep{unwin_ensemble_2018} that bundle multi-views for a data set
-together, are essential for exploring and mining temporal data in
-multi-facets. Adding interactivity, such as linked brushing via common
-observations and wrapping by time, can accelerate this process.
+together, are essential for exploring and mining many different aspects
+of temporal data. Adding interactivity, such as linked brushing via
+common observations and wrapping by time, can accelerate this process.
 
@@
 Close to the heart of the \CRANpkg{tsibbletalk} package is blending the
-best bits from \CRANpkg{tsibble} and \CRANpkg{crosstalk}.
+best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly},
+and \CRANpkg{shiny}.
+
+The \texttt{as\_shared\_tsibble()} turns a \texttt{tsibble} object to a
+shared data class, with an option to express any nesting and crossing
+structures from the \texttt{key} attribute. If nesting is found in the
+data, the \texttt{plotly\_key\_tree()} creates an interactive
+hierarchical tree to help with data overview. This sets the stage for
+hierarchical and categorical linking between multi-views from one shared
+tsibble.
+
+A new shiny module, \texttt{tsibbleWrapUI()} and
+\texttt{tsibbleWrapServer()}, provides a lens for exploring temporal
+aspects for a \texttt{tsibble}, in particular seasonal variations.
 

### 2020-09-16 tsibbletalk.Rmd
i think i'm done. cc @dicook

@@
 Ensemble graphics [@unwin_ensemble_2018] that bundle multi-views for a data set together, are essential for exploring and mining many different aspects of temporal data.
-Adding interactivity, such as linked brushing via common observations and wrapping by time, can accelerate this process.
-
-There are a plenty of R packages, but few
+Adding interactivity, such as linked brushing via shared observations and wrapping by time, can accelerate this process.
 
@@
 
-```{r wrap-ped, fig.show = "hold", fig.cap = "The shiny module for wrapping the \\code{pedestrian20} data.", fig.subcap = c("Initial overview state", "1-day state", "7-day state"), fig.ncol = 1}
+```{r wrap-ped, fig.show = "hold", fig.cap = "The shiny module for wrapping the \\code{pedestrian20} data.", fig.subcap = c("Initial overview state", "1-day state", "7-day state, anchoring to Monday"), fig.ncol = 1}
 include_graphics("img/wrap-0.png")
@@
 Figure \ref{fig:wrap-ped} suggests that there could be more than one eye-catching slices out of many possible combinations, and thus repeated wrappings can be unwieldy.
-To visually locate an interesting slice, the \CRANpkg{tsibbletalk} package implements a shiny module, a pair of UI and server functions, to automate this wrapping procedure.
+To visually locate a salient slice, the \CRANpkg{tsibbletalk} package implements a shiny module, a pair of UI and server functions, to automate this wrapping procedure.
 
@@
 
-The new shiny module exploits the temporal aspect for a tsibble object. It allows users to slide through relative periods to explore seasonal insights, with slick user experience.
+The new shiny module exploits the temporal aspect for a tsibble object, exposed by the `index` attribute.
+It allows users to slide through relative periods to digest seasonal behaviours, with slick user experience.
 
@@
 
-A new shiny module, `tsibbleWrapUI()` and `tsibbleWrapServer()`, provides a lens for exploring temporal aspects for a `tsibble`, in particular seasonal variations.
+A new shiny module, `tsibbleWrapUI()` and `tsibbleWrapServer()`, provides a lens for looking at temporal aspects for a `tsibble`, in particular seasonal or cyclical variations.
+The slicing and dicing technique efficiently wrap time lines for user-defined plots.
+The `plotlyProxy()` react method makes it possible to send wrapped data to the server and amend the plot straight way.
 

### 2020-09-16 tsibbletalk.tex
i think i'm done. cc @dicook

@@
 of temporal data. Adding interactivity, such as linked brushing via
-common observations and wrapping by time, can accelerate this process.
-
-There are a plenty of R packages, but few
+shared observations and wrapping by time, can accelerate this process.
 
@@
 
-{\centering \subfloat[Initial overview state\label{fig:wrap-ped-1}]{\includegraphics[width=\textwidth]{img/wrap-0} }\newline\subfloat[1-day state\label{fig:wrap-ped-2}]{\includegraphics[width=\textwidth]{img/wrap-1} }\newline\subfloat[7-day state\label{fig:wrap-ped-3}]{\includegraphics[width=\textwidth]{img/wrap-7} }
+{\centering \subfloat[Initial overview state\label{fig:wrap-ped-1}]{\includegraphics[width=\textwidth]{img/wrap-0} }\newline\subfloat[1-day state\label{fig:wrap-ped-2}]{\includegraphics[width=\textwidth]{img/wrap-1} }\newline\subfloat[7-day state, anchoring to Monday\label{fig:wrap-ped-3}]{\includegraphics[width=\textwidth]{img/wrap-7} }
 
@@
 eye-catching slices out of many possible combinations, and thus repeated
-wrappings can be unwieldy. To visually locate an interesting slice, the
+wrappings can be unwieldy. To visually locate a salient slice, the
 \CRANpkg{tsibbletalk} package implements a shiny module, a pair of UI
@@
 
-The new shiny module exploits the temporal aspect for a tsibble object.
-It allows users to slide through relative periods to explore seasonal
-insights, with slick user experience.
+The new shiny module exploits the temporal aspect for a tsibble object,
+exposed by the \texttt{index} attribute. It allows users to slide
+through relative periods to digest seasonal behaviours, with slick user
+experience.
 
@@
 A new shiny module, \texttt{tsibbleWrapUI()} and
-\texttt{tsibbleWrapServer()}, provides a lens for exploring temporal
-aspects for a \texttt{tsibble}, in particular seasonal variations.
+\texttt{tsibbleWrapServer()}, provides a lens for looking at temporal
+aspects for a \texttt{tsibble}, in particular seasonal or cyclical
+variations. The slicing and dicing technique efficiently wrap time lines
+for user-defined plots. The \texttt{plotlyProxy()} react method makes it
+possible to send wrapped data to the server and amend the plot straight
+way.
 

### 2020-09-17 tsibbletalk.Rmd
re-snapshots
revised abstract

@@
 abstract: >
-  The \pkg{tsibbletalk} package can be used to interactively explore temporal data
-  of nesting and crossing structures, and discover interesting periodic/aperiodic
-  slices. This package implements a shared `tsibble` object that allows for
-  linked brushing for coordinated views, and a shiny module that aids in wrapping
-  time lines for seasonal patterns. Basic usage of functions is demonstrated
-  in the article, with data examples.
+  Temporal data often has a hierarchical structure, defined by categorical variables describing different levels, such as political regions or sales products.
+Nesting of categorical variables produces a hierarchical structure.
+The \pkg{tsibbletalk} package is developed to allow a user to interactively explore temporal data, relative to the nested or crossed structures.
+It can help to discover differences between category levels, and uncover interesting periodic or aperiodic slices.
+The package implements a shared `tsibble` object that allows for
+  linked brushing between coordinated views, and a shiny module that aids in wrapping
+  time lines for seasonal patterns. The tools are demonstrated using two data examples: retail turnover and tourism in Australia.
 preamble: |

### 2020-09-17 tsibbletalk.tex
re-snapshots
revised abstract

@@
 \abstract{%
-The \pkg{tsibbletalk} package can be used to interactively explore
-temporal data of nesting and crossing structures, and discover
-interesting periodic/aperiodic slices. This package implements a shared
-\texttt{tsibble} object that allows for linked brushing for coordinated
-views, and a shiny module that aids in wrapping time lines for seasonal
-patterns. Basic usage of functions is demonstrated in the article, with
-data examples.
+Temporal data often has a hierarchical structure, defined by categorical
+variables describing different levels, such as political regions or
+sales products. Nesting of categorical variables produces a hierarchical
+structure. The \pkg{tsibbletalk} package is developed to allow a user to
+interactively explore temporal data, relative to the nested or crossed
+structures. It can help to discover differences between category levels,
+and uncover interesting periodic or aperiodic slices. The package
+implements a shared \texttt{tsibble} object that allows for linked
+brushing between coordinated views, and a shiny module that aids in
+wrapping time lines for seasonal patterns. The tools are demonstrated
+using two data examples: retail turnover and tourism in Australia.
 }

### 2020-09-17 tsibbletalk.Rmd
revised intro

@@
 
-A group of cohesive observational units measured over time typically form one temporal data set.
-Ensemble graphics [@unwin_ensemble_2018] that bundle multi-views for a data set together, are essential for exploring and mining many different aspects of temporal data.
-Adding interactivity, such as linked brushing via shared observations and wrapping by time, can accelerate this process.
+Temporal data typically arrives as a set of many observational units measured over time.
+Some variables may be categorical, indicating a hierarchy in the collection process, that may be measurements taken in different geographic regions, or types of products sold by the one company.
+Exploring these multiple features can be daunting.
+Ensemble graphics [@unwin_ensemble_2018] bundle multiple views of a data set together into one composite figure.
+These provide an effective approach for exploring and digesting many different aspects of temporal data.
+Adding interactivity to the ensemble can greatly enhance the exploration process.
 
-This article demonstrates basic usage of the \CRANpkg{tsibbletalk} package with temporal data applications for linked views and time wrapping, proceeded by brief reviews of temporal data structure and interactive graphics.
+This article describes new software, the \CRANpkg{tsibbletalk} package, for exploring temporal data using linked views and time wrapping.
+The next section provides some background to the approach based on setting up data structures and workflow.
+The section following explains the mechanism for constructing interactivity, to link between multiple hiearchical data objects and hence plots.
+The section after this describes the set up for interactively slicing and dicing time to wrap a series on itself to investigate periodicities.
 
@@
 Upon running the shiny application, Figure \ref{fig:wrap-ped-1} corresponds to the initial state, with the slider incremented by 1-day unit.
-The "play" button near the end of slider begins animating the slicing and dicing process by walking through all 24 hours by 153 days.
+The "play" button near the end of slider begins animating the slicing and dicing process by walking through all 24 hours by 152 days.
 Alternatively, users can drag the handler to poke around certain slices themselves.

### 2020-09-17 tsibbletalk.tex
revised intro

@@
 
-A group of cohesive observational units measured over time typically
-form one temporal data set. Ensemble graphics
-\citep{unwin_ensemble_2018} that bundle multi-views for a data set
-together, are essential for exploring and mining many different aspects
-of temporal data. Adding interactivity, such as linked brushing via
-shared observations and wrapping by time, can accelerate this process.
+Temporal data typically arrives as a set of many observational units
+measured over time. Some variables may be categorical, indicating a
+hierarchy in the collection process, that may be measurements taken in
+different geographic regions, or types of products sold by the one
+company. Exploring these multiple features can be daunting. Ensemble
+graphics \citep{unwin_ensemble_2018} bundle multiple views of a data set
+together into one composite figure. These provide an effective approach
+for exploring and digesting many different aspects of temporal data.
+Adding interactivity to the ensemble can greatly enhance the exploration
+process.
 
-This article demonstrates basic usage of the \CRANpkg{tsibbletalk}
-package with temporal data applications for linked views and time
-wrapping, proceeded by brief reviews of temporal data structure and
-interactive graphics.
+This article describes new software, the \CRANpkg{tsibbletalk} package,
+for exploring temporal data using linked views and time wrapping. The
+next section provides some background to the approach based on setting
+up data structures and workflow. The section following explains the
+mechanism for constructing interactivity, to link between multiple
+hiearchical data objects and hence plots. The section after this
+describes the set up for interactively slicing and dicing time to wrap a
+series on itself to investigate periodicities.
 
@@
 unit. The ``play'' button near the end of slider begins animating the
-slicing and dicing process by walking through all 24 hours by 153 days.
+slicing and dicing process by walking through all 24 hours by 152 days.
 Alternatively, users can drag the handler to poke around certain slices

### 2020-09-23 tsibbletalk.Rmd
background section

@@
 
-The \CRANpkg{tsibble} package [@wang2020tsibble] introduces a unified temporal data structure, referred to as a `tsibble`, to represent time series and longitudinal data in a tidy format [@wickham2014tidy].
-That said, a `tsibble` extends the `data.frame` and \CRANpkg{tibble} classes with temporally contextual metadata: `index` and `key`.
+The \CRANpkg{tsibble} package [@wang2020tsibble] introduced a unified temporal data structure, referred to as a `tsibble`, to represent time series and longitudinal data in a tidy format [@wickham2014tidy].
+A `tsibble` extends the `data.frame` and \CRANpkg{tibble} classes with the temporal contextual metadata: `index` and `key`.
 The `index` declares a data column that holds time-related indices.
 The `key` identifies a collection of related series or panels observed over the `index`-defined period, which can comprise multiple columns.
-Below displays the monthly Australian retail trade turnover data (`aus_retail`), available in the \CRANpkg{tsibbledata} package [@R-tsibbledata].
-The `Month` column holds year-months as `index`.
-The `State` together with `Industry` are the identifiers for these 152 series, highlighted as `key`.
-Note that the column `Series ID` could be an alternative option for setting up `key`, but `State` and `Industry` are more readable and informative.
-The `index` and `key` are "sticky" columns to a `tsibble`, forming critical pieces for fluent temporal data analysis later.
+An example of a `tsibble` can be found in the monthly Australian retail trade turnover data (`aus_retail`), available in the \CRANpkg{tsibbledata} package [@R-tsibbledata], shown below.
+The `Month` column holds year-months as the `index`.
+`State` and `Industry`, together, are the identifiers for these 152 series, which form the `key`.
+Note that the column `Series ID` could be an alternative option for setting up the `key`, but `State` and `Industry` are more readable and informative.
+The `index` and `key` are "sticky" columns to a `tsibble`, forming critical pieces for fluent downstream temporal data analysis.
 
@@
 
-In the spirit of tidy data to the \CRANpkg{tidyverse} [@Wickham2019], the **tidyverts** suite features `tsibble` as the foundational data structure, in order to build a fluid and fluent pipeline for time series analysis.
+In the spirit of tidy data from the \CRANpkg{tidyverse} [@Wickham2019], the **tidyverts** suite features `tsibble` as the foundational data structure, and helps to build a fluid and fluent pipeline for time series analysis.
 Besides \CRANpkg{tsibble}, the \CRANpkg{feasts} [@R-feasts] and \CRANpkg{fable} [@R-fable] packages fill the role of statistical analysis and forecasting in the **tidyverts** ecosystem.
-When time series analysis starts taking off, series of interest denoted by the `key` variables often remain unchanged over the course of analysis, from trend inspection to forecasting performance.
+During all the steps of a time series analysis, the series of interest, denoted by the `key` variable, typically persist, through the trend modeling and also forecasting.
+We would typically want to examine the series across all of the keys.
 
-Figure \ref{fig:highlight-retail-1} gives an overview of 152 series for the retail data using an overlaid time series plot, while Figure \ref{fig:highlight-retail-2} presents a scatterplot, where each series is represented by a dot in the feature space (trend versus seasonal strength).
-The plot making of Figure \ref{fig:highlight-retail-2} is aided with the `features()` function from \CRANpkg{feasts}, which summarises original data by each series down to various statistical features.
-This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, thus graphically displayed as Figure \ref{fig:highlight-retail-2}.
+Figure \ref{fig:highlight-retail} illustrates examining temporal data with many keys.
+The data has 152 series coorresponding to different items in retail data.
+The multiple series are displayed using an overlaid time series plot, along with a scatterplot of two variables (trend versus seasonal strength) from feature space, where each series is represented by a dot.
+The feature space is computed using the `features()` function from \CRANpkg{feasts}, which summarises the original data for each series using various statistical features.
+This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, which can be graphically displayed as in Figure \ref{fig:highlight-retail-2}.
 
@@
 
-Figure \ref{fig:highlight-retail} highlights not only a series with strongest seasonality, but also a need to querying interesting series on the fly.
-Without interactivity, one needs to first filter the interesting series out from the features table, and join back to the original `tsibble` in order to examine its trend in relation to others.
-This procedure can soon grow cumbersome if many series to be discovered.
-Despite that the two plots are static, they can be considered as linked views via the common `key` variables between two tables.
-This motivates enabling interactivity of `tsibble` and `tsibble`-derived objects for rapid exploratory data analysis.
+Figure \ref{fig:highlight-retail} has also been highlighted to focus on the one series with the strongest seasonality.
+To create this highlighting, one needs to first filter the interesting series from the features table, and join back to the original `tsibble` in order to examine its trend in relation to others.
+This procedure can soon grow cumbersome if many series are to be explored.
+It illustrates a need to query interesting series on the fly.
+Although these two plots are static, we can consider them as linked views becaause the common `key` variables link between the two data tables producing the two ploots.
+This motivates the work in this package, described in this paper, to enable interactivity of `tsibble` and `tsibble`-derived objects for rapid exploratory data analysis.
 

### 2020-09-23 tsibbletalk.tex
background section

@@
 
-The \CRANpkg{tsibble} package \citep{wang2020tsibble} introduces a
+The \CRANpkg{tsibble} package \citep{wang2020tsibble} introduced a
 unified temporal data structure, referred to as a \texttt{tsibble}, to
 represent time series and longitudinal data in a tidy format
-\citep{wickham2014tidy}. That said, a \texttt{tsibble} extends the
-\texttt{data.frame} and \CRANpkg{tibble} classes with temporally
+\citep{wickham2014tidy}. A \texttt{tsibble} extends the
+\texttt{data.frame} and \CRANpkg{tibble} classes with the temporal
 contextual metadata: \texttt{index} and \texttt{key}. The \texttt{index}
@@
 identifies a collection of related series or panels observed over the
-\texttt{index}-defined period, which can comprise multiple columns.
-Below displays the monthly Australian retail trade turnover data
-(\texttt{aus\_retail}), available in the \CRANpkg{tsibbledata} package
-\citep{R-tsibbledata}. The \texttt{Month} column holds year-months as
-\texttt{index}. The \texttt{State} together with \texttt{Industry} are
-the identifiers for these 152 series, highlighted as \texttt{key}. Note
-that the column \texttt{Series\ ID} could be an alternative option for
-setting up \texttt{key}, but \texttt{State} and \texttt{Industry} are
-more readable and informative. The \texttt{index} and \texttt{key} are
-``sticky'' columns to a \texttt{tsibble}, forming critical pieces for
-fluent temporal data analysis later.
+\texttt{index}-defined period, which can comprise multiple columns. An
+example of a \texttt{tsibble} can be found in the monthly Australian
+retail trade turnover data (\texttt{aus\_retail}), available in the
+\CRANpkg{tsibbledata} package \citep{R-tsibbledata}, shown below. The
+\texttt{Month} column holds year-months as the \texttt{index}.
+\texttt{State} and \texttt{Industry}, together, are the identifiers for
+these 152 series, which form the \texttt{key}. Note that the column
+\texttt{Series\ ID} could be an alternative option for setting up the
+\texttt{key}, but \texttt{State} and \texttt{Industry} are more readable
+and informative. The \texttt{index} and \texttt{key} are ``sticky''
+columns to a \texttt{tsibble}, forming critical pieces for fluent
+downstream temporal data analysis.
 
@@
 
-In the spirit of tidy data to the \CRANpkg{tidyverse}
+In the spirit of tidy data from the \CRANpkg{tidyverse}
 \citep{Wickham2019}, the \textbf{tidyverts} suite features
-\texttt{tsibble} as the foundational data structure, in order to build a
-fluid and fluent pipeline for time series analysis. Besides
+\texttt{tsibble} as the foundational data structure, and helps to build
+a fluid and fluent pipeline for time series analysis. Besides
 \CRANpkg{tsibble}, the \CRANpkg{feasts} \citep{R-feasts} and
 \CRANpkg{fable} \citep{R-fable} packages fill the role of statistical
-analysis and forecasting in the \textbf{tidyverts} ecosystem. When time
-series analysis starts taking off, series of interest denoted by the
-\texttt{key} variables often remain unchanged over the course of
-analysis, from trend inspection to forecasting performance.
+analysis and forecasting in the \textbf{tidyverts} ecosystem. During all
+the steps of a time series analysis, the series of interest, denoted by
+the \texttt{key} variable, typically persist, through the trend modeling
+and also forecasting. We would typically want to examine the series
+across all of the keys.
 
-Figure \ref{fig:highlight-retail-1} gives an overview of 152 series for
-the retail data using an overlaid time series plot, while Figure
-\ref{fig:highlight-retail-2} presents a scatterplot, where each series
-is represented by a dot in the feature space (trend versus seasonal
-strength). The plot making of Figure \ref{fig:highlight-retail-2} is
-aided with the \texttt{features()} function from \CRANpkg{feasts}, which
-summarises original data by each series down to various statistical
-features. This function along with other \textbf{tidyverts} functions is
+Figure \ref{fig:highlight-retail} illustrates examining temporal data
+with many keys. The data has 152 series coorresponding to different
+items in retail data. The multiple series are displayed using an
+overlaid time series plot, along with a scatterplot of two variables
+(trend versus seasonal strength) from feature space, where each series
+is represented by a dot. The feature space is computed using the
+\texttt{features()} function from \CRANpkg{feasts}, which summarises the
+original data for each series using various statistical features. This
+function along with other \textbf{tidyverts} functions is
 \texttt{tsibble}-aware, and outputs a table in a reduced form where each
-row corresponds to a series, thus graphically displayed as Figure
-\ref{fig:highlight-retail-2}.
+row corresponds to a series, which can be graphically displayed as in
+Figure \ref{fig:highlight-retail-2}.
 
@@
 
-Figure \ref{fig:highlight-retail} highlights not only a series with
-strongest seasonality, but also a need to querying interesting series on
-the fly. Without interactivity, one needs to first filter the
-interesting series out from the features table, and join back to the
-original \texttt{tsibble} in order to examine its trend in relation to
-others. This procedure can soon grow cumbersome if many series to be
-discovered. Despite that the two plots are static, they can be
-considered as linked views via the common \texttt{key} variables between
-two tables. This motivates enabling interactivity of \texttt{tsibble}
-and \texttt{tsibble}-derived objects for rapid exploratory data
-analysis.
+Figure \ref{fig:highlight-retail} has also been highlighted to focus on
+the one series with the strongest seasonality. To create this
+highlighting, one needs to first filter the interesting series from the
+features table, and join back to the original \texttt{tsibble} in order
+to examine its trend in relation to others. This procedure can soon grow
+cumbersome if many series are to be explored. It illustrates a need to
+query interesting series on the fly. Although these two plots are
+static, we can consider them as linked views becaause the common
+\texttt{key} variables link between the two data tables producing the
+two ploots. This motivates the work in this package, described in this
+paper, to enable interactivity of \texttt{tsibble} and
+\texttt{tsibble}-derived objects for rapid exploratory data analysis.
 

### 2020-09-23 tsibbletalk.Rmd
overview of interactivity

@@
 
-The R community has a long-standing contribution to interactive visualisation systems in abundance.
-These systems can be roughly divided into two classes: ones with web technology and the others without.
+There is a long history of interactive data visualisation research and corresponding systems. Within R, the systems can be roughly divided into systems utilising web technology and those that don't.
 
-R \CRANpkg{shiny} [@R-shiny] and \CRANpkg{htmlwidgets} [@R-htmlwidgets] lay a fundamental infrastructure in marrying R with HTML elements and JavaScript for interactivity, although they seldom emphasise on interactive graphics.
-The \CRANpkg{htmlwidgets} package opens a door to embedding JavaScript libraries into R such that users are able to write only R code to generate web-based plots.
+R \CRANpkg{shiny} [@R-shiny] and \CRANpkg{htmlwidgets} [@R-htmlwidgets] provides and infrastructure connecting R with HTML elements and JavaScript that provide the interactivity.<!--, although they seldom emphasise on interactive graphics.--> The \CRANpkg{htmlwidgets} package makes it possible to embed JavaScript libraries into R so that users are able to write only R code to generate web-based plots.
 Many JavaScript charting libraries have been ported to R as HTML widgets, including \CRANpkg{plotly} [@plotly2020], \CRANpkg{rbokeh} [@R-rbokeh], and \CRANpkg{leaflet} [@R-leaflet] for maps.
-To enable interactions between different widgets, it can be achieved with \CRANpkg{shiny} or \CRANpkg{crosstalk} [@R-crosstalk].
+Interactions between different widgets can be achieved with \CRANpkg{shiny} or \CRANpkg{crosstalk} [@R-crosstalk].
 The \CRANpkg{crosstalk} extends \CRANpkg{htmlwidgets} with shared R6 instances to support linked brushing and filtering across widgets, without relying on \CRANpkg{shiny}.
 
-If not using web technology, two examples are provided here: \CRANpkg{loon} [@R-loon] based on Tcl/Tk and \pkg{cranvas} [@xie_reactive_2014] based on Qt.
-They offer a wide array of pre-defined interactions, such as selecting and zooming, to manipulate plots via mouse, keyboard strokes, and menus.
-The \pkg{cranvastime} package [@cheng_enabling_2016] is an add-on of \pkg{cranvas}, which develops a set of novel interactions for temporal data, for example wrapping and mirroring.
+Systems without the web technology include \CRANpkg{loon} [@R-loon], based on Tcl/Tk, and \pkg{cranvas} [@xie_reactive_2014] based on Qt.
+They offer a wide array of pre-defined interactions, such as selecting and zooming, to manipulate plots via mouse action, keyboard strokes, and menus.
+The \pkg{cranvastime} package [@cheng_enabling_2016] is an add-on to \pkg{cranvas}, which provides specialised interactions for temporal data, such as wrapping and mirroring.
 

### 2020-09-23 tsibbletalk.tex
overview of interactivity

@@
 
-The R community has a long-standing contribution to interactive
-visualisation systems in abundance. These systems can be roughly divided
-into two classes: ones with web technology and the others without.
+There is a long history of interactive data visualisation research and
+corresponding systems. Within R, the systems can be roughly divided into
+systems utilising web technology and those that don't.
 
 R \CRANpkg{shiny} \citep{R-shiny} and \CRANpkg{htmlwidgets}
-\citep{R-htmlwidgets} lay a fundamental infrastructure in marrying R
-with HTML elements and JavaScript for interactivity, although they
-seldom emphasise on interactive graphics. The \CRANpkg{htmlwidgets}
-package opens a door to embedding JavaScript libraries into R such that
-users are able to write only R code to generate web-based plots. Many
-JavaScript charting libraries have been ported to R as HTML widgets,
-including \CRANpkg{plotly} \citep{plotly2020}, \CRANpkg{rbokeh}
-\citep{R-rbokeh}, and \CRANpkg{leaflet} \citep{R-leaflet} for maps. To
-enable interactions between different widgets, it can be achieved with
-\CRANpkg{shiny} or \CRANpkg{crosstalk} \citep{R-crosstalk}. The
-\CRANpkg{crosstalk} extends \CRANpkg{htmlwidgets} with shared R6
-instances to support linked brushing and filtering across widgets,
-without relying on \CRANpkg{shiny}.
+\citep{R-htmlwidgets} provides and infrastructure connecting R with HTML
+elements and JavaScript that provide the interactivity. The
+\CRANpkg{htmlwidgets} package makes it possible to embed JavaScript
+libraries into R so that users are able to write only R code to generate
+web-based plots. Many JavaScript charting libraries have been ported to
+R as HTML widgets, including \CRANpkg{plotly} \citep{plotly2020},
+\CRANpkg{rbokeh} \citep{R-rbokeh}, and \CRANpkg{leaflet}
+\citep{R-leaflet} for maps. Interactions between different widgets can
+be achieved with \CRANpkg{shiny} or \CRANpkg{crosstalk}
+\citep{R-crosstalk}. The \CRANpkg{crosstalk} extends
+\CRANpkg{htmlwidgets} with shared R6 instances to support linked
+brushing and filtering across widgets, without relying on
+\CRANpkg{shiny}.
 
-If not using web technology, two examples are provided here:
-\CRANpkg{loon} \citep{R-loon} based on Tcl/Tk and \pkg{cranvas}
+Systems without the web technology include \CRANpkg{loon}
+\citep{R-loon}, based on Tcl/Tk, and \pkg{cranvas}
 \citep{xie_reactive_2014} based on Qt. They offer a wide array of
 pre-defined interactions, such as selecting and zooming, to manipulate
-plots via mouse, keyboard strokes, and menus. The \pkg{cranvastime}
-package \citep{cheng_enabling_2016} is an add-on of \pkg{cranvas}, which
-develops a set of novel interactions for temporal data, for example
-wrapping and mirroring.
+plots via mouse action, keyboard strokes, and menus. The
+\pkg{cranvastime} package \citep{cheng_enabling_2016} is an add-on to
+\pkg{cranvas}, which provides specialised interactions for temporal
+data, such as wrapping and mirroring.
 

### 2020-09-23 tsibbletalk.Rmd
start of shared data object

@@
 
-# Interactivity for coordinated views via shared temporal data
+The techniques implemented in work described in this paper utilise web technology.
 
-The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble instance on top of a `tsibble` to allow for frictionless communication between different plots for temporal data.
-The `as_shared_tsibble()` function provides an entry point in the integrated flow, turning a `tsibble` to a shared instance (i.e.
-`SharedTsibbleData` subclassing of `SharedData` from \CRANpkg{crosstalk}) that powers data transmission across multiple views.
-The \CRANpkg{tsibbletalk} package aims to streamline interactive graphical analysis with the focus of temporal and structured linking.
+# Using a shared temporal data object for interactivity
+
+The \CRANpkg{tsibbletalk} package introduces a shared tsibble instance built on a `tsibble`.
+This allows for seamless communication between different plots of temporal data.
+The `as_shared_tsibble()` function turns a `tsibble` into a shared instance,  `SharedTsibbleData`, which is a subclass of `SharedData` from \CRANpkg{crosstalk}).
+This is the object driving data transmission across multiple views.
+The \CRANpkg{tsibbletalk} package aims to streamline interactive exploration of temporal data, with the focus of temporal elements and structured linking.
+
+## Linking between plots
 
@@
 
-# Slicing and dicing time
+## Slicing and dicing time
 

### 2020-09-23 tsibbletalk.tex
start of shared data object

@@
 
-\hypertarget{interactivity-for-coordinated-views-via-shared-temporal-data}{%
-\section{Interactivity for coordinated views via shared temporal
-data}\label{interactivity-for-coordinated-views-via-shared-temporal-data}}
+The techniques implemented in work described in this paper utilise web
+technology.
 
-The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk}
-package, introduces a shared tsibble instance on top of a
-\texttt{tsibble} to allow for frictionless communication between
-different plots for temporal data. The \texttt{as\_shared\_tsibble()}
-function provides an entry point in the integrated flow, turning a
-\texttt{tsibble} to a shared instance (i.e.~\texttt{SharedTsibbleData}
-subclassing of \texttt{SharedData} from \CRANpkg{crosstalk}) that powers
-data transmission across multiple views. The \CRANpkg{tsibbletalk}
-package aims to streamline interactive graphical analysis with the focus
-of temporal and structured linking.
+\hypertarget{using-a-shared-temporal-data-object-for-interactivity}{%
+\section{Using a shared temporal data object for
+interactivity}\label{using-a-shared-temporal-data-object-for-interactivity}}
+
+The \CRANpkg{tsibbletalk} package introduces a shared tsibble instance
+built on a \texttt{tsibble}. This allows for seamless communication
+between different plots of temporal data. The
+\texttt{as\_shared\_tsibble()} function turns a \texttt{tsibble} into a
+shared instance, \texttt{SharedTsibbleData}, which is a subclass of
+\texttt{SharedData} from \CRANpkg{crosstalk}). This is the object
+driving data transmission across multiple views. The
+\CRANpkg{tsibbletalk} package aims to streamline interactive exploration
+of temporal data, with the focus of temporal elements and structured
+linking.
+
+\hypertarget{linking-between-plots}{%
+\subsection{Linking between plots}\label{linking-between-plots}}
 
@@
 \hypertarget{slicing-and-dicing-time}{%
-\section{Slicing and dicing time}\label{slicing-and-dicing-time}}
+\subsection{Slicing and dicing time}\label{slicing-and-dicing-time}}
 

### 2020-09-23 tsibbletalk.Rmd
linking section, still more to do

@@
 
-As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical linking where marking one or more observations in one category will broadcast to all other observations in this category.
-Given time series plots, click any data point on a line, highlighting the whole line as a result.
-The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking, and the `spec` argument takes one step further in constructing hybrid linking, for example hierarchical and categorical linking.
-For example, each series in the `aus_retail` data corresponds to all possible combinations of the `State` and `Industry` variables.
-They are intrinsically crossed with each other.
-If one variable is nested within another, this lends itself to a hierarchical structure, like geographical hierarchy.
-Such collection of inter-related time series are referred to as hierarchical and grouped time series in the literature [@fpp].
+As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical variable linking where selecting one or more observations in one category will broadcast to all other observations in this category.
+In addition, linking is also by time: within the time series plot, click on any data point, and  whole line will be highlighted in response.
+The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking.
 
-To incorporate structured specifications in the `key`, a symbolic formula can be passed to the `spec` argument.
-Adopting Wilkinson notations for factorial models [@Wilkinson1973], the `spec` follows the `/` and `*` operators tradition to declare nesting and crossing variables respectively.
+The approach can also accommodate hierarchical temporal data.
+When one key variable is nested within another, such as regional areas within a state, this is considered to be a hierarchical structure.
+These time series are referred to as hierarchical and grouped time series in the literature [@fpp].
+The `aus_retail` above is an example.
+Each series in the  data corresponds to all possible combinations of the `State` and `Industry` variables, which means they are intrinsically crossed with each other.
+
+The `spec` argument in `as_shared_tsibble()`  constructing hybrid linking, that incorporates hierarchical and categorical linking.
+A symbolic formula can be passed to the `spec` argument, to incorporate structured specifications in the `key`.
+Adopting @Wilkinson1973's notation for factorial models, the `spec` follows the `/` and `*` operator conventions to declare nesting and crossing variables, respectively.
 The `spec` for the `aus_retail` data is therefore specified as `State * Industry` or `Industry * State`, which is the default for the presence of multiple `key` variables.
-If there is a hierarchy in the data, using `/` is required to indicate the parent-child relation, as strictly one direction `parent/child`.
+If there is a hierarchy in the data, using `/` is required to indicate the parent-child relation, for a strictly one directional `parent/child`.
 
-The `tourism_monthly` dataset [@tourism] packaged in \CRANpkg{tsibbletalk}, contains monthly domestic overnight trips across Australia, to give an illustrator of nesting and crossing.
+To illustrate nesting and crossing we use the `tourism_monthly` dataset [@tourism] packaged in \CRANpkg{tsibbletalk}.
+It contains monthly domestic overnight trips across Australia.
 The `key` is comprised of three identifying variables: `State`, `Region`, and `Purpose` (of trip), in particular `State` nesting of `Region`, together crossed with `Purpose`.
@@
 
-This dataset contains a three-level hierarchy: the root node is implicitly Australia, and geographically disaggregated to states and lower-level tourism regions.
-A new handy function `plotly_key_tree()` has been implemented to address the need of hierarchical discovery arising from the data.
+There is a three-level hierarchy: the root node is implicitly Australia, and geographically disaggregated to states and lower-level tourism regions.
+A new handy function `plotly_key_tree()` has been implemented to help explore the hierarchy.
 It interprets hierarchies in the shared tsibble's `spec` as a tree view, built with \CRANpkg{plotly}.
-The following code line produces the linked tree diagram and spans the left panel of Figure \ref{fig:tourism-linking-fig}.
-The visual of tree hierarchy untangles a group of related series and snapshots the data organisation from a bird's eye view.
+The following code line produces the linked tree diagram (left panel of Figure \ref{fig:tourism-linking-fig}).
+The visual for the tree hierarchy detangles a group of related series and provides a bird's eye view of the data organisation.
 

### 2020-09-23 tsibbletalk.tex
linking section, still more to do

@@
 As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to
-categorical linking where marking one or more observations in one
-category will broadcast to all other observations in this category.
-Given time series plots, click any data point on a line, highlighting
-the whole line as a result. The \texttt{as\_shared\_tsibble()} uses
-\texttt{tsibble}'s \texttt{key} variables to achieve these types of
-linking, and the \texttt{spec} argument takes one step further in
-constructing hybrid linking, for example hierarchical and categorical
-linking. For example, each series in the \texttt{aus\_retail} data
-corresponds to all possible combinations of the \texttt{State} and
-\texttt{Industry} variables. They are intrinsically crossed with each
-other. If one variable is nested within another, this lends itself to a
-hierarchical structure, like geographical hierarchy. Such collection of
-inter-related time series are referred to as hierarchical and grouped
-time series in the literature \citep{fpp}.
+categorical variable linking where selecting one or more observations in
+one category will broadcast to all other observations in this category.
+In addition, linking is also by time: within the time series plot, click
+on any data point, and whole line will be highlighted in response. The
+\texttt{as\_shared\_tsibble()} uses \texttt{tsibble}'s \texttt{key}
+variables to achieve these types of linking.
 
-To incorporate structured specifications in the \texttt{key}, a symbolic
-formula can be passed to the \texttt{spec} argument. Adopting Wilkinson
-notations for factorial models \citep{Wilkinson1973}, the \texttt{spec}
-follows the \texttt{/} and \texttt{*} operators tradition to declare
-nesting and crossing variables respectively. The \texttt{spec} for the
-\texttt{aus\_retail} data is therefore specified as
-\texttt{State\ *\ Industry} or \texttt{Industry\ *\ State}, which is the
-default for the presence of multiple \texttt{key} variables. If there is
-a hierarchy in the data, using \texttt{/} is required to indicate the
-parent-child relation, as strictly one direction \texttt{parent/child}.
+The approach can also accommodate hierarchical temporal data. When one
+key variable is nested within another, such as regional areas within a
+state, this is considered to be a hierarchical structure. These time
+series are referred to as hierarchical and grouped time series in the
+literature \citep{fpp}. The \texttt{aus\_retail} above is an example.
+Each series in the data corresponds to all possible combinations of the
+\texttt{State} and \texttt{Industry} variables, which means they are
+intrinsically crossed with each other.
 
-The \texttt{tourism\_monthly} dataset \citep{tourism} packaged in
-\CRANpkg{tsibbletalk}, contains monthly domestic overnight trips across
-Australia, to give an illustrator of nesting and crossing. The
-\texttt{key} is comprised of three identifying variables:
-\texttt{State}, \texttt{Region}, and \texttt{Purpose} (of trip), in
-particular \texttt{State} nesting of \texttt{Region}, together crossed
-with \texttt{Purpose}. This specification can be translated as follows:
+The \texttt{spec} argument in \texttt{as\_shared\_tsibble()}
+constructing hybrid linking, that incorporates hierarchical and
+categorical linking. A symbolic formula can be passed to the
+\texttt{spec} argument, to incorporate structured specifications in the
+\texttt{key}. Adopting \citet{Wilkinson1973}'s notation for factorial
+models, the \texttt{spec} follows the \texttt{/} and \texttt{*} operator
+conventions to declare nesting and crossing variables, respectively. The
+\texttt{spec} for the \texttt{aus\_retail} data is therefore specified
+as \texttt{State\ *\ Industry} or \texttt{Industry\ *\ State}, which is
+the default for the presence of multiple \texttt{key} variables. If
+there is a hierarchy in the data, using \texttt{/} is required to
+indicate the parent-child relation, for a strictly one directional
+\texttt{parent/child}.
+
+To illustrate nesting and crossing we use the \texttt{tourism\_monthly}
+dataset \citep{tourism} packaged in \CRANpkg{tsibbletalk}. It contains
+monthly domestic overnight trips across Australia. The \texttt{key} is
+comprised of three identifying variables: \texttt{State},
+\texttt{Region}, and \texttt{Purpose} (of trip), in particular
+\texttt{State} nesting of \texttt{Region}, together crossed with
+\texttt{Purpose}. This specification can be translated as follows:
 
@@
 
-This dataset contains a three-level hierarchy: the root node is
-implicitly Australia, and geographically disaggregated to states and
-lower-level tourism regions. A new handy function
-\texttt{plotly\_key\_tree()} has been implemented to address the need of
-hierarchical discovery arising from the data. It interprets hierarchies
-in the shared tsibble's \texttt{spec} as a tree view, built with
+There is a three-level hierarchy: the root node is implicitly Australia,
+and geographically disaggregated to states and lower-level tourism
+regions. A new handy function \texttt{plotly\_key\_tree()} has been
+implemented to help explore the hierarchy. It interprets hierarchies in
+the shared tsibble's \texttt{spec} as a tree view, built with
 \CRANpkg{plotly}. The following code line produces the linked tree
-diagram and spans the left panel of Figure
-\ref{fig:tourism-linking-fig}. The visual of tree hierarchy untangles a
-group of related series and snapshots the data organisation from a
-bird's eye view.
+diagram (left panel of Figure \ref{fig:tourism-linking-fig}). The visual
+for the tree hierarchy detangles a group of related series and provides
+a bird's eye view of the data organisation.
 

### 2020-09-23 scripts/demo.R

@@ -11,9 +11,9 @@ library(tidyverse)
 library(ggrepel)
 library(gghighlight)
 
-aus_retail %>% 
+aus_retail %>%
   as_tibble() %>% # gghighlight issue for group_by() + filter()
-  mutate(group = paste(State, ":", Industry)) %>% 
+  mutate(group = paste(State, ":", Industry)) %>%
   ggplot(aes(x = Month, y = Turnover)) +
   geom_line(aes(group = group)) +
   gghighlight(
@@ -22,9 +22,9 @@ aus_retail %>%
   ) +
   scale_x_yearmonth(breaks = yearmonth(c("1990 Jan", "2000 Jan", "2010 Jan")))
 
-aus_retail %>% 
-  features(Turnover, feat_stl) %>% 
-  mutate(group = paste(State, ":", Industry)) %>% 
+aus_retail %>%
+  features(Turnover, feat_stl) %>%
+  mutate(group = paste(State, ":", Industry)) %>%
   ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
   geom_point(aes(group = group)) +
   labs(x = "Trend strength", y = "Seasonal strength") +
@@ -33,3 +33,4 @@ aus_retail %>%
     use_direct_label = FALSE
   ) +
   geom_label_repel(aes(label = group), vjust = 2, nudge_x = -0.1)
+

### 2020-09-23 tsibbletalk.Rmd
still on linking

@@
 
-There is a three-level hierarchy: the root node is implicitly Australia, and geographically disaggregated to states and lower-level tourism regions.
+\noindent There is a three-level hierarchy: the root node is implicitly Australia, and geographically disaggregated to states and lower-level tourism regions.
 A new handy function `plotly_key_tree()` has been implemented to help explore the hierarchy.
@@
 
-The tree plot provides backbones of the data, and much flesh yet to be attached.
-Small multiples of time series lines are composed and placed at the top right of Figure \ref{fig:tourism-linking-fig} to unpack the temporal trend across regions by purposes of trips.
-The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code.
+The tree plot provides the data skeleton, upon which the rest of data can be attached.
+In this example, small multiples of line plots are placed at the top right of Figure \ref{fig:tourism-linking-fig} to explore the temporal trend across regions, by trip purpose.
+The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code as follows:
 

### 2020-09-23 tsibbletalk.tex
still on linking

@@
 
-There is a three-level hierarchy: the root node is implicitly Australia,
-and geographically disaggregated to states and lower-level tourism
-regions. A new handy function \texttt{plotly\_key\_tree()} has been
-implemented to help explore the hierarchy. It interprets hierarchies in
-the shared tsibble's \texttt{spec} as a tree view, built with
-\CRANpkg{plotly}. The following code line produces the linked tree
+\noindent There is a three-level hierarchy: the root node is implicitly
+Australia, and geographically disaggregated to states and lower-level
+tourism regions. A new handy function \texttt{plotly\_key\_tree()} has
+been implemented to help explore the hierarchy. It interprets
+hierarchies in the shared tsibble's \texttt{spec} as a tree view, built
+with \CRANpkg{plotly}. The following code line produces the linked tree
 diagram (left panel of Figure \ref{fig:tourism-linking-fig}). The visual
@@
 
-The tree plot provides backbones of the data, and much flesh yet to be
-attached. Small multiples of time series lines are composed and placed
-at the top right of Figure \ref{fig:tourism-linking-fig} to unpack the
-temporal trend across regions by purposes of trips. The shared tsibble
-data can be directly piped into \CRANpkg{ggplot2} code.
+The tree plot provides the data skeleton, upon which the rest of data
+can be attached. In this example, small multiples of line plots are
+placed at the top right of Figure \ref{fig:tourism-linking-fig} to
+explore the temporal trend across regions, by trip purpose. The shared
+tsibble data can be directly piped into \CRANpkg{ggplot2} code as
+follows.
 

### 2020-09-23 tsibbletalk.Rmd
done with linking section

@@
 
-The tree plot provides the data skeleton, upon which the rest of data can be attached.
+The tree plot provides the graphics skeleton, upon which the rest of data plots can be attached.
 In this example, small multiples of line plots are placed at the top right of Figure \ref{fig:tourism-linking-fig} to explore the temporal trend across regions, by trip purpose.
-The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code as follows:
+The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code to create this.
 
@@
 
-To tease apart these overlaid time series, they are funnelled through the `features()` S3 method to extract some key characteristics, including the measurements of trend and seasonality.
-A scatterplot is populated from these statistics for each series.
+These line plots are heavily overplotted.
+To tease apart structure in the multiple time series, the `features()` function `fabletools` computes interesting characteristics, including the measures of trend and seasonality.
+These are displayed in the scatterplot at bottom right, where one dot represents one series.
 
@@
 
-Lastly, three graphics are composed as an ensemble of coordinated views for multi-facetted exploration, shown as Figure \ref{fig:tourism-linking-fig} (the interactive realisation of Figure \ref{fig:highlight-retail}).
-Routine functions bring about new interaction with temporal data on the client side.
+There is one final step, to compose the three plots into an ensemble of coordinated views for exploration, shown in Figure \ref{fig:tourism-linking-fig}.
+(This is the interactive realisation of Figure \ref{fig:highlight-retail}).
+<!-- Routine functions bring about new interaction with temporal data on the client side.-->
 
@@
 
-Since all plots are stemmed from one shared tsibble data source, they are self-linking views.
+Since all plots are created from one shared tsibble data source, they are self-linking views.
 Nodes, lines, and points are hoverable and clickable.
-Given the `spec`, clicking either one element in any plot highlights all points that match the `Region` category, briefly "categorical linking".
-In Figure \ref{fig:tourism-linking-fig}, when hovering and selecting the circle associated with "Sydney" in the scatter plot, all data records with shared values of "Sydney" listen and react to this interaction via self updating in red.
-In order for comparison with other regions or states, press the "Shift" key to enable persistent selection, and simultaneously select the parent node on the tree, saying "Western Australia", to include all the children by switching to the blue colour.
-The domestic tourism sees Sydney as one of the most popular destinations in realm of business and friends visiting over years.
-Despite of relatively weaker performance in Western Australia, Australia's North West region sees the strongest upward trend, bypassing Sydney in some years.
+Given the `spec`, clicking either one element in any plot highlights all points that match the `Region` category, that is, categorical linking.
+Figure \ref{fig:tourism-linking-fig} is a static view of an interactive exploration.
+The steps in getting to this point were:
 
-In summary, shared tsibble data nicely bridges between the \CRANpkg{crosstalk} and **tidyverts** ecosystems for temporal data over their common term "key".
-The `as_shared_tsibble()` provides a symbolic user interface for effortless construction of a hybrid of hierarchical and categorical linkings.
-And the `plotly_key_tree()` in turn decodes the specification to plot a tree for data overview and navigation, accompanied with more detailed plots.
+1.
+A branch of the tree corresponding to Western Australia was first selected.
+(The names of the regions are a little odd, which is a quirk of the data set, but all four areas, Australia's South West, ...., correspond to tourist destinations in Western Australia.
+Hovering over the node on the branch brings up the state name.) This generated the response in the line plots and the scatterplot that colored corresponding time series and points as blue.
+2.
+To enable persistent selection, in oder to compare regions or states, "Shift" and click on the tree was done, after switching the colour to red.
+This generated the response that points and time series corresponding to Sydney were highlighted in red.
+3. Hovering over the points brings up the label for Sydney.
+
+<!-- In Figure \ref{fig:tourism-linking-fig}, when hovering and selecting the circle associated with "Sydney" in the scatter plot, all data records with shared values of "Sydney" listen and react to this interaction via self updating in red.
+In order for comparison with other regions or states, press the "Shift" key to enable persistent selection, and simultaneously select the parent node on the tree, saying "Western Australia", to include all the children by switching to the blue colour.-->
+The domestic tourism sees Sydney as one of the most popular destinations in realm of business and friends visiting over the years.
+Despite the relatively weaker performance in Western Australia, Australia's North West region sees a strongest upward trend in Business, bypassing Sydney in some years.
+
+In summary, shared the tsibble data nicely bridges between the \CRANpkg{crosstalk} and **tidyverts** ecosystems for temporal data using the common "key".
+The `as_shared_tsibble()` provides a symbolic user interface for effortless construction of a hybrid of hierarchical and categorical linking between plots.
+The `plotly_key_tree()` fuction, in turn, decodes the hierarchical specification to plot a tree for data overview and navigation, when accompanied by more detailed plots.
 

### 2020-09-23 tsibbletalk.tex
done with linking section

@@
 
-The tree plot provides the data skeleton, upon which the rest of data
-can be attached. In this example, small multiples of line plots are
-placed at the top right of Figure \ref{fig:tourism-linking-fig} to
-explore the temporal trend across regions, by trip purpose. The shared
-tsibble data can be directly piped into \CRANpkg{ggplot2} code as
-follows.
+The tree plot provides the graphics skeleton, upon which the rest of
+data plots can be attached. In this example, small multiples of line
+plots are placed at the top right of Figure
+\ref{fig:tourism-linking-fig} to explore the temporal trend across
+regions, by trip purpose. The shared tsibble data can be directly piped
+into \CRANpkg{ggplot2} code to create this.
 
@@
 
-To tease apart these overlaid time series, they are funnelled through
-the \texttt{features()} S3 method to extract some key characteristics,
-including the measurements of trend and seasonality. A scatterplot is
-populated from these statistics for each series.
+These line plots are heavily overplotted. To tease apart structure in
+the multiple time series, the \texttt{features()} function
+\texttt{fabletools} computes interesting characteristics, including the
+measures of trend and seasonality. These are displayed in the
+scatterplot at bottom right, where one dot represents one series.
 
@@
 
-Lastly, three graphics are composed as an ensemble of coordinated views
-for multi-facetted exploration, shown as Figure
-\ref{fig:tourism-linking-fig} (the interactive realisation of Figure
-\ref{fig:highlight-retail}). Routine functions bring about new
-interaction with temporal data on the client side.
+There is one final step, to compose the three plots into an ensemble of
+coordinated views for exploration, shown in Figure
+\ref{fig:tourism-linking-fig}. (This is the interactive realisation of
+Figure \ref{fig:highlight-retail}).
 
@@
 
-Since all plots are stemmed from one shared tsibble data source, they
+Since all plots are created from one shared tsibble data source, they
 are self-linking views. Nodes, lines, and points are hoverable and
 clickable. Given the \texttt{spec}, clicking either one element in any
-plot highlights all points that match the \texttt{Region} category,
-briefly ``categorical linking''. In Figure
-\ref{fig:tourism-linking-fig}, when hovering and selecting the circle
-associated with ``Sydney'' in the scatter plot, all data records with
-shared values of ``Sydney'' listen and react to this interaction via
-self updating in red. In order for comparison with other regions or
-states, press the ``Shift'' key to enable persistent selection, and
-simultaneously select the parent node on the tree, saying ``Western
-Australia'', to include all the children by switching to the blue
-colour. The domestic tourism sees Sydney as one of the most popular
-destinations in realm of business and friends visiting over years.
-Despite of relatively weaker performance in Western Australia,
-Australia's North West region sees the strongest upward trend, bypassing
-Sydney in some years.
+plot highlights all points that match the \texttt{Region} category, that
+is, categorical linking. Figure \ref{fig:tourism-linking-fig} is a
+static view of an interactive exploration. The steps in getting to this
+point were:
 
-In summary, shared tsibble data nicely bridges between the
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  A branch of the tree corresponding to Western Australia was first
+  selected. (The names of the regions are a little odd, which is a quirk
+  of the data set, but all four areas, Australia's South West, \ldots.,
+  correspond to tourist destinations in Western Australia. Hovering over
+  the node on the branch brings up the state name.) This generated the
+  response in the line plots and the scatterplot that colored
+  corresponding time series and points as blue.
+\item
+  To enable persistent selection, in oder to compare regions or states,
+  ``Shift'' and click on the tree was done, after switching the colour
+  to red. This generated the response that points and time series
+  corresponding to Sydney were highlighted in red.
+\item
+  Hovering over the points brings up the label for Sydney.
+\end{enumerate}
+
+The domestic tourism sees Sydney as one of the most popular destinations
+in realm of business and friends visiting over the years. Despite the
+relatively weaker performance in Western Australia, Australia's North
+West region sees a strongest upward trend in Business, bypassing Sydney
+in some years.
+
+In summary, shared the tsibble data nicely bridges between the
 \CRANpkg{crosstalk} and \textbf{tidyverts} ecosystems for temporal data
-over their common term ``key''. The \texttt{as\_shared\_tsibble()}
-provides a symbolic user interface for effortless construction of a
-hybrid of hierarchical and categorical linkings. And the
-\texttt{plotly\_key\_tree()} in turn decodes the specification to plot a
-tree for data overview and navigation, accompanied with more detailed
-plots.
+using the common ``key''. The \texttt{as\_shared\_tsibble()} provides a
+symbolic user interface for effortless construction of a hybrid of
+hierarchical and categorical linking between plots. The
+\texttt{plotly\_key\_tree()} fuction, in turn, decodes the hierarchical
+specification to plot a tree for data overview and navigation, when
+accompanied by more detailed plots.
 

### 2020-09-23 tsibbletalk.Rmd
caption for figure 2

@@
 
-```{r tourism-linking-fig, fig.cap = "Coordinated views generated from the \\code{tourism\\_shared} with persistent linked brushing."}
+```{r tourism-linking-fig, fig.cap = "Snapshot of exploring an ensemble of linked plots of the Australian tourism data, built on a \\code{tourism\\_shared} object.
+It also illustrates persistent linked brushing to compare two groups."}
 include_graphics("img/tourism-linking.png")

### 2020-09-23 tsibbletalk.tex
caption for figure 2

@@
 
-\caption[Coordinated views generated from the \code{tourism\_shared} with persistent linked brushing]{Coordinated views generated from the \code{tourism\_shared} with persistent linked brushing.}\label{fig:tourism-linking-fig}
+\caption[Snapshot of exploring an ensemble of linked plots of the Australian tourism data, built on a \code{tourism\_shared} object]{Snapshot of exploring an ensemble of linked plots of the Australian tourism data, built on a \code{tourism\_shared} object.
+It also illustrates persistent linked brushing to compare two groups.}\label{fig:tourism-linking-fig}
 \end{figure}

### 2020-09-24 tsibbletalk.Rmd
slicing

@@
 
-The shared tsibble data leverages the `key` attribute to converse with many coordinated views, with or without \CRANpkg{shiny}.
+An important aspect of temporal data is the time context.
+Time has cyclical structure, that may correspond to seasonal patterns to be discovered.
+The `index` component of the shared tsibble data forms the basis for exploring seasonality.
+To investigate for periodic or aperiodic patterns, series should be wrapped on themselves, where the index is broken into temporal components like quarter or day.
+We'll explore this with pedestrian traffic in Melbourne, Australia.
+
+<!-- The shared tsibble data leverages the `key` attribute to converse with many coordinated views, with or without \CRANpkg{shiny}.
 On the other hand, a second critical attribute--`index`--lays the foundational temporal context that augments the conversation.
 When temporal data are plotted and stretched against the entire span like Figure \ref{fig:highlight-retail-1}, it puts emphasis on the trend perception.
-Yet to digest periodic/aperiodic patterns, data should be wrapped over relative time units that are origin-less, such as one quarter or one day.
+Yet to digest periodic/aperiodic patterns, data should be wrapped over relative time units that are origin-less, such as one quarter or one day.-->
 
-```{r wrap-ped, fig.show = "hold", fig.cap = "The shiny module for wrapping the \\code{pedestrian20} data.", fig.subcap = c("Initial overview state", "1-day state", "7-day state, anchoring to Monday"), fig.ncol = 1}
+```{r wrap-ped, fig.show = "hold", fig.cap = "Snapshots wrapping after slicing the \\code{pedestrian20} data at different intervals, none (top), daily (middle) and weekly (bottom).
+This type of interaction is made possible with Shiny elements.", fig.subcap = c("Initial overview state", "1-day state", "7-day state, anchoring to Monday"), fig.ncol = 1}
 include_graphics("img/wrap-0.png")
@@
 
-The city of Melbourne has sensors installed to count hourly tallies of pedestrians in order to capture downtown daily rhythms [@ped].
-Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four locations, with the depiction of three pronounced slices in time.
-Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on their absolute time line, facetted by four sensors.
-On March 16, Melbourne went to the stage three lockdown due to COVID-19, seeing a significant decline in traffic volume across the city.
-These lines are then folded into daily and weekly sections, shown as Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} respectively.
-Seasonal variations have been popped out to viewers, complementing the not-just-magnitude-drop story.
-The pre-lockdown period is coloured with dark green and lockdown with orange.
+The city of Melbourne has sensors installed at various locations, to record hourly counts of pedestrians, in order to capture daily rhythms of the downtown [@ped].
+Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four different locations, for three different time slices, daily, weekly and full five months.
+Plot \ref{fig:wrap-ped-1} shows hourly counts from January to May on an absolute time line, facetted by location.
+The stage 3 COVID-19 lockdown, on March 16, is marked by a change of colour.
+(The pre-lockdown period is coloured with dark green and lockdown with orange.) We can see a significant decline in foot traffic  at all four locations.
+QV Market is less affected probably because this is a major produce market, an essential service that continued to operate.
+Bourke St, a major shopping center, sees a gradual uptick in the last weeks of the period indicating that people were getting back into the shops.
 
-The wrapping procedure involves slicing time indices into seasonal periods of interest and their corresponding time dices.
-For example, hourly pedestrian data can be decomposed into 24-hour blocks grouped by all respective days, like Figure \ref{fig:wrap-ped-2}.
-Figure \ref{fig:wrap-ped} suggests that there could be more than one eye-catching slices out of many possible combinations, and thus repeated wrappings can be unwieldy.
-To visually locate a salient slice, the \CRANpkg{tsibbletalk} package implements a shiny module, a pair of UI and server functions, to automate this wrapping procedure.
+Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} show slicing and wrapping of the series into daily and weekly sections, respectively.
+Multiple seasonality pops out.
+There tends to be a daily pattern, especially visible at the main train station, Southern Cross.
+There is also a week day vs weekend pattern, also most visible at Southern Cross.
+These seasonal patterns are still present during the lockdown, but the magnitude is greatly reduced.
+Numbers are also down at the produce market and the shopping center.
+Birrarung Marr is the most affected.
+This is the location of special events and its clear that these have completely disappeared during lockdown.
 
-This shiny module, decoupled to `tsibbleWrapUI()` and `tsibbleWrapServer()`, presents a clean interface and forms a resusable piece in a shiny application.
-A shiny module provides the vehicle in modularising shiny applications for both users and developers.
-Like all shiny modules, the first argument in both functions requires a user-supplied session id that must be unique.
+The wrapping procedure involves slicing the time index into seasonal periods of interest, and the result is diced time.
+For example, hourly pedestrian data can be decomposed into 24-hour blocks, which then overlays the counts for all respective days, as done in plot \ref{fig:wrap-ped-2}.
+For exploration, this slice position should be controlled interactively, so that many different slices can be examined rapidly.
+This can be achieved using shiny, with the functions provided in a \CRANpkg{tsibbletalk}.
+
+This shiny module, decoupled to `tsibbleWrapUI()` and `tsibbleWrapServer()`, presents a clean interface and forms a reusable code that could be embedded in any shiny application.
+In general, a shiny module provides a vehicle for modularising shiny applications, relevant for both users and developers.
+As with all shiny modules, the first argument in both functions in \CRANpkg{tsibbletalk} requires a user-supplied session id that must be unique.
 The UI function `tsibbleWrapUI()` simply shows a slider that animates or controls the number of periods to be diced.
-The workhorse is certainly the server function `tsibbleWrapServer()`, encapsulating the algorithm that transforms data and sends messages to update the plot accordingly.
+The workhorse is the server function `tsibbleWrapServer()`, encapsulating the algorithm that transforms data and sends messages to update the plot accordingly.
 The `plot` argument expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots).
-As the function name suggests, a (shared) tsibble is needed to start the engine, and thereby the time `index` can be retrieved for dissection.
+As the function name suggests, a (shared) tsibble is needed to start the engine, so that the time `index` can be retrieved for dissection.
 The `period` option semantically takes a desired number of seasonal periods to be shifted, for example data shifted by "1 day", "2 days", or "1 week", etc.
 In other words, the `period` defines the grind level.
-For date-times (represented by `POSIXt`), the granularity ranges from fine "day" to coarse "year".
+For date-times (represented by `POSIXt`), the granularity ranges from fine "day" to a much coarser "year".
 The following code snippet generates Figure \ref{fig:wrap-ped}.
@@
 
-Upon running the shiny application, Figure \ref{fig:wrap-ped-1} corresponds to the initial state, with the slider incremented by 1-day unit.
-The "play" button near the end of slider begins animating the slicing and dicing process by walking through all 24 hours by 152 days.
-Alternatively, users can drag the handler to poke around certain slices themselves.
+Figure \ref{fig:wrap-ped-1} corresponds to the initial state, with the slider incremented by 1-day units.
+The "play" button near the end of slider can automately animate the slicing and dicing process, walking the viewer through all 24 hours of the 152 days.
+Alternatively, users can drag the slider to examine selected slices.
 
@@
 The underlying tsibble data is being called back and processed in R.
-Only transformed data gets fed back to the shiny server, along with reseting the x-axis ranges and breaks.
-The rest plot configurations, such as marks, y-axes, and layouts, are properly cached.
+Only transformed data gets fed back to the shiny server, for updating with resetting the x-axis ranges and breaks.
+The other plot configurations, such as marks, y-axes, and layouts, are cached and used as is.
 
-The new shiny module exploits the temporal aspect for a tsibble object, exposed by the `index` attribute.
-It allows users to slide through relative periods to digest seasonal behaviours, with slick user experience.
+The new shiny module exploits the temporal aspect for a tsibble object, available throu the `index` attribute.
+It allows users to slide through relative periods to digest seasonal behaviours, with nimble user experience.
 

### 2020-09-24 tsibbletalk.tex
slicing

@@
 
-The shared tsibble data leverages the \texttt{key} attribute to converse
-with many coordinated views, with or without \CRANpkg{shiny}. On the
-other hand, a second critical attribute--\texttt{index}--lays the
-foundational temporal context that augments the conversation. When
-temporal data are plotted and stretched against the entire span like
-Figure \ref{fig:highlight-retail-1}, it puts emphasis on the trend
-perception. Yet to digest periodic/aperiodic patterns, data should be
-wrapped over relative time units that are origin-less, such as one
-quarter or one day.
+An important aspect of temporal data is the time context. Time has
+cyclical structure, that may correspond to seasonal patterns to be
+discovered. The \texttt{index} component of the shared tsibble data
+forms the basis for exploring seasonality. To investigate for periodic
+or aperiodic patterns, series should be wrapped on themselves, where the
+index is broken into temporal components like quarter or day. We'll
+explore this with pedestrian traffic in Melbourne, Australia.
 
@@
 
-\caption[The shiny module for wrapping the \code{pedestrian20} data]{The shiny module for wrapping the \code{pedestrian20} data.}\label{fig:wrap-ped}
+\caption[Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, none (top), daily (middle) and weekly (bottom)]{Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, none (top), daily (middle) and weekly (bottom).
+This type of interaction is made possible with Shiny elements.}\label{fig:wrap-ped}
 \end{figure}
@@
 
-The city of Melbourne has sensors installed to count hourly tallies of
-pedestrians in order to capture downtown daily rhythms \citep{ped}.
-Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot
-traffic at four locations, with the depiction of three pronounced slices
-in time. Figure \ref{fig:wrap-ped-1} unfolds all counts from January to
-May on their absolute time line, facetted by four sensors. On March 16,
-Melbourne went to the stage three lockdown due to COVID-19, seeing a
-significant decline in traffic volume across the city. These lines are
-then folded into daily and weekly sections, shown as Figure
-\ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} respectively. Seasonal
-variations have been popped out to viewers, complementing the
-not-just-magnitude-drop story. The pre-lockdown period is coloured with
-dark green and lockdown with orange.
+The city of Melbourne has sensors installed at various locations, to
+record hourly counts of pedestrians, in order to capture daily rhythms
+of the downtown \citep{ped}. Figure \ref{fig:wrap-ped} shows the first
+five months of 2020 foot traffic at four different locations, for three
+different time slices, daily, weekly and full five months. Plot
+\ref{fig:wrap-ped-1} shows hourly counts from January to May on an
+absolute time line, facetted by location. The stage 3 COVID-19 lockdown,
+on March 16, is marked by a change of colour. (The pre-lockdown period
+is coloured with dark green and lockdown with orange.) We can see a
+significant decline in foot traffic at all four locations. QV Market is
+less affected probably because this is a major produce market, an
+essential service that continued to operate. Bourke St, a major shopping
+center, sees a gradual uptick in the last weeks of the period indicating
+that people were getting back into the shops.
 
-The wrapping procedure involves slicing time indices into seasonal
-periods of interest and their corresponding time dices. For example,
-hourly pedestrian data can be decomposed into 24-hour blocks grouped by
-all respective days, like Figure \ref{fig:wrap-ped-2}. Figure
-\ref{fig:wrap-ped} suggests that there could be more than one
-eye-catching slices out of many possible combinations, and thus repeated
-wrappings can be unwieldy. To visually locate a salient slice, the
-\CRANpkg{tsibbletalk} package implements a shiny module, a pair of UI
-and server functions, to automate this wrapping procedure.
+Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} show slicing and
+wrapping of the series into daily and weekly sections, respectively.
+Multiple seasonality pops out. There tends to be a daily pattern,
+especially visible at the main train station, Southern Cross. There is
+also a week day vs weekend pattern, also most visible at Southern Cross.
+These seasonal patterns are still present during the lockdown, but the
+magnitude is greatly reduced. Numbers are also down at the produce
+market and the shopping center. Birrarung Marr is the most affected.
+This is the location of special events and its clear that these have
+completely disappeared during lockdown.
+
+The wrapping procedure involves slicing the time index into seasonal
+periods of interest, and the result is diced time. For example, hourly
+pedestrian data can be decomposed into 24-hour blocks, which then
+overlays the counts for all respective days, as done in plot
+\ref{fig:wrap-ped-2}. For exploration, this slice position should be
+controlled interactively, so that many different slices can be examined
+rapidly. This can be achieved using shiny, with the functions provided
+in a \CRANpkg{tsibbletalk}.
 
@@
 \texttt{tsibbleWrapServer()}, presents a clean interface and forms a
-resusable piece in a shiny application. A shiny module provides the
-vehicle in modularising shiny applications for both users and
-developers. Like all shiny modules, the first argument in both functions
+reusable code that could be embedded in any shiny application. In
+general, a shiny module provides a vehicle for modularising shiny
+applications, relevant for both users and developers. As with all shiny
+modules, the first argument in both functions in \CRANpkg{tsibbletalk}
 requires a user-supplied session id that must be unique. The UI function
 \texttt{tsibbleWrapUI()} simply shows a slider that animates or controls
-the number of periods to be diced. The workhorse is certainly the server
-function \texttt{tsibbleWrapServer()}, encapsulating the algorithm that
+the number of periods to be diced. The workhorse is the server function
+\texttt{tsibbleWrapServer()}, encapsulating the algorithm that
 transforms data and sends messages to update the plot accordingly. The
@@
 elements (such as boxplots). As the function name suggests, a (shared)
-tsibble is needed to start the engine, and thereby the time
-\texttt{index} can be retrieved for dissection. The \texttt{period}
-option semantically takes a desired number of seasonal periods to be
-shifted, for example data shifted by ``1 day'', ``2 days'', or ``1
-week'', etc. In other words, the \texttt{period} defines the grind
-level. For date-times (represented by \texttt{POSIXt}), the granularity
-ranges from fine ``day'' to coarse ``year''. The following code snippet
-generates Figure \ref{fig:wrap-ped}.
+tsibble is needed to start the engine, so that the time \texttt{index}
+can be retrieved for dissection. The \texttt{period} option semantically
+takes a desired number of seasonal periods to be shifted, for example
+data shifted by ``1 day'', ``2 days'', or ``1 week'', etc. In other
+words, the \texttt{period} defines the grind level. For date-times
+(represented by \texttt{POSIXt}), the granularity ranges from fine
+``day'' to a much coarser ``year''. The following code snippet generates
+Figure \ref{fig:wrap-ped}.
 
@@
 
-Upon running the shiny application, Figure \ref{fig:wrap-ped-1}
-corresponds to the initial state, with the slider incremented by 1-day
-unit. The ``play'' button near the end of slider begins animating the
-slicing and dicing process by walking through all 24 hours by 152 days.
-Alternatively, users can drag the handler to poke around certain slices
-themselves.
+Figure \ref{fig:wrap-ped-1} corresponds to the initial state, with the
+slider incremented by 1-day units. The ``play'' button near the end of
+slider can automately animate the slicing and dicing process, walking
+the viewer through all 24 hours of the 152 days. Alternatively, users
+can drag the slider to examine selected slices.
 
@@
 and processed in R. Only transformed data gets fed back to the shiny
-server, along with reseting the x-axis ranges and breaks. The rest plot
-configurations, such as marks, y-axes, and layouts, are properly cached.
+server, for updating with resetting the x-axis ranges and breaks. The
+other plot configurations, such as marks, y-axes, and layouts, are
+cached and used as is.
 
 The new shiny module exploits the temporal aspect for a tsibble object,
-exposed by the \texttt{index} attribute. It allows users to slide
-through relative periods to digest seasonal behaviours, with slick user
+available throu the \texttt{index} attribute. It allows users to slide
+through relative periods to digest seasonal behaviours, with nimble user
 experience.

### 2020-09-24 tsibbletalk.Rmd
done proofing and checking

@@
 
-Close to the heart of the \CRANpkg{tsibbletalk} package is blending the best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly}, and \CRANpkg{shiny}.
+At the heart of the \CRANpkg{tsibbletalk} package is a blending the best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly}, and \CRANpkg{shiny}.
 
@@
 If nesting is found in the data, the `plotly_key_tree()` creates an interactive hierarchical tree to help with data overview.
-This sets the stage for hierarchical and categorical linking between multi-views from one shared tsibble.
+This sets the stage for hierarchical and categorical linking between multiple views from one shared tsibble.
 
-A new shiny module, `tsibbleWrapUI()` and `tsibbleWrapServer()`, provides a lens for looking at temporal aspects for a `tsibble`, in particular seasonal or cyclical variations.
+A new shiny module, `tsibbleWrapUI()` and `tsibbleWrapServer()`, provides a lens for looking at temporal aspects of a `tsibble`, in particular seasonal or cyclical variations.
 The slicing and dicing technique efficiently wrap time lines for user-defined plots.

### 2020-09-24 tsibbletalk.tex
done proofing and checking

@@
 
-Close to the heart of the \CRANpkg{tsibbletalk} package is blending the
-best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly},
-and \CRANpkg{shiny}.
+At the heart of the \CRANpkg{tsibbletalk} package is a blending the best
+bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly}, and
+\CRANpkg{shiny}.
 
@@
 hierarchical tree to help with data overview. This sets the stage for
-hierarchical and categorical linking between multi-views from one shared
-tsibble.
+hierarchical and categorical linking between multiple views from one
+shared tsibble.
 
@@
 \texttt{tsibbleWrapServer()}, provides a lens for looking at temporal
-aspects for a \texttt{tsibble}, in particular seasonal or cyclical
+aspects of a \texttt{tsibble}, in particular seasonal or cyclical
 variations. The slicing and dicing technique efficiently wrap time lines

### 2020-09-24 tsibbletalk.Rmd
typos

@@
 Nesting of categorical variables produces a hierarchical structure.
-The \pkg{tsibbletalk} package is developed to allow a user to interactively explore temporal data, relative to the nested or crossed structures.
+The \CRANpkg{tsibbletalk} package is developed to allow a user to interactively explore temporal data, relative to the nested or crossed structures.
 It can help to discover differences between category levels, and uncover interesting periodic or aperiodic slices.
-The package implements a shared `tsibble` object that allows for
-  linked brushing between coordinated views, and a shiny module that aids in wrapping
-  time lines for seasonal patterns. The tools are demonstrated using two data examples: retail turnover and tourism in Australia.
+The package implements a shared `tsibble` object that allows for linked brushing between coordinated views, and a shiny module that aids in wrapping time lines for seasonal patterns.
+The tools are demonstrated using two data examples: domestic tourism in Australia and foot traffic in Melbourne.
 preamble: |
@@
 The next section provides some background to the approach based on setting up data structures and workflow.
-The section following explains the mechanism for constructing interactivity, to link between multiple hiearchical data objects and hence plots.
+The section following explains the mechanism for constructing interactivity, to link between multiple hierarchical data objects and hence plots.
 The section after this describes the set up for interactively slicing and dicing time to wrap a series on itself to investigate periodicities.
@@
 Figure \ref{fig:highlight-retail} illustrates examining temporal data with many keys.
-The data has 152 series coorresponding to different items in retail data.
+The data has 152 series corresponding to different items in retail data.
 The multiple series are displayed using an overlaid time series plot, along with a scatterplot of two variables (trend versus seasonal strength) from feature space, where each series is represented by a dot.
@@
 It illustrates a need to query interesting series on the fly.
-Although these two plots are static, we can consider them as linked views becaause the common `key` variables link between the two data tables producing the two ploots.
+Although these two plots are static, we can consider them as linked views because the common `key` variables link between the two data tables producing the two plots.
 This motivates the work in this package, described in this paper, to enable interactivity of `tsibble` and `tsibble`-derived objects for rapid exploratory data analysis.
@@
 These line plots are heavily overplotted.
-To tease apart structure in the multiple time series, the `features()` function `fabletools` computes interesting characteristics, including the measures of trend and seasonality.
+To tease apart structure in the multiple time series, the `features()` function computes interesting characteristics, including the measures of trend and seasonality.
 These are displayed in the scatterplot at bottom right, where one dot represents one series.
@@
 (The names of the regions are a little odd, which is a quirk of the data set, but all four areas, Australia's South West, ...., correspond to tourist destinations in Western Australia.
-Hovering over the node on the branch brings up the state name.) This generated the response in the line plots and the scatterplot that colored corresponding time series and points as blue.
+Hovering over the node on the branch brings up the state name.) This generated the response in the line plots and the scatterplot that coloured corresponding time series and points as blue.
 2.
@@
 The `as_shared_tsibble()` provides a symbolic user interface for effortless construction of a hybrid of hierarchical and categorical linking between plots.
-The `plotly_key_tree()` fuction, in turn, decodes the hierarchical specification to plot a tree for data overview and navigation, when accompanied by more detailed plots.
+The `plotly_key_tree()` function, in turn, decodes the hierarchical specification to plot a tree for data overview and navigation, when accompanied by more detailed plots.
 
@@
 To investigate for periodic or aperiodic patterns, series should be wrapped on themselves, where the index is broken into temporal components like quarter or day.
-We'll explore this with pedestrian traffic in Melbourne, Australia.
+We shall explore this with pedestrian traffic in Melbourne, Australia.
 
@@
 Figure \ref{fig:wrap-ped-1} corresponds to the initial state, with the slider incremented by 1-day units.
-The "play" button near the end of slider can automately animate the slicing and dicing process, walking the viewer through all 24 hours of the 152 days.
+The "play" button near the end of slider can automatically animate the slicing and dicing process, walking the viewer through all 24 hours of the 152 days.
 Alternatively, users can drag the slider to examine selected slices.
@@
 
-The new shiny module exploits the temporal aspect for a tsibble object, available throu the `index` attribute.
+The new shiny module exploits the temporal aspect for a tsibble object, available through the `index` attribute.
 It allows users to slide through relative periods to digest seasonal behaviours, with nimble user experience.

### 2020-09-24 tsibbletalk.tex
typos

@@
 sales products. Nesting of categorical variables produces a hierarchical
-structure. The \pkg{tsibbletalk} package is developed to allow a user to
-interactively explore temporal data, relative to the nested or crossed
-structures. It can help to discover differences between category levels,
-and uncover interesting periodic or aperiodic slices. The package
-implements a shared \texttt{tsibble} object that allows for linked
-brushing between coordinated views, and a shiny module that aids in
-wrapping time lines for seasonal patterns. The tools are demonstrated
-using two data examples: retail turnover and tourism in Australia.
+structure. The \CRANpkg{tsibbletalk} package is developed to allow a
+user to interactively explore temporal data, relative to the nested or
+crossed structures. It can help to discover differences between category
+levels, and uncover interesting periodic or aperiodic slices. The
+package implements a shared \texttt{tsibble} object that allows for
+linked brushing between coordinated views, and a shiny module that aids
+in wrapping time lines for seasonal patterns. The tools are demonstrated
+using two data examples: domestic tourism in Australia and foot traffic
+in Melbourne.
 }
@@
 mechanism for constructing interactivity, to link between multiple
-hiearchical data objects and hence plots. The section after this
+hierarchical data objects and hence plots. The section after this
 describes the set up for interactively slicing and dicing time to wrap a
@@
 Figure \ref{fig:highlight-retail} illustrates examining temporal data
-with many keys. The data has 152 series coorresponding to different
-items in retail data. The multiple series are displayed using an
-overlaid time series plot, along with a scatterplot of two variables
-(trend versus seasonal strength) from feature space, where each series
-is represented by a dot. The feature space is computed using the
-\texttt{features()} function from \CRANpkg{feasts}, which summarises the
-original data for each series using various statistical features. This
-function along with other \textbf{tidyverts} functions is
-\texttt{tsibble}-aware, and outputs a table in a reduced form where each
-row corresponds to a series, which can be graphically displayed as in
-Figure \ref{fig:highlight-retail-2}.
+with many keys. The data has 152 series corresponding to different items
+in retail data. The multiple series are displayed using an overlaid time
+series plot, along with a scatterplot of two variables (trend versus
+seasonal strength) from feature space, where each series is represented
+by a dot. The feature space is computed using the \texttt{features()}
+function from \CRANpkg{feasts}, which summarises the original data for
+each series using various statistical features. This function along with
+other \textbf{tidyverts} functions is \texttt{tsibble}-aware, and
+outputs a table in a reduced form where each row corresponds to a
+series, which can be graphically displayed as in Figure
+\ref{fig:highlight-retail-2}.
 
@@
 query interesting series on the fly. Although these two plots are
-static, we can consider them as linked views becaause the common
+static, we can consider them as linked views because the common
 \texttt{key} variables link between the two data tables producing the
-two ploots. This motivates the work in this package, described in this
+two plots. This motivates the work in this package, described in this
 paper, to enable interactivity of \texttt{tsibble} and
@@
 These line plots are heavily overplotted. To tease apart structure in
-the multiple time series, the \texttt{features()} function
-\texttt{fabletools} computes interesting characteristics, including the
-measures of trend and seasonality. These are displayed in the
-scatterplot at bottom right, where one dot represents one series.
+the multiple time series, the \texttt{features()} function computes
+interesting characteristics, including the measures of trend and
+seasonality. These are displayed in the scatterplot at bottom right,
+where one dot represents one series.
 
@@
   the node on the branch brings up the state name.) This generated the
-  response in the line plots and the scatterplot that colored
+  response in the line plots and the scatterplot that coloured
   corresponding time series and points as blue.
@@
 hierarchical and categorical linking between plots. The
-\texttt{plotly\_key\_tree()} fuction, in turn, decodes the hierarchical
+\texttt{plotly\_key\_tree()} function, in turn, decodes the hierarchical
 specification to plot a tree for data overview and navigation, when
@@
 or aperiodic patterns, series should be wrapped on themselves, where the
-index is broken into temporal components like quarter or day. We'll
+index is broken into temporal components like quarter or day. We shall
 explore this with pedestrian traffic in Melbourne, Australia.
@@
 slider incremented by 1-day units. The ``play'' button near the end of
-slider can automately animate the slicing and dicing process, walking
+slider can automatically animate the slicing and dicing process, walking
 the viewer through all 24 hours of the 152 days. Alternatively, users
@@
 The new shiny module exploits the temporal aspect for a tsibble object,
-available throu the \texttt{index} attribute. It allows users to slide
+available through the \texttt{index} attribute. It allows users to slide
 through relative periods to digest seasonal behaviours, with nimble user

### 2020-09-25 tsibbletalk.Rmd
minor corrections

@@
 The package implements a shared `tsibble` object that allows for linked brushing between coordinated views, and a shiny module that aids in wrapping time lines for seasonal patterns.
-The tools are demonstrated using two data examples: domestic tourism in Australia and foot traffic in Melbourne.
+The tools are demonstrated using two data examples: domestic tourism in Australia and pedestrian traffic in Melbourne.
 preamble: |
@@
 Temporal data typically arrives as a set of many observational units measured over time.
-Some variables may be categorical, indicating a hierarchy in the collection process, that may be measurements taken in different geographic regions, or types of products sold by the one company.
+Some variables may be categorical, containing a hierarchy in the collection process, that may be measurements taken in different geographic regions, or types of products sold by the one company.
 Exploring these multiple features can be daunting.
@@
 
-This article describes new software, the \CRANpkg{tsibbletalk} package, for exploring temporal data using linked views and time wrapping.
-The next section provides some background to the approach based on setting up data structures and workflow.
-The section following explains the mechanism for constructing interactivity, to link between multiple hierarchical data objects and hence plots.
-The section after this describes the set up for interactively slicing and dicing time to wrap a series on itself to investigate periodicities.
+This paper describes new software, the \CRANpkg{tsibbletalk} package, for exploring temporal data using linked views and time wrapping.
+We first provide some background to the approach based on setting up data structures and workflow, and give an overview of interactive systems in R.
+The section following introduces the \CRANpkg{tsibbletalk} package.
+We explain the mechanism for constructing interactivity, to link between multiple hierarchical data objects and hence plots, and describe the set up for interactively slicing and dicing time to wrap a series on itself to investigate periodicities.
 
@@
 The \CRANpkg{tsibble} package [@wang2020tsibble] introduced a unified temporal data structure, referred to as a `tsibble`, to represent time series and longitudinal data in a tidy format [@wickham2014tidy].
-A `tsibble` extends the `data.frame` and \CRANpkg{tibble} classes with the temporal contextual metadata: `index` and `key`.
+A `tsibble` extends the `data.frame` and `tibble` classes with the temporal contextual metadata: `index` and `key`.
 The `index` declares a data column that holds time-related indices.
@@
 The `Month` column holds year-months as the `index`.
-`State` and `Industry`, together, are the identifiers for these 152 series, which form the `key`.
+`State` and `Industry` are the identifiers for these 152 series, which form the `key`.
 Note that the column `Series ID` could be an alternative option for setting up the `key`, but `State` and `Industry` are more readable and informative.
@@
 Figure \ref{fig:highlight-retail} illustrates examining temporal data with many keys.
-The data has 152 series corresponding to different items in retail data.
+The data has 152 series corresponding to different industries in retail data.
 The multiple series are displayed using an overlaid time series plot, along with a scatterplot of two variables (trend versus seasonal strength) from feature space, where each series is represented by a dot.
@@
 
-R \CRANpkg{shiny} [@R-shiny] and \CRANpkg{htmlwidgets} [@R-htmlwidgets] provides and infrastructure connecting R with HTML elements and JavaScript that provide the interactivity.<!--, although they seldom emphasise on interactive graphics.--> The \CRANpkg{htmlwidgets} package makes it possible to embed JavaScript libraries into R so that users are able to write only R code to generate web-based plots.
+R \CRANpkg{shiny} [@R-shiny] and \CRANpkg{htmlwidgets} [@R-htmlwidgets] provide infrastructure connecting R with HTML elements and JavaScript that support the interactivity.
+The \CRANpkg{htmlwidgets} package makes it possible to embed JavaScript libraries into R so that users are able to write only R code to generate web-based plots.
 Many JavaScript charting libraries have been ported to R as HTML widgets, including \CRANpkg{plotly} [@plotly2020], \CRANpkg{rbokeh} [@R-rbokeh], and \CRANpkg{leaflet} [@R-leaflet] for maps.
@@
 This allows for seamless communication between different plots of temporal data.
-The `as_shared_tsibble()` function turns a `tsibble` into a shared instance,  `SharedTsibbleData`, which is a subclass of `SharedData` from \CRANpkg{crosstalk}).
+The `as_shared_tsibble()` function turns a `tsibble` into a shared instance,  `SharedTsibbleData`, which is a subclass of `SharedData` from \CRANpkg{crosstalk}.
 This is the object driving data transmission across multiple views.
@@
 As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical variable linking where selecting one or more observations in one category will broadcast to all other observations in this category.
-In addition, linking is also by time: within the time series plot, click on any data point, and  whole line will be highlighted in response.
+That is, linking is by time: within the time series plot, click on any data point, and  whole line will be highlighted in response.
 The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking.
 
-The approach can also accommodate hierarchical temporal data.
+The approach can also accommodate temporal data of nesting and crossing structures.
+These time series are referred to as hierarchical and grouped time series in the literature [@fpp].
+The `aus_retail` above is an example of grouped time series.
+Each series in the  data corresponds to all possible combinations of the `State` and `Industry` variables, which means they are intrinsically crossed with each other.
 When one key variable is nested within another, such as regional areas within a state, this is considered to be a hierarchical structure.
-These time series are referred to as hierarchical and grouped time series in the literature [@fpp].
-The `aus_retail` above is an example.
-Each series in the  data corresponds to all possible combinations of the `State` and `Industry` variables, which means they are intrinsically crossed with each other.
 
-The `spec` argument in `as_shared_tsibble()`  constructing hybrid linking, that incorporates hierarchical and categorical linking.
-A symbolic formula can be passed to the `spec` argument, to incorporate structured specifications in the `key`.
+The `spec` argument in `as_shared_tsibble()` provides a means to construct hybrid linking, that incorporates hierarchical and categorical linking.
+A symbolic formula can be passed to the `spec` argument, to allow for structured specifications in the `key`.
 Adopting @Wilkinson1973's notation for factorial models, the `spec` follows the `/` and `*` operator conventions to declare nesting and crossing variables, respectively.
@@
 
-In summary, shared the tsibble data nicely bridges between the \CRANpkg{crosstalk} and **tidyverts** ecosystems for temporal data using the common "key".
+In summary, shared tsibble data nicely bridges between the \CRANpkg{crosstalk} and **tidyverts** ecosystems for temporal data using the common "key".
 The `as_shared_tsibble()` provides a symbolic user interface for effortless construction of a hybrid of hierarchical and categorical linking between plots.
@@
 Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} show slicing and wrapping of the series into daily and weekly sections, respectively.
-Multiple seasonality pops out.
-There tends to be a daily pattern, especially visible at the main train station, Southern Cross.
-There is also a week day vs weekend pattern, also most visible at Southern Cross.
+Multiple seasonalities pop out.
+There tends to be a daily pattern, especially visible at the main train station, Southern Cross Station.
+There is also a weekday vs weekend pattern, also most visible at Southern Cross Station.
 These seasonal patterns are still present during the lockdown, but the magnitude is greatly reduced.
@@
 Birrarung Marr is the most affected.
-This is the location of special events and its clear that these have completely disappeared during lockdown.
+This is the location of special events, and it is clear that these have completely disappeared during lockdown.
 
@@
 For exploration, this slice position should be controlled interactively, so that many different slices can be examined rapidly.
-This can be achieved using shiny, with the functions provided in a \CRANpkg{tsibbletalk}.
+This can be achieved using shiny, with the functions provided in the \CRANpkg{tsibbletalk}.
 
-This shiny module, decoupled to `tsibbleWrapUI()` and `tsibbleWrapServer()`, presents a clean interface and forms a reusable code that could be embedded in any shiny application.
+This shiny module, decoupled to `tsibbleWrapUI()` and `tsibbleWrapServer()`, presents a clean interface and forms a reusable component that could be embedded in any shiny application.
 In general, a shiny module provides a vehicle for modularising shiny applications, relevant for both users and developers.
@@
 
-At the heart of the \CRANpkg{tsibbletalk} package is a blending the best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly}, and \CRANpkg{shiny}.
+At the heart of the \CRANpkg{tsibbletalk} package is a blending of the best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly}, and \CRANpkg{shiny}.
 

### 2020-09-25 tsibbletalk.tex
minor corrections

@@
 in wrapping time lines for seasonal patterns. The tools are demonstrated
-using two data examples: domestic tourism in Australia and foot traffic
-in Melbourne.
+using two data examples: domestic tourism in Australia and pedestrian
+traffic in Melbourne.
 }
@@
 Temporal data typically arrives as a set of many observational units
-measured over time. Some variables may be categorical, indicating a
+measured over time. Some variables may be categorical, containing a
 hierarchy in the collection process, that may be measurements taken in
@@
 
-This article describes new software, the \CRANpkg{tsibbletalk} package,
-for exploring temporal data using linked views and time wrapping. The
-next section provides some background to the approach based on setting
-up data structures and workflow. The section following explains the
-mechanism for constructing interactivity, to link between multiple
-hierarchical data objects and hence plots. The section after this
-describes the set up for interactively slicing and dicing time to wrap a
-series on itself to investigate periodicities.
+This paper describes new software, the \CRANpkg{tsibbletalk} package,
+for exploring temporal data using linked views and time wrapping. We
+first provide some background to the approach based on setting up data
+structures and workflow, and give an overview of interactive systems in
+R. The section following introduces the \CRANpkg{tsibbletalk} package.
+We explain the mechanism for constructing interactivity, to link between
+multiple hierarchical data objects and hence plots, and describe the set
+up for interactively slicing and dicing time to wrap a series on itself
+to investigate periodicities.
 
@@
 \citep{wickham2014tidy}. A \texttt{tsibble} extends the
-\texttt{data.frame} and \CRANpkg{tibble} classes with the temporal
+\texttt{data.frame} and \texttt{tibble} classes with the temporal
 contextual metadata: \texttt{index} and \texttt{key}. The \texttt{index}
@@
 \texttt{Month} column holds year-months as the \texttt{index}.
-\texttt{State} and \texttt{Industry}, together, are the identifiers for
-these 152 series, which form the \texttt{key}. Note that the column
+\texttt{State} and \texttt{Industry} are the identifiers for these 152
+series, which form the \texttt{key}. Note that the column
 \texttt{Series\ ID} could be an alternative option for setting up the
@@
 Figure \ref{fig:highlight-retail} illustrates examining temporal data
-with many keys. The data has 152 series corresponding to different items
-in retail data. The multiple series are displayed using an overlaid time
-series plot, along with a scatterplot of two variables (trend versus
-seasonal strength) from feature space, where each series is represented
-by a dot. The feature space is computed using the \texttt{features()}
-function from \CRANpkg{feasts}, which summarises the original data for
-each series using various statistical features. This function along with
-other \textbf{tidyverts} functions is \texttt{tsibble}-aware, and
-outputs a table in a reduced form where each row corresponds to a
-series, which can be graphically displayed as in Figure
-\ref{fig:highlight-retail-2}.
+with many keys. The data has 152 series corresponding to different
+industries in retail data. The multiple series are displayed using an
+overlaid time series plot, along with a scatterplot of two variables
+(trend versus seasonal strength) from feature space, where each series
+is represented by a dot. The feature space is computed using the
+\texttt{features()} function from \CRANpkg{feasts}, which summarises the
+original data for each series using various statistical features. This
+function along with other \textbf{tidyverts} functions is
+\texttt{tsibble}-aware, and outputs a table in a reduced form where each
+row corresponds to a series, which can be graphically displayed as in
+Figure \ref{fig:highlight-retail-2}.
 
@@
 R \CRANpkg{shiny} \citep{R-shiny} and \CRANpkg{htmlwidgets}
-\citep{R-htmlwidgets} provides and infrastructure connecting R with HTML
-elements and JavaScript that provide the interactivity. The
+\citep{R-htmlwidgets} provide infrastructure connecting R with HTML
+elements and JavaScript that support the interactivity. The
 \CRANpkg{htmlwidgets} package makes it possible to embed JavaScript
@@
 shared instance, \texttt{SharedTsibbleData}, which is a subclass of
-\texttt{SharedData} from \CRANpkg{crosstalk}). This is the object
-driving data transmission across multiple views. The
-\CRANpkg{tsibbletalk} package aims to streamline interactive exploration
-of temporal data, with the focus of temporal elements and structured
-linking.
+\texttt{SharedData} from \CRANpkg{crosstalk}. This is the object driving
+data transmission across multiple views. The \CRANpkg{tsibbletalk}
+package aims to streamline interactive exploration of temporal data,
+with the focus of temporal elements and structured linking.
 
@@
 one category will broadcast to all other observations in this category.
-In addition, linking is also by time: within the time series plot, click
-on any data point, and whole line will be highlighted in response. The
+That is, linking is by time: within the time series plot, click on any
+data point, and whole line will be highlighted in response. The
 \texttt{as\_shared\_tsibble()} uses \texttt{tsibble}'s \texttt{key}
@@
 
-The approach can also accommodate hierarchical temporal data. When one
-key variable is nested within another, such as regional areas within a
-state, this is considered to be a hierarchical structure. These time
-series are referred to as hierarchical and grouped time series in the
-literature \citep{fpp}. The \texttt{aus\_retail} above is an example.
-Each series in the data corresponds to all possible combinations of the
+The approach can also accommodate temporal data of nesting and crossing
+structures. These time series are referred to as hierarchical and
+grouped time series in the literature \citep{fpp}. The
+\texttt{aus\_retail} above is an example of grouped time series. Each
+series in the data corresponds to all possible combinations of the
 \texttt{State} and \texttt{Industry} variables, which means they are
-intrinsically crossed with each other.
+intrinsically crossed with each other. When one key variable is nested
+within another, such as regional areas within a state, this is
+considered to be a hierarchical structure.
 
-The \texttt{spec} argument in \texttt{as\_shared\_tsibble()}
-constructing hybrid linking, that incorporates hierarchical and
+The \texttt{spec} argument in \texttt{as\_shared\_tsibble()} provides a
+means to construct hybrid linking, that incorporates hierarchical and
 categorical linking. A symbolic formula can be passed to the
-\texttt{spec} argument, to incorporate structured specifications in the
+\texttt{spec} argument, to allow for structured specifications in the
 \texttt{key}. Adopting \citet{Wilkinson1973}'s notation for factorial
@@
 
-In summary, shared the tsibble data nicely bridges between the
+In summary, shared tsibble data nicely bridges between the
 \CRANpkg{crosstalk} and \textbf{tidyverts} ecosystems for temporal data
@@
 wrapping of the series into daily and weekly sections, respectively.
-Multiple seasonality pops out. There tends to be a daily pattern,
-especially visible at the main train station, Southern Cross. There is
-also a week day vs weekend pattern, also most visible at Southern Cross.
-These seasonal patterns are still present during the lockdown, but the
-magnitude is greatly reduced. Numbers are also down at the produce
-market and the shopping center. Birrarung Marr is the most affected.
-This is the location of special events and its clear that these have
-completely disappeared during lockdown.
+Multiple seasonalities pop out. There tends to be a daily pattern,
+especially visible at the main train station, Southern Cross Station.
+There is also a weekday vs weekend pattern, also most visible at
+Southern Cross Station. These seasonal patterns are still present during
+the lockdown, but the magnitude is greatly reduced. Numbers are also
+down at the produce market and the shopping center. Birrarung Marr is
+the most affected. This is the location of special events, and it is
+clear that these have completely disappeared during lockdown.
 
@@
 rapidly. This can be achieved using shiny, with the functions provided
-in a \CRANpkg{tsibbletalk}.
+in the \CRANpkg{tsibbletalk}.
 
@@
 \texttt{tsibbleWrapServer()}, presents a clean interface and forms a
-reusable code that could be embedded in any shiny application. In
+reusable component that could be embedded in any shiny application. In
 general, a shiny module provides a vehicle for modularising shiny
@@
 
-At the heart of the \CRANpkg{tsibbletalk} package is a blending the best
-bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly}, and
-\CRANpkg{shiny}.
+At the heart of the \CRANpkg{tsibbletalk} package is a blending of the
+best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly},
+and \CRANpkg{shiny}.
 

### 2020-09-26 tsibbletalk.Rmd
minor tweaks

@@
 
-<!-- In Figure \ref{fig:tourism-linking-fig}, when hovering and selecting the circle associated with "Sydney" in the scatter plot, all data records with shared values of "Sydney" listen and react to this interaction via self updating in red.
-In order for comparison with other regions or states, press the "Shift" key to enable persistent selection, and simultaneously select the parent node on the tree, saying "Western Australia", to include all the children by switching to the blue colour.-->
 The domestic tourism sees Sydney as one of the most popular destinations in realm of business and friends visiting over the years.
@@
 Time has cyclical structure, that may correspond to seasonal patterns to be discovered.
-The `index` component of the shared tsibble data forms the basis for exploring seasonality.
+The `index` component of the (shared) tsibble data forms the basis for exploring seasonality.
 To investigate for periodic or aperiodic patterns, series should be wrapped on themselves, where the index is broken into temporal components like quarter or day.
@@
 
-<!-- The shared tsibble data leverages the `key` attribute to converse with many coordinated views, with or without \CRANpkg{shiny}.
-On the other hand, a second critical attribute--`index`--lays the foundational temporal context that augments the conversation.
-When temporal data are plotted and stretched against the entire span like Figure \ref{fig:highlight-retail-1}, it puts emphasis on the trend perception.
-Yet to digest periodic/aperiodic patterns, data should be wrapped over relative time units that are origin-less, such as one quarter or one day.-->
-
-```{r wrap-ped, fig.show = "hold", fig.cap = "Snapshots wrapping after slicing the \\code{pedestrian20} data at different intervals, none (top), daily (middle) and weekly (bottom).
+```{r wrap-ped, fig.show = "hold", fig.cap = "Snapshots wrapping after slicing the \\code{pedestrian20} data at different intervals, (a) none, (b) daily and (c) weekly.
 This type of interaction is made possible with Shiny elements.", fig.subcap = c("Initial overview state", "1-day state", "7-day state, anchoring to Monday"), fig.ncol = 1}
@@
 Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four different locations, for three different time slices, daily, weekly and full five months.
-Plot \ref{fig:wrap-ped-1} shows hourly counts from January to May on an absolute time line, facetted by location.
+Plot \ref{fig:wrap-ped-1} shows hourly counts from January to May on an absolute time line, facetted by locations.
 The stage 3 COVID-19 lockdown, on March 16, is marked by a change of colour.

### 2020-09-26 tsibbletalk.tex
minor tweaks

@@
 cyclical structure, that may correspond to seasonal patterns to be
-discovered. The \texttt{index} component of the shared tsibble data
+discovered. The \texttt{index} component of the (shared) tsibble data
 forms the basis for exploring seasonality. To investigate for periodic
@@
 
-\caption[Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, none (top), daily (middle) and weekly (bottom)]{Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, none (top), daily (middle) and weekly (bottom).
+\caption[Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, (a) none, (b) daily and (c) weekly]{Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, (a) none, (b) daily and (c) weekly.
 This type of interaction is made possible with Shiny elements.}\label{fig:wrap-ped}
@@
 \ref{fig:wrap-ped-1} shows hourly counts from January to May on an
-absolute time line, facetted by location. The stage 3 COVID-19 lockdown,
-on March 16, is marked by a change of colour. (The pre-lockdown period
-is coloured with dark green and lockdown with orange.) We can see a
-significant decline in foot traffic at all four locations. QV Market is
-less affected probably because this is a major produce market, an
-essential service that continued to operate. Bourke St, a major shopping
-center, sees a gradual uptick in the last weeks of the period indicating
-that people were getting back into the shops.
+absolute time line, facetted by locations. The stage 3 COVID-19
+lockdown, on March 16, is marked by a change of colour. (The
+pre-lockdown period is coloured with dark green and lockdown with
+orange.) We can see a significant decline in foot traffic at all four
+locations. QV Market is less affected probably because this is a major
+produce market, an essential service that continued to operate. Bourke
+St, a major shopping center, sees a gradual uptick in the last weeks of
+the period indicating that people were getting back into the shops.
 

### 2020-10-04 cover-letter/cover-letter.Rmd
added cover letter

---
from: Earo Wang
fromqualifications: PhD (Statistics)
fromtitle: Lecturer in Statistics
fromWWW: earo.me
fromemail: earo.wang@auckland.ac.nz
fromphone: +64 9 923 9692
fromsignature: signature.png
toname: Editor
toaddress: The R Journal
opening: Dear Dr Michael Kane,
closing: Sincerely
spacing: 1.4
output: UOARmdTemplates::letter
---

Please consider our manuscript titled, "Conversations in time: interactive visualisation to explore structured temporal data" for publication in the R Journal.

This paper describes a new R package, **tsibbletalk**, for interactively exploring temporal data of nesting and crossing structures using linked views and time wrapping. It implements a shared tsibble data class to allow for linked brushing between html widgets on both client and server sides. It also provides a shiny module to wrap time lines for revealing periodic or aperiodic patterns.

We discuss **tsibbletalk** functions to easily build linked and dynamic graphics for temporal data, and include code examples using two contemporary data applications. The paper is written in RMarkdown, which contains the R code to reproduce the results.

We believe that this manuscript is a good fit for the R Journal, because it features novel implementation and nimble user experience in R.

### 2020-10-04 tsibbletalk.R

@@ -1,17 +1,3 @@
-## ----setup, echo = FALSE, cache = FALSE, include = FALSE----------------------
-options("knitr.graphics.auto_pdf" = TRUE)
-library(knitr)
-opts_chunk$set(
-  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
-  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
-  cache = TRUE, cache.path = 'cache/',
-  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
-)
-opts_knit$set(root.dir = here::here())
-read_chunk("scripts/demo.R")
-read_chunk("scripts/tourism.R")
-
-
 ## ---- load-pkgs---------------------------------------------------------------
 library(feasts)
 library(tsibble)
@@ -22,7 +8,7 @@ library(tsibbledata)
 print(aus_retail, n = 5)
 
 
-## ----highlight-retail, fig.height = 3.3, fig.cap = "ToDo"---------------------
+## ----highlight-retail---------------------
 library(tidyverse)
 library(ggrepel)
 library(gghighlight)
@@ -67,36 +53,33 @@ library(ggplot2)
 library(tsibble)
 
 
-## ----tourism-linking-fig, fig.cap = "ToDo"------------------------------------
-include_graphics("img/tourism-linking.png")
-
 
 ## ----plotly-key-tree, echo = TRUE---------------------------------------------
 p_l <- plotly_key_tree(tourism_shared, height = 1100, width = 800)
 
 
 ## ----tourism-series, echo = TRUE, eval = FALSE--------------------------------
-#> p_tr <- tourism_shared %>%
-#>   ggplot(aes(x = Month, y = Trips)) +
-#>   geom_line(aes(group = Region), alpha = .5, size = .4) +
-#>   facet_wrap(~ Purpose, scales = "free_y") +
-#>   scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
+p_tr <- tourism_shared %>%
+  ggplot(aes(x = Month, y = Trips)) +
+  geom_line(aes(group = Region), alpha = .5, size = .4) +
+  facet_wrap(~ Purpose, scales = "free_y") +
+  scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
 
 
[…]

### 2021-02-15 reviews/response.Rmd
added reviews

---
title: Response to reviewers
author: RJournal 2020-171
output:
  UOARmdTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

We thank the Editor and reviewers for their comments which have improved this paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.

# Reviewer 1 comments {-}

> Bottom of page 8: "Figure 3 peels the tidy module with" I'm not sure about the use of the word peels there?

We replaced "peels" with "reveals".

> Top of page 9: "The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage." Should "warrants" be replaced with "ensures" and perhaps replace "temporal data analysis for the later stage" with "subsequent temporal data analysis".

Done.

# Reviewer 2 comments {-}

> P4 L59 and the enhancements of -> and the enhancements by

Done.

> P6 L34. Second item missing full stop at the end.

Done.

> I think for Figure 4, you should mention that supplementary figure contains the gif if the animation does not work or perhaps include a link to a working animation.

Thanks. We added a sentence in Figure 4 caption "The animations are available with the supplementary materials online, and can also be viewed directly at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>."

> P15 L58. Mention `future_*` is from `furrr` package.

Done.

> P16 L24. Missing reference for `lubridate`.

Done.

### 2021-02-15 reviews/response.Rmd
started revising

@@
 
-We thank the Editor and reviewers for their comments which have improved this paper.
-The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
+We thank the reviewers for their comments which have improved this paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
 
@@
 
-> Bottom of page 8: "Figure 3 peels the tidy module with" I'm not sure about the use of the word peels there?
+> Please mention that tidyverts also contains the tsibbletalk package
 
-We replaced "peels" with "reveals".
+The tidyverts only contains {tsibble}, {fable}, and {feasts}. The {tsibbletalk} packages is NOT part of the tidyverts suite.
 
-> Top of page 9: "The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage." Should "warrants" be replaced with "ensures" and perhaps replace "temporal data analysis for the later stage" with "subsequent temporal data analysis".
+> The sentence "The techniques implemented in work described in this paper utilise web technology." could be more specific.
+One fix could include moving the sentence in the Summary section "At the heart of the tsibbletalk package is a blending of the best bits from tsibble, crosstalk, plotly, and shiny." to an earlier section in the paper.
+
+Thanks. Done.
+
+> In the section "Linking between plots" you write "... linking is by time, ...". It seems to me that linking is by key, so all time series points associated to a particular key are linked.
+
+Yes. We corrected from "by time" to "by key variables".
+
+> The phrasing "to allow for structured specifications in the key" is unclear.
+It would be better to be more explicit saying that it is to define the crossing and/or nesting relationships among the key variables.
 
@@
 
+> The phrasing "user-supplied session id" is wrong, a session means something very particular in shiny whereas the id argument if a module takes an id string (that should not be repeated in the same shiny namespace).
+
+Thanks. We corrected "user-supplied session id" to "user-supplied id string".
+
 # Reviewer 2 comments {-}
 
-> P4 L59 and the enhancements of -> and the enhancements by
+> Within the article, I would like to see more discussion of the actual linking implementation.
+I believe the package is using the R6 class and creating a subclass of SharedData from 'crosstalk', but it would be nice to see this explained within the article itself.
+A little note on why this is the preferred solution would also be of interest to readers
 
@@
 
-> P6 L34. Second item missing full stop at the end.
+> The code snippets lack library calls or the package::function_in_package format, for reproducibility.
 
-Done.
+Thanks. Added missing {ggplot2}, {plotly}, {shiny}, and {feasts} library calls in the code snippets.
 
-> I think for Figure 4, you should mention that supplementary figure contains the gif if the animation does not work or perhaps include a link to a working animation.
+> In the "Overview of interactivity", no mention is made of the non-web-based interactivity made available by the 'grDevices' package (https://protect-au.mimecast.com/s/xMG5CGv0kQtJA4DNh7AMnU?domain=github.com).
 
-Thanks.
-We added a sentence in Figure 4 caption "The animations are available with the supplementary materials online, and can also be viewed directly at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>."
-
-> P15 L58. Mention `future_*` is from `furrr` package.
-
-Done.
-
-> P16 L24. Missing reference for `lubridate`.
-
-Done.
+We added the mention of {grDevices} in the "Overview of interactivity" section: "Systems without the web technology include grDevices ...".

### 2021-02-15 tsibbletalk.Rmd
started revising

@@
 
-Systems without the web technology include \CRANpkg{loon} [@R-loon], based on Tcl/Tk, and \pkg{cranvas} [@xie_reactive_2014] based on Qt.
+Systems without the web technology include \CRANpkg{grDevices}, \CRANpkg{loon} [@R-loon], based on Tcl/Tk, and \pkg{cranvas} [@xie_reactive_2014] based on Qt.
 They offer a wide array of pre-defined interactions, such as selecting and zooming, to manipulate plots via mouse action, keyboard strokes, and menus.
@@
 
-The techniques implemented in work described in this paper utilise web technology.
+The techniques implemented in work described in this paper utilise web technology, including \CRANpkg{crosstalk}, \CRANpkg{plotly}, and R \CRANpkg{shiny}.
 
@@
 As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical variable linking where selecting one or more observations in one category will broadcast to all other observations in this category.
-That is, linking is by time: within the time series plot, click on any data point, and  whole line will be highlighted in response.
+That is, linking is by key variables: within the time series plot, click on any data point, and  whole line will be highlighted in response.
 The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking.
@@
 The `spec` argument in `as_shared_tsibble()` provides a means to construct hybrid linking, that incorporates hierarchical and categorical linking.
-A symbolic formula can be passed to the `spec` argument, to allow for structured specifications in the `key`.
+A symbolic formula can be passed to the `spec` argument, to define the crossing and/or nesting relationships among the key variables.
 Adopting @Wilkinson1973's notation for factorial models, the `spec` follows the `/` and `*` operator conventions to declare nesting and crossing variables, respectively.
@@
 ```{r tourism-shared, echo = TRUE}
-```
-
-```{r load-other-pkgs}
 ```
@@
 In general, a shiny module provides a vehicle for modularising shiny applications, relevant for both users and developers.
-As with all shiny modules, the first argument in both functions in \CRANpkg{tsibbletalk} requires a user-supplied session id that must be unique.
+As with all shiny modules, the first argument in both functions in \CRANpkg{tsibbletalk} requires a user-supplied id string that must be unique.
 The UI function `tsibbleWrapUI()` simply shows a slider that animates or controls the number of periods to be diced.

### 2021-02-15 tsibbletalk.tex
started revising

@@
 
-Systems without the web technology include \CRANpkg{loon}
-\citep{R-loon}, based on Tcl/Tk, and \pkg{cranvas}
+Systems without the web technology include \CRANpkg{grDevices},
+\CRANpkg{loon} \citep{R-loon}, based on Tcl/Tk, and \pkg{cranvas}
 \citep{xie_reactive_2014} based on Qt. They offer a wide array of
@@
 The techniques implemented in work described in this paper utilise web
-technology.
+technology, including \CRANpkg{crosstalk}, \CRANpkg{plotly}, and R
+\CRANpkg{shiny}.
 
@@
 one category will broadcast to all other observations in this category.
-That is, linking is by time: within the time series plot, click on any
-data point, and whole line will be highlighted in response. The
+That is, linking is by key variables: within the time series plot, click
+on any data point, and whole line will be highlighted in response. The
 \texttt{as\_shared\_tsibble()} uses \texttt{tsibble}'s \texttt{key}
@@
 categorical linking. A symbolic formula can be passed to the
-\texttt{spec} argument, to allow for structured specifications in the
-\texttt{key}. Adopting \citet{Wilkinson1973}'s notation for factorial
-models, the \texttt{spec} follows the \texttt{/} and \texttt{*} operator
-conventions to declare nesting and crossing variables, respectively. The
-\texttt{spec} for the \texttt{aus\_retail} data is therefore specified
-as \texttt{State\ *\ Industry} or \texttt{Industry\ *\ State}, which is
-the default for the presence of multiple \texttt{key} variables. If
-there is a hierarchy in the data, using \texttt{/} is required to
-indicate the parent-child relation, for a strictly one directional
-\texttt{parent/child}.
+\texttt{spec} argument, to define the crossing and/or nesting
+relationships among the key variables. Adopting \citet{Wilkinson1973}'s
+notation for factorial models, the \texttt{spec} follows the \texttt{/}
+and \texttt{*} operator conventions to declare nesting and crossing
+variables, respectively. The \texttt{spec} for the \texttt{aus\_retail}
+data is therefore specified as \texttt{State\ *\ Industry} or
+\texttt{Industry\ *\ State}, which is the default for the presence of
+multiple \texttt{key} variables. If there is a hierarchy in the data,
+using \texttt{/} is required to indicate the parent-child relation, for
+a strictly one directional \texttt{parent/child}.
 
@@
 \begin{Sinput}
+library(ggplot2)
 p_tr <- tourism_shared %>%
@@
 \begin{Sinput}
+library(feasts)
 tourism_feat <- tourism_shared %>%
@@
 \begin{Sinput}
+library(plotly)
 subplot(p_l,
@@
 modules, the first argument in both functions in \CRANpkg{tsibbletalk}
-requires a user-supplied session id that must be unique. The UI function
+requires a user-supplied id string that must be unique. The UI function
 \texttt{tsibbleWrapUI()} simply shows a slider that animates or controls
@@
 \begin{Sinput}
+library(shiny)
 p_line <- pedestrian20 %>%

### 2021-02-15 scripts/pedestrian.R

@@ -2,7 +2,6 @@
 library(rwalkr)
 library(tsibble)
 library(tsibbletalk)
-library(shiny)
 library(plotly)
 library(tidyverse)
 library(lubridate)
@@ -16,6 +15,7 @@ pedestrian20 <- melb_walk_fast(year = 2020, sensor = sensors) %>%
   as_tsibble(index = Date_Time, key = Sensor)
 
 ## ---- ped-slice
+library(shiny)
 p_line <- pedestrian20 %>%
   ggplot(aes(x = Date_Time, y = Count, colour = Lockdown)) +
   geom_line(size = .3) +

### 2021-02-15 scripts/tourism.R

@@ -3,16 +3,11 @@ library(tsibbletalk)
 tourism_shared <- tourism_monthly %>% 
   as_shared_tsibble(spec = (State / Region) * Purpose)
 
-## ---- load-other-pkgs
-library(feasts)
-library(plotly)
-library(ggplot2)
-library(tsibble)
-
 ## ---- plotly-key-tree
 p_l <- plotly_key_tree(tourism_shared, height = 1100, width = 800)
 
 ## ---- tourism-series
+library(ggplot2)
 p_tr <- tourism_shared %>%
   ggplot(aes(x = Month, y = Trips)) +
   geom_line(aes(group = Region), alpha = .5, size = .4) +
@@ -20,6 +15,7 @@ p_tr <- tourism_shared %>%
   scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
 
 ## ---- tourism-scatter
+library(feasts)
 tourism_feat <- tourism_shared %>%
   features(Trips, feat_stl)
 p_br <- tourism_feat %>%
@@ -27,6 +23,7 @@ p_br <- tourism_feat %>%
   geom_point(aes(group = Region), alpha = .8, size = 2)
 
 ## ---- tourism-multi
+library(plotly)
 subplot(p_l,
   subplot(
     ggplotly(p_tr, tooltip = "Region", width = 1100),

### 2021-03-01 reviews/response.Rmd
done with revision

@@
 
-We thank the reviewers for their comments which have improved this paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
+We thank the reviewers for their comments which have improved this paper.
+
+I've incorporated all feedback from reviewers regarding the package, except for adding unit tests.
+I'd like to include unit tests for code quality, but as far as I know, it's tricky to enable unit tests for interactive graphics.
+
+Regarding the article, the point by point description of changes are below: the reviewers' comments are in red and our response is in black.
 
@@
 
-The tidyverts only contains {tsibble}, {fable}, and {feasts}. The {tsibbletalk} packages is NOT part of the tidyverts suite.
+The core tidyverts only contains {tsibble}, {fable}, and {feasts}. The {tsibbletalk} packages is NOT part of the tidyverts suite.
 
@@
 
-Done.
+The actual linking is done through 'crosstalk'.
+We don't implement the linking rather than subclassing of its R6 object.
+I've added a note on why we prefer 'crosstalk' (or R6) as "This is an R6 object driving data transmission across multiple views, due to its mutable and lightweight properties." in the first paragraph of the section "Using a shared temporal data object for interactivity".
 

### 2021-03-20 scripts/tourism.R
minor tweak for publication

@@ -1,4 +1,5 @@
 ## ---- tourism-shared
+library(tsibble)
 library(tsibbletalk)
 tourism_shared <- tourism_monthly %>% 
   as_shared_tsibble(spec = (State / Region) * Purpose)

### 2021-07-18 tsibbletalk.Rmd
done with revision
minor tweak for publication
tweaks for publication

@@
 ---
-title: "Conversations in time: interactive visualisation to explore structured temporal data"
+title: "Conversations in time: interactive visualization to explore structured temporal data"
 author:
@@
   Temporal data often has a hierarchical structure, defined by categorical variables describing different levels, such as political regions or sales products.
-Nesting of categorical variables produces a hierarchical structure.
+The nesting of categorical variables produces a hierarchical structure.
 The \CRANpkg{tsibbletalk} package is developed to allow a user to interactively explore temporal data, relative to the nested or crossed structures.
 It can help to discover differences between category levels, and uncover interesting periodic or aperiodic slices.
-The package implements a shared `tsibble` object that allows for linked brushing between coordinated views, and a shiny module that aids in wrapping time lines for seasonal patterns.
+The package implements a shared `tsibble` object that allows for linked brushing between coordinated views, and a shiny module that aids in wrapping timelines for seasonal patterns.
 The tools are demonstrated using two data examples: domestic tourism in Australia and pedestrian traffic in Melbourne.
@@
 Temporal data typically arrives as a set of many observational units measured over time.
-Some variables may be categorical, containing a hierarchy in the collection process, that may be measurements taken in different geographic regions, or types of products sold by the one company.
+Some variables may be categorical, containing a hierarchy in the collection process, that may be measurements taken in different geographic regions, or types of products sold by one company.
 Exploring these multiple features can be daunting.
@@
 
-There is a long history of interactive data visualisation research and corresponding systems. Within R, the systems can be roughly divided into systems utilising web technology and those that don't.
+There is a long history of interactive data visualization research and corresponding systems. Within R, the systems can be roughly divided into systems utilizing web technology and those that do not.
 
@@
 They offer a wide array of pre-defined interactions, such as selecting and zooming, to manipulate plots via mouse action, keyboard strokes, and menus.
-The \pkg{cranvastime} package [@cheng_enabling_2016] is an add-on to \pkg{cranvas}, which provides specialised interactions for temporal data, such as wrapping and mirroring.
+The \pkg{cranvastime} package [@cheng_enabling_2016] is an add-on to \pkg{cranvas}, which provides specialized interactions for temporal data, such as wrapping and mirroring.
 
-The techniques implemented in work described in this paper utilise web technology, including \CRANpkg{crosstalk}, \CRANpkg{plotly}, and R \CRANpkg{shiny}.
+The techniques implemented in the work described in this paper utilize web technology, including \CRANpkg{crosstalk}, \CRANpkg{plotly}, and R \CRANpkg{shiny}.
 
@@
 The `as_shared_tsibble()` function turns a `tsibble` into a shared instance,  `SharedTsibbleData`, which is a subclass of `SharedData` from \CRANpkg{crosstalk}.
-This is the object driving data transmission across multiple views.
+This is an R6 object driving data transmission across multiple views, due to its mutable and lightweight properties.
 The \CRANpkg{tsibbletalk} package aims to streamline interactive exploration of temporal data, with the focus of temporal elements and structured linking.
@@
 
-As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical variable linking where selecting one or more observations in one category will broadcast to all other observations in this category.
-That is, linking is by key variables: within the time series plot, click on any data point, and  whole line will be highlighted in response.
+As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical variable linking, where selecting one or more observations in one category will broadcast to all other observations in this category.
+That is, linking is by key variables: within the time series plot, click on any data point, and the whole line will be highlighted in response.
 The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking.
@@
 It contains monthly domestic overnight trips across Australia.
-The `key` is comprised of three identifying variables: `State`, `Region`, and `Purpose` (of trip), in particular `State` nesting of `Region`, together crossed with `Purpose`.
+The `key` is comprised of three identifying variables: `State`, `Region`, and `Purpose` (of the trip), in particular `State` nesting of `Region`, crossed together with `Purpose`.
 This specification can be translated as follows:
@@
 
-\noindent There is a three-level hierarchy: the root node is implicitly Australia, and geographically disaggregated to states and lower-level tourism regions.
+\noindent There is a three-level hierarchy: the root node is implicitly Australia, geographically disaggregated to states, and lower-level tourism regions.
 A new handy function `plotly_key_tree()` has been implemented to help explore the hierarchy.
@@
 The following code line produces the linked tree diagram (left panel of Figure \ref{fig:tourism-linking-fig}).
-The visual for the tree hierarchy detangles a group of related series and provides a bird's eye view of the data organisation.
+The visual for the tree hierarchy detangles a group of related series and provides a bird's eye view of the data organization.
 
@@
 
-The tree plot provides the graphics skeleton, upon which the rest of data plots can be attached.
-In this example, small multiples of line plots are placed at the top right of Figure \ref{fig:tourism-linking-fig} to explore the temporal trend across regions, by trip purpose.
+The tree plot provides the graphics skeleton, upon which the rest of the data plots can be attached.
+In this example, small multiples of line plots are placed at the top right of Figure \ref{fig:tourism-linking-fig} to explore the temporal trend across regions by the trip purpose.
 The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code to create this.
@@
 To tease apart structure in the multiple time series, the `features()` function computes interesting characteristics, including the measures of trend and seasonality.
-These are displayed in the scatterplot at bottom right, where one dot represents one series.
+These are displayed in the scatterplot at the bottom right, where one dot represents one series.
 
@@
 There is one final step, to compose the three plots into an ensemble of coordinated views for exploration, shown in Figure \ref{fig:tourism-linking-fig}.
-(This is the interactive realisation of Figure \ref{fig:highlight-retail}).
+(This is the interactive realization of Figure \ref{fig:highlight-retail}).
 <!-- Routine functions bring about new interaction with temporal data on the client side.-->
@@
 (The names of the regions are a little odd, which is a quirk of the data set, but all four areas, Australia's South West, ...., correspond to tourist destinations in Western Australia.
-Hovering over the node on the branch brings up the state name.) This generated the response in the line plots and the scatterplot that coloured corresponding time series and points as blue.
+Hovering over the node on the branch brings up the state name.) This generated the response in the line plots and the scatterplot that colored corresponding time series and points as blue.
 2.
-To enable persistent selection, in oder to compare regions or states, "Shift" and click on the tree was done, after switching the colour to red.
+To enable persistent selection, in oder to compare regions or states, "Shift" and click on the tree was done, after switching the color to red.
 This generated the response that points and time series corresponding to Sydney were highlighted in red.
@@
 
-The domestic tourism sees Sydney as one of the most popular destinations in realm of business and friends visiting over the years.
-Despite the relatively weaker performance in Western Australia, Australia's North West region sees a strongest upward trend in Business, bypassing Sydney in some years.
+Domestic tourism sees Sydney as one of the most popular destinations in the realm of business and friends visiting over the years.
+Despite the relatively weaker performance in Western Australia, Australia's North West region sees a strongest upward trend in business, bypassing Sydney in some years.
 
 In summary, shared tsibble data nicely bridges between the \CRANpkg{crosstalk} and **tidyverts** ecosystems for temporal data using the common "key".
-The `as_shared_tsibble()` provides a symbolic user interface for effortless construction of a hybrid of hierarchical and categorical linking between plots.
+The `as_shared_tsibble()` provides a symbolic user interface for the effortless construction of a hybrid of hierarchical and categorical linking between plots.
 The `plotly_key_tree()` function, in turn, decodes the hierarchical specification to plot a tree for data overview and navigation, when accompanied by more detailed plots.
@@
 An important aspect of temporal data is the time context.
-Time has cyclical structure, that may correspond to seasonal patterns to be discovered.
+Time has a cyclical structure, that may correspond to seasonal patterns to be discovered.
 The `index` component of the (shared) tsibble data forms the basis for exploring seasonality.
@@
 
-The city of Melbourne has sensors installed at various locations, to record hourly counts of pedestrians, in order to capture daily rhythms of the downtown [@ped].
+The city of Melbourne has sensors installed at various locations, to record hourly counts of pedestrians, in order to capture the daily rhythms of the downtown [@ped].
 Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four different locations, for three different time slices, daily, weekly and full five months.
-Plot \ref{fig:wrap-ped-1} shows hourly counts from January to May on an absolute time line, facetted by locations.
-The stage 3 COVID-19 lockdown, on March 16, is marked by a change of colour.
-(The pre-lockdown period is coloured with dark green and lockdown with orange.) We can see a significant decline in foot traffic  at all four locations.
+Plot \ref{fig:wrap-ped-1} shows hourly counts from January to May on an absolute timeline, facetted by locations.
+The stage 3 COVID-19 lockdown, on March 16, is marked by a change of color.
+(The pre-lockdown period is colored with dark green and lockdown with orange.) We can see a significant decline in foot traffic  at all four locations.
 QV Market is less affected probably because this is a major produce market, an essential service that continued to operate.
@@
 
-Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} show slicing and wrapping of the series into daily and weekly sections, respectively.
+Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} show the slicing and wrapping of the series into daily and weekly sections, respectively.
 Multiple seasonalities pop out.
@@
 Birrarung Marr is the most affected.
-This is the location of special events, and it is clear that these have completely disappeared during lockdown.
+This is the location of special events, and it is clear that these have completely disappeared during the lockdown.
 
@@
 The following code snippet generates Figure \ref{fig:wrap-ped}.
+The creation of the `pedestrian20` data is available in supplementary R files.
 
@@
 Figure \ref{fig:wrap-ped-1} corresponds to the initial state, with the slider incremented by 1-day units.
-The "play" button near the end of slider can automatically animate the slicing and dicing process, walking the viewer through all 24 hours of the 152 days.
+The "play" button near the end of the slider can automatically animate the slicing and dicing process, walking the viewer through all 24 hours of the 152 days.
 Alternatively, users can drag the slider to examine selected slices.
@@
 The new shiny module exploits the temporal aspect for a tsibble object, available through the `index` attribute.
-It allows users to slide through relative periods to digest seasonal behaviours, with nimble user experience.
+It allows users to slide through relative periods to digest seasonal behaviors, with a nimble user experience.
 
@@
 The `as_shared_tsibble()` turns a `tsibble` object to a shared data class, with an option to express any nesting and crossing structures from the `key` attribute.
-If nesting is found in the data, the `plotly_key_tree()` creates an interactive hierarchical tree to help with data overview.
+If nesting is found in the data, the `plotly_key_tree()` creates an interactive hierarchical tree to help with the data overview.
 This sets the stage for hierarchical and categorical linking between multiple views from one shared tsibble.

### 2021-07-18 tsibbletalk.tex
done with revision
tweaks for publication

@@
 % !TeX root = RJwrapper.tex
-\title{Conversations in time: interactive visualisation to explore
+\title{Conversations in time: interactive visualization to explore
 structured temporal data}
@@
 variables describing different levels, such as political regions or
-sales products. Nesting of categorical variables produces a hierarchical
-structure. The \CRANpkg{tsibbletalk} package is developed to allow a
-user to interactively explore temporal data, relative to the nested or
-crossed structures. It can help to discover differences between category
-levels, and uncover interesting periodic or aperiodic slices. The
-package implements a shared \texttt{tsibble} object that allows for
-linked brushing between coordinated views, and a shiny module that aids
-in wrapping time lines for seasonal patterns. The tools are demonstrated
-using two data examples: domestic tourism in Australia and pedestrian
-traffic in Melbourne.
+sales products. The nesting of categorical variables produces a
+hierarchical structure. The \CRANpkg{tsibbletalk} package is developed
+to allow a user to interactively explore temporal data, relative to the
+nested or crossed structures. It can help to discover differences
+between category levels, and uncover interesting periodic or aperiodic
+slices. The package implements a shared \texttt{tsibble} object that
+allows for linked brushing between coordinated views, and a shiny module
+that aids in wrapping timelines for seasonal patterns. The tools are
+demonstrated using two data examples: domestic tourism in Australia and
+pedestrian traffic in Melbourne.
 }
@@
 hierarchy in the collection process, that may be measurements taken in
-different geographic regions, or types of products sold by the one
-company. Exploring these multiple features can be daunting. Ensemble
-graphics \citep{unwin_ensemble_2018} bundle multiple views of a data set
-together into one composite figure. These provide an effective approach
-for exploring and digesting many different aspects of temporal data.
-Adding interactivity to the ensemble can greatly enhance the exploration
+different geographic regions, or types of products sold by one company.
+Exploring these multiple features can be daunting. Ensemble graphics
+\citep{unwin_ensemble_2018} bundle multiple views of a data set together
+into one composite figure. These provide an effective approach for
+exploring and digesting many different aspects of temporal data. Adding
+interactivity to the ensemble can greatly enhance the exploration
 process.
@@
 
-There is a long history of interactive data visualisation research and
+There is a long history of interactive data visualization research and
 corresponding systems. Within R, the systems can be roughly divided into
-systems utilising web technology and those that don't.
+systems utilizing web technology and those that do not.
 
@@
 \pkg{cranvastime} package \citep{cheng_enabling_2016} is an add-on to
-\pkg{cranvas}, which provides specialised interactions for temporal
+\pkg{cranvas}, which provides specialized interactions for temporal
 data, such as wrapping and mirroring.
 
-The techniques implemented in work described in this paper utilise web
-technology, including \CRANpkg{crosstalk}, \CRANpkg{plotly}, and R
+The techniques implemented in the work described in this paper utilize
+web technology, including \CRANpkg{crosstalk}, \CRANpkg{plotly}, and R
 \CRANpkg{shiny}.
@@
 shared instance, \texttt{SharedTsibbleData}, which is a subclass of
-\texttt{SharedData} from \CRANpkg{crosstalk}. This is the object driving
-data transmission across multiple views. The \CRANpkg{tsibbletalk}
-package aims to streamline interactive exploration of temporal data,
-with the focus of temporal elements and structured linking.
+\texttt{SharedData} from \CRANpkg{crosstalk}. This is an R6 object
+driving data transmission across multiple views, due to its mutable and
+lightweight properties. The \CRANpkg{tsibbletalk} package aims to
+streamline interactive exploration of temporal data, with the focus of
+temporal elements and structured linking.
 
@@
 As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to
-categorical variable linking where selecting one or more observations in
-one category will broadcast to all other observations in this category.
-That is, linking is by key variables: within the time series plot, click
-on any data point, and whole line will be highlighted in response. The
-\texttt{as\_shared\_tsibble()} uses \texttt{tsibble}'s \texttt{key}
-variables to achieve these types of linking.
+categorical variable linking, where selecting one or more observations
+in one category will broadcast to all other observations in this
+category. That is, linking is by key variables: within the time series
+plot, click on any data point, and the whole line will be highlighted in
+response. The \texttt{as\_shared\_tsibble()} uses \texttt{tsibble}'s
+\texttt{key} variables to achieve these types of linking.
 
@@
 comprised of three identifying variables: \texttt{State},
-\texttt{Region}, and \texttt{Purpose} (of trip), in particular
-\texttt{State} nesting of \texttt{Region}, together crossed with
+\texttt{Region}, and \texttt{Purpose} (of the trip), in particular
+\texttt{State} nesting of \texttt{Region}, crossed together with
 \texttt{Purpose}. This specification can be translated as follows:
@@
 \begin{Sinput}
+library(tsibble)
 library(tsibbletalk)
@@
 \noindent There is a three-level hierarchy: the root node is implicitly
-Australia, and geographically disaggregated to states and lower-level
+Australia, geographically disaggregated to states, and lower-level
 tourism regions. A new handy function \texttt{plotly\_key\_tree()} has
@@
 for the tree hierarchy detangles a group of related series and provides
-a bird's eye view of the data organisation.
+a bird's eye view of the data organization.
 
@@
 
-The tree plot provides the graphics skeleton, upon which the rest of
+The tree plot provides the graphics skeleton, upon which the rest of the
 data plots can be attached. In this example, small multiples of line
@@
 \ref{fig:tourism-linking-fig} to explore the temporal trend across
-regions, by trip purpose. The shared tsibble data can be directly piped
-into \CRANpkg{ggplot2} code to create this.
+regions by the trip purpose. The shared tsibble data can be directly
+piped into \CRANpkg{ggplot2} code to create this.
 
@@
 interesting characteristics, including the measures of trend and
-seasonality. These are displayed in the scatterplot at bottom right,
+seasonality. These are displayed in the scatterplot at the bottom right,
 where one dot represents one series.
@@
 coordinated views for exploration, shown in Figure
-\ref{fig:tourism-linking-fig}. (This is the interactive realisation of
+\ref{fig:tourism-linking-fig}. (This is the interactive realization of
 Figure \ref{fig:highlight-retail}).
@@
   the node on the branch brings up the state name.) This generated the
-  response in the line plots and the scatterplot that coloured
+  response in the line plots and the scatterplot that colored
   corresponding time series and points as blue.
@@
   To enable persistent selection, in oder to compare regions or states,
-  ``Shift'' and click on the tree was done, after switching the colour
-  to red. This generated the response that points and time series
+  ``Shift'' and click on the tree was done, after switching the color to
+  red. This generated the response that points and time series
   corresponding to Sydney were highlighted in red.
@@
 
-The domestic tourism sees Sydney as one of the most popular destinations
-in realm of business and friends visiting over the years. Despite the
+Domestic tourism sees Sydney as one of the most popular destinations in
+the realm of business and friends visiting over the years. Despite the
 relatively weaker performance in Western Australia, Australia's North
-West region sees a strongest upward trend in Business, bypassing Sydney
+West region sees a strongest upward trend in business, bypassing Sydney
 in some years.
@@
 using the common ``key''. The \texttt{as\_shared\_tsibble()} provides a
-symbolic user interface for effortless construction of a hybrid of
+symbolic user interface for the effortless construction of a hybrid of
 hierarchical and categorical linking between plots. The
@@
 
-An important aspect of temporal data is the time context. Time has
+An important aspect of temporal data is the time context. Time has a
 cyclical structure, that may correspond to seasonal patterns to be
@@
 The city of Melbourne has sensors installed at various locations, to
-record hourly counts of pedestrians, in order to capture daily rhythms
-of the downtown \citep{ped}. Figure \ref{fig:wrap-ped} shows the first
-five months of 2020 foot traffic at four different locations, for three
-different time slices, daily, weekly and full five months. Plot
+record hourly counts of pedestrians, in order to capture the daily
+rhythms of the downtown \citep{ped}. Figure \ref{fig:wrap-ped} shows the
+first five months of 2020 foot traffic at four different locations, for
+three different time slices, daily, weekly and full five months. Plot
 \ref{fig:wrap-ped-1} shows hourly counts from January to May on an
-absolute time line, facetted by locations. The stage 3 COVID-19
-lockdown, on March 16, is marked by a change of colour. (The
-pre-lockdown period is coloured with dark green and lockdown with
-orange.) We can see a significant decline in foot traffic at all four
-locations. QV Market is less affected probably because this is a major
-produce market, an essential service that continued to operate. Bourke
-St, a major shopping center, sees a gradual uptick in the last weeks of
-the period indicating that people were getting back into the shops.
+absolute timeline, facetted by locations. The stage 3 COVID-19 lockdown,
+on March 16, is marked by a change of color. (The pre-lockdown period is
+colored with dark green and lockdown with orange.) We can see a
+significant decline in foot traffic at all four locations. QV Market is
+less affected probably because this is a major produce market, an
+essential service that continued to operate. Bourke St, a major shopping
+center, sees a gradual uptick in the last weeks of the period indicating
+that people were getting back into the shops.
 
-Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} show slicing and
-wrapping of the series into daily and weekly sections, respectively.
+Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} show the slicing
+and wrapping of the series into daily and weekly sections, respectively.
 Multiple seasonalities pop out. There tends to be a daily pattern,
@@
 the most affected. This is the location of special events, and it is
-clear that these have completely disappeared during lockdown.
+clear that these have completely disappeared during the lockdown.
 
@@
 ``day'' to a much coarser ``year''. The following code snippet generates
-Figure \ref{fig:wrap-ped}.
+Figure \ref{fig:wrap-ped}. The creation of the \texttt{pedestrian20}
+data is available in supplementary R files.
 
@@
 slider incremented by 1-day units. The ``play'' button near the end of
-slider can automatically animate the slicing and dicing process, walking
-the viewer through all 24 hours of the 152 days. Alternatively, users
-can drag the slider to examine selected slices.
+the slider can automatically animate the slicing and dicing process,
+walking the viewer through all 24 hours of the 152 days. Alternatively,
+users can drag the slider to examine selected slices.
 
@@
 available through the \texttt{index} attribute. It allows users to slide
-through relative periods to digest seasonal behaviours, with nimble user
-experience.
+through relative periods to digest seasonal behaviors, with a nimble
+user experience.
 
@@
 data, the \texttt{plotly\_key\_tree()} creates an interactive
-hierarchical tree to help with data overview. This sets the stage for
-hierarchical and categorical linking between multiple views from one
+hierarchical tree to help with the data overview. This sets the stage
+for hierarchical and categorical linking between multiple views from one
 shared tsibble.
@@
 %
-\\\href{mailto:earo.wang@auckland.ac.nz}{\nolinkurl{earo.wang@auckland.ac.nz}}
+\href{mailto:earo.wang@auckland.ac.nz}{\nolinkurl{earo.wang@auckland.ac.nz}}%
 }
@@
 %
-\\\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
+\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}%
 }

### 2021-07-18 tsibbletalk.R

@@ -1,3 +1,18 @@
+## ----setup, echo = FALSE, cache = FALSE, include = FALSE----------------------
+options("knitr.graphics.auto_pdf" = TRUE)
+library(knitr)
+opts_chunk$set(
+  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
+  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
+  cache = TRUE, cache.path = 'cache/',
+  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
+)
+opts_knit$set(root.dir = here::here())
+read_chunk("scripts/demo.R")
+read_chunk("scripts/tourism.R")
+read_chunk("scripts/pedestrian.R")
+
+
 ## ---- load-pkgs---------------------------------------------------------------
 library(feasts)
 library(tsibble)
@@ -8,15 +23,14 @@ library(tsibbledata)
 print(aus_retail, n = 5)
 
 
-## ----highlight-retail---------------------
+## ----highlight-retail, fig.height = 6, out.width = ".49\\linewidth", fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted. (a) An overlaid time series plot. (b) A scatter plot drawn from their time series features, where each dot represents a time series from (a).", fig.subcap = c("", ""), fig.ncol = 2----
 library(tidyverse)
 library(ggrepel)
 library(gghighlight)
-library(patchwork)
 
-p_retail_ts <- aus_retail %>% 
+aus_retail %>%
   as_tibble() %>% # gghighlight issue for group_by() + filter()
-  mutate(group = paste(State, ":", Industry)) %>% 
+  mutate(group = paste(State, ":", Industry)) %>%
   ggplot(aes(x = Month, y = Turnover)) +
   geom_line(aes(group = group)) +
   gghighlight(
@@ -25,9 +39,9 @@ p_retail_ts <- aus_retail %>%
   ) +
   scale_x_yearmonth(breaks = yearmonth(c("1990 Jan", "2000 Jan", "2010 Jan")))
 
-p_retail_feat <- aus_retail %>% 
-  features(Turnover, feat_stl) %>% 
-  mutate(group = paste(State, ":", Industry)) %>% 
+aus_retail %>%
+  features(Turnover, feat_stl) %>%
+  mutate(group = paste(State, ":", Industry)) %>%
[…]

### 2021-07-28 rjdistill_version/tsibbletalk.Rmd
Rmd reworked to fit new format

--- 2021-07-18 tsibbletalk.Rmd
@@
 ---
-title: "Conversations in time: interactive visualization to explore structured temporal data"
+title: "Conversations in Time: Interactive Visualization to Explore Structured Temporal Data"
 author:
@@
   email: earo.wang@auckland.ac.nz
+  orcid: 0000-0001-6448-5260
 - name: Dianne Cook
@@
   email: dicook@monash.edu
-bibliography: tsibbletalk.bib
+  orcid: 0000-0002-3813-7155
+date: "`r Sys.Date()`"
 abstract: >
@@
   \usepackage{subfig}
-output: rticles::rjournal_article
+creative_commons: CC BY
+output: 
+  - rjtools::rjournal_web_article
+  - rjtools::rjournal_pdf_article
+# per R journal requirement, the bib filename should be the same as the output 
+# tex file. Don't forget to rename the bib file and change this example value.
+bibliography: tsibbletalk.bib
 ---

### 2021-07-28 rjdistill_version/tsibbletalk.Rmd
new r chunk for external scripts

@@
 opts_knit$set(root.dir = here::here())
+```
+
+```{r external}
 read_chunk("scripts/demo.R")
@@
 
-```{r, load-pkgs}
+```{r load-pkgs}
 ```

### 2021-07-28 rjdistill_version/tsibbletalk.Rmd
standalone rjdistill

@@
   fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
-  cache = TRUE, cache.path = 'cache/',
   out.width = ifelse(is_html_output(), "100%", "\\textwidth")
@@
 
-Figure \ref{fig:highlight-retail} illustrates examining temporal data with many keys.
+Figure \@ref(fig:highlight-retail) illustrates examining temporal data with many keys.
 The data has 152 series corresponding to different industries in retail data.
@@
 The feature space is computed using the `features()` function from \CRANpkg{feasts}, which summarises the original data for each series using various statistical features.
-This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, which can be graphically displayed as in Figure \ref{fig:highlight-retail-2}.
+This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, which can be graphically displayed as in Figure \@ref(fig:highlight-retail-2).
 
-```{r highlight-retail, fig.height = 6, out.width = ".49\\linewidth", fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted.
+```{r highlight-retail, fig.height = 6, out.width = "49%", fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted.
 (a) An overlaid time series plot.
@@
 
-Figure \ref{fig:highlight-retail} has also been highlighted to focus on the one series with the strongest seasonality.
+Figure \@ref(fig:highlight-retail) has also been highlighted to focus on the one series with the strongest seasonality.
 To create this highlighting, one needs to first filter the interesting series from the features table, and join back to the original `tsibble` in order to examine its trend in relation to others.
@@
 ```{r tourism-linking-fig, fig.cap = "Snapshot of exploring an ensemble of linked plots of the Australian tourism data, built on a \\code{tourism\\_shared} object.
-It also illustrates persistent linked brushing to compare two groups."}
+It also illustrates persistent linked brushing to compare two groups.", eval = knitr::is_latex_output()}
 include_graphics("img/tourism-linking.png")
@@
 It interprets hierarchies in the shared tsibble's `spec` as a tree view, built with \CRANpkg{plotly}.
-The following code line produces the linked tree diagram (left panel of Figure \ref{fig:tourism-linking-fig}).
+The following code line produces the linked tree diagram (left panel of Figure \@ref(fig:tourism-multi)).
 The visual for the tree hierarchy detangles a group of related series and provides a bird's eye view of the data organization.
@@
 The tree plot provides the graphics skeleton, upon which the rest of the data plots can be attached.
-In this example, small multiples of line plots are placed at the top right of Figure \ref{fig:tourism-linking-fig} to explore the temporal trend across regions by the trip purpose.
+In this example, small multiples of line plots are placed at the top right of Figure \@ref(fig:tourism-linking-fig) to explore the temporal trend across regions by the trip purpose.
 The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code to create this.
 
-```{r tourism-series, echo = TRUE, eval = FALSE}
+```{r tourism-series, echo = TRUE, eval = knitr::is_html_output()}
 ```
@@
 
-```{r tourism-scatter, echo = TRUE, eval = FALSE}
+```{r tourism-scatter, echo = TRUE, eval = knitr::is_html_output()}
 ```
 
-There is one final step, to compose the three plots into an ensemble of coordinated views for exploration, shown in Figure \ref{fig:tourism-linking-fig}.
-(This is the interactive realization of Figure \ref{fig:highlight-retail}).
+There is one final step, to compose the three plots into an ensemble of coordinated views for exploration, shown in Figure \@ref(fig:tourism-linking-fig).
+(This is the interactive realization of Figure \@ref(fig:highlight-retail)).
 <!-- Routine functions bring about new interaction with temporal data on the client side.-->
 
-```{r tourism-multi, echo = TRUE, eval = FALSE}
+```{r tourism-multi, echo = TRUE, eval = knitr::is_html_output(), fig.cap = "Exploring an ensemble of linked plots of the Australian tourism data, built on a \\code{tourism\\_shared} object.
+Click one of the nodes in the hierarchical tree to enable persistent linked brushing to compare two groups.
+Points and lines can also be selected in other plots."}
 ```
@@
 Given the `spec`, clicking either one element in any plot highlights all points that match the `Region` category, that is, categorical linking.
-Figure \ref{fig:tourism-linking-fig} is a static view of an interactive exploration.
+Figure \@ref(fig:tourism-multi) provides an interactive exploration.
 The steps in getting to this point were:
@@
 2.
-To enable persistent selection, in oder to compare regions or states, "Shift" and click on the tree was done, after switching the color to red.
+To enable persistent selection, in order to compare regions or states, "Shift" and click on the tree was done, after switching the color to red.
 This generated the response that points and time series corresponding to Sydney were highlighted in red.
@@
 ```{r wrap-ped, fig.show = "hold", fig.cap = "Snapshots wrapping after slicing the \\code{pedestrian20} data at different intervals, (a) none, (b) daily and (c) weekly.
-This type of interaction is made possible with Shiny elements.", fig.subcap = c("Initial overview state", "1-day state", "7-day state, anchoring to Monday"), fig.ncol = 1}
+This type of interaction is made possible with Shiny elements.", fig.subcap = c("Initial overview state", "1-day state", "7-day state, anchoring to Monday"), fig.ncol = 1, eval = knitr::is_latex_output()}
 include_graphics("img/wrap-0.png")
@@
 
+```{r wrap-ped-html, fig.show = "hold", fig.cap = "Animations showing wrapping after slicing the \\code{pedestrian20} data at different intervals, (a) none, (b) daily and (c) weekly.
+This type of interaction is made possible with Shiny elements.", fig.subcap = c("Initial overview state", "1-day state", "7-day state, anchoring to Monday"), fig.ncol = 1, eval = knitr::is_html_output()}
+include_graphics("img/shiny-wrap.gif")
+```
+
 The city of Melbourne has sensors installed at various locations, to record hourly counts of pedestrians, in order to capture the daily rhythms of the downtown [@ped].
-Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four different locations, for three different time slices, daily, weekly and full five months.
-Plot \ref{fig:wrap-ped-1} shows hourly counts from January to May on an absolute timeline, facetted by locations.
+Figure \@ref(fig:wrap-ped-html) shows the first five months of 2020 foot traffic at four different locations, for three different time slices, daily, weekly and full five months.
+Plot \@ref(fig:wrap-ped-1) shows hourly counts from January to May on an absolute timeline, facetted by locations.
 The stage 3 COVID-19 lockdown, on March 16, is marked by a change of color.
@@
 
-Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} show the slicing and wrapping of the series into daily and weekly sections, respectively.
+Figure \@ref(fig:wrap-ped-html) shows the slicing and wrapping of the series into daily and weekly sections, respectively.
 Multiple seasonalities pop out.
@@
 The wrapping procedure involves slicing the time index into seasonal periods of interest, and the result is diced time.
-For example, hourly pedestrian data can be decomposed into 24-hour blocks, which then overlays the counts for all respective days, as done in plot \ref{fig:wrap-ped-2}.
+For example, hourly pedestrian data can be decomposed into 24-hour blocks, which then overlays the counts for all respective days, as done in plot \@ref(fig:wrap-ped-2).
 For exploration, this slice position should be controlled interactively, so that many different slices can be examined rapidly.
@@
 For date-times (represented by `POSIXt`), the granularity ranges from fine "day" to a much coarser "year".
-The following code snippet generates Figure \ref{fig:wrap-ped}.
+The following code snippet generates Figure \@ref(fig:wrap-ped).
 The creation of the `pedestrian20` data is available in supplementary R files.
@@
 
-Figure \ref{fig:wrap-ped-1} corresponds to the initial state, with the slider incremented by 1-day units.
+Figure \@ref(fig:wrap-ped-1) corresponds to the initial state, with the slider incremented by 1-day units.
 The "play" button near the end of the slider can automatically animate the slicing and dicing process, walking the viewer through all 24 hours of the 152 days.

### 2021-07-28 rjdistill_version/tsibbletalk.tex
standalone rjdistill

--- 2021-07-18 tsibbletalk.tex
@@
 % !TeX root = RJwrapper.tex
-\title{Conversations in time: interactive visualization to explore
-structured temporal data}
+\title{Conversations in Time: Interactive Visualization to Explore
+Structured Temporal Data}
 \author{by Earo Wang and Dianne Cook}
@@
 
-Figure \ref{fig:highlight-retail} illustrates examining temporal data
+Figure @ref(fig:highlight-retail) illustrates examining temporal data
 with many keys. The data has 152 series corresponding to different
@@
   subplot(
-    ggplotly(p_tr, tooltip = "Region", width = 1100),
-    ggplotly(p_br, tooltip = "Region", width = 1100),
+    ggplotly(p_tr, tooltip = "Region", width = 700, height = 800),
+    ggplotly(p_br, tooltip = "Region", width = 700, height = 800),
     nrows = 2),
@@
 \item
-  To enable persistent selection, in oder to compare regions or states,
+  To enable persistent selection, in order to compare regions or states,
   ``Shift'' and click on the tree was done, after switching the color to
@@
 %
-%
+\textit{ORCiD: \href{https://orcid.org/0000-0001-6448-5260}{0000-0001-6448-5260}}\\%
 \href{mailto:earo.wang@auckland.ac.nz}{\nolinkurl{earo.wang@auckland.ac.nz}}%
@@
 %
-%
+\textit{ORCiD: \href{https://orcid.org/0000-0002-3813-7155}{0000-0002-3813-7155}}\\%
 \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}%

### 2021-07-28 rjdistill_version/tsibbletalk.R

@@ -1,0 +1,151 @@
+## ----setup, echo = FALSE, cache = FALSE, include = FALSE----------------------
+options("knitr.graphics.auto_pdf" = TRUE)
+library(knitr)
+opts_chunk$set(
+  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
+  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
+  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
+)
+opts_knit$set(root.dir = here::here())
+
+
+## ----external-----------------------------------------------------------------
+read_chunk("scripts/demo.R")
+read_chunk("scripts/tourism.R")
+read_chunk("scripts/pedestrian.R")
+
+
+## ----load-pkgs----------------------------------------------------------------
+library(feasts)
+library(tsibble)
+library(tsibbledata)
+
+
+## ---- print-retail------------------------------------------------------------
+print(aus_retail, n = 5)
+
+
+## ----highlight-retail, fig.height = 6, out.width = ".49\\linewidth", fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted. (a) An overlaid time series plot. (b) A scatter plot drawn from their time series features, where each dot represents a time series from (a).", fig.subcap = c("", ""), fig.ncol = 2----
+library(tidyverse)
+library(ggrepel)
+library(gghighlight)
+
+aus_retail %>%
+  as_tibble() %>% # gghighlight issue for group_by() + filter()
+  mutate(group = paste(State, ":", Industry)) %>%
+  ggplot(aes(x = Month, y = Turnover)) +
+  geom_line(aes(group = group)) +
+  gghighlight(
+    State == "Queensland", Industry == "Department stores",
+    label_params = list(vjust = -2)
+  ) +
+  scale_x_yearmonth(breaks = yearmonth(c("1990 Jan", "2000 Jan", "2010 Jan")))
+
+aus_retail %>%
+  features(Turnover, feat_stl) %>%
+  mutate(group = paste(State, ":", Industry)) %>%
+  ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
+  geom_point(aes(group = group)) +
+  labs(x = "Trend strength", y = "Seasonal strength") +
+  gghighlight(
[…]

### 2021-07-28 scripts/tourism.R

@@ -1,7 +1,7 @@
 ## ---- tourism-shared
 library(tsibble)
 library(tsibbletalk)
-tourism_shared <- tourism_monthly %>% 
+tourism_shared <- tourism_monthly %>%
   as_shared_tsibble(spec = (State / Region) * Purpose)
 
 ## ---- plotly-key-tree
@@ -27,8 +27,8 @@ p_br <- tourism_feat %>%
 library(plotly)
 subplot(p_l,
   subplot(
-    ggplotly(p_tr, tooltip = "Region", width = 1100),
-    ggplotly(p_br, tooltip = "Region", width = 1100),
+    ggplotly(p_tr, tooltip = "Region", width = 700, height = 800),
+    ggplotly(p_br, tooltip = "Region", width = 700, height = 800),
     nrows = 2),
   widths = c(.4, .6)) %>%
   highlight(dynamic = TRUE)

### 2021-07-29 rjdistill_version/tsibbletalk.Rmd
more fixes about rjdistill

@@
 The feature space is computed using the `features()` function from \CRANpkg{feasts}, which summarises the original data for each series using various statistical features.
-This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, which can be graphically displayed as in Figure \@ref(fig:highlight-retail-2).
+This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, which can be graphically displayed as in Figure \@ref(fig:highlight-retail) (right).
 
-```{r highlight-retail, fig.height = 6, out.width = "49%", fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted.
+```{r highlight-retail, fig.height = 3.6, fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted.
 (a) An overlaid time series plot.
-(b) A scatter plot drawn from their time series features, where each dot represents a time series from (a).", fig.subcap = c("", ""), fig.ncol = 2}
+(b) A scatter plot drawn from their time series features, where each dot represents a time series from (a)."}
 ```
@@
 The tree plot provides the graphics skeleton, upon which the rest of the data plots can be attached.
-In this example, small multiples of line plots are placed at the top right of Figure \@ref(fig:tourism-linking-fig) to explore the temporal trend across regions by the trip purpose.
+In this example, small multiples of line plots are placed at the top right of Figure \@ref(fig:tourism-multi) to explore the temporal trend across regions by the trip purpose.
 The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code to create this.
@@
 
-There is one final step, to compose the three plots into an ensemble of coordinated views for exploration, shown in Figure \@ref(fig:tourism-linking-fig).
+There is one final step, to compose the three plots into an ensemble of coordinated views for exploration, shown in Figure \@ref(fig:tourism-multi).
 (This is the interactive realization of Figure \@ref(fig:highlight-retail)).
@@
 Figure \@ref(fig:wrap-ped-html) shows the first five months of 2020 foot traffic at four different locations, for three different time slices, daily, weekly and full five months.
-Plot \@ref(fig:wrap-ped-1) shows hourly counts from January to May on an absolute timeline, facetted by locations.
+Plot \@ref(fig:wrap-ped-html) shows hourly counts from January to May on an absolute timeline, facetted by locations.
 The stage 3 COVID-19 lockdown, on March 16, is marked by a change of color.
@@
 The wrapping procedure involves slicing the time index into seasonal periods of interest, and the result is diced time.
-For example, hourly pedestrian data can be decomposed into 24-hour blocks, which then overlays the counts for all respective days, as done in plot \@ref(fig:wrap-ped-2).
+For example, hourly pedestrian data can be decomposed into 24-hour blocks, which then overlays the counts for all respective days, as done in plot \@ref(fig:wrap-ped-html).
 For exploration, this slice position should be controlled interactively, so that many different slices can be examined rapidly.
@@
 For date-times (represented by `POSIXt`), the granularity ranges from fine "day" to a much coarser "year".
-The following code snippet generates Figure \@ref(fig:wrap-ped).
+The following code snippet generates Figure \@ref(fig:wrap-ped-html).
 The creation of the `pedestrian20` data is available in supplementary R files.
@@
 
-Figure \@ref(fig:wrap-ped-1) corresponds to the initial state, with the slider incremented by 1-day units.
+Figure \@ref(fig:wrap-ped-html) corresponds to the initial state, with the slider incremented by 1-day units.
 The "play" button near the end of the slider can automatically animate the slicing and dicing process, walking the viewer through all 24 hours of the 152 days.

### 2021-07-29 scripts/demo.R

@@ -10,8 +10,9 @@ print(aus_retail, n = 5)
 library(tidyverse)
 library(ggrepel)
 library(gghighlight)
+library(patchwork)
 
-aus_retail %>%
+p1 <- aus_retail %>%
   as_tibble() %>% # gghighlight issue for group_by() + filter()
   mutate(group = paste(State, ":", Industry)) %>%
   ggplot(aes(x = Month, y = Turnover)) +
@@ -22,7 +23,7 @@ aus_retail %>%
   ) +
   scale_x_yearmonth(breaks = yearmonth(c("1990 Jan", "2000 Jan", "2010 Jan")))
 
-aus_retail %>%
+p2 <- aus_retail %>%
   features(Turnover, feat_stl) %>%
   mutate(group = paste(State, ":", Industry)) %>%
   ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
@@ -34,3 +35,5 @@ aus_retail %>%
   ) +
   geom_label_repel(aes(label = group), vjust = 2, nudge_x = -0.1)
 
+p1 + p2 +
+  plot_annotation(tag_levels = "a")

### 2021-08-04 rjdistill_version/scripts/demo.R
reduced size of main interactive plot and fixed label locations

@@ -2,6 +2,7 @@
 library(feasts)
 library(tsibble)
 library(tsibbledata)
+library(patchwork)
 
 ## ---- print-retail
 print(aus_retail, n = 5)
@@ -11,18 +12,18 @@ library(tidyverse)
 library(ggrepel)
 library(gghighlight)
 
-aus_retail %>%
+a <- aus_retail %>%
   as_tibble() %>% # gghighlight issue for group_by() + filter()
   mutate(group = paste(State, ":", Industry)) %>%
   ggplot(aes(x = Month, y = Turnover)) +
   geom_line(aes(group = group)) +
   gghighlight(
     State == "Queensland", Industry == "Department stores",
-    label_params = list(vjust = -2)
+    label_params = list(vjust = 0, nudge_y = 1, label.size = 0.15)
   ) +
   scale_x_yearmonth(breaks = yearmonth(c("1990 Jan", "2000 Jan", "2010 Jan")))
 
-aus_retail %>%
+b <- aus_retail %>%
   features(Turnover, feat_stl) %>%
   mutate(group = paste(State, ":", Industry)) %>%
   ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
@@ -32,5 +33,7 @@ aus_retail %>%
     State == "Queensland", Industry == "Department stores",
     use_direct_label = FALSE
   ) +
-  geom_label_repel(aes(label = group), vjust = 2, nudge_x = -0.1)
+  geom_label_repel(aes(label = group), vjust = 0.4,
+                   nudge_x = -0.1, label.size = 0.15)
 
+a+b

### 2021-08-04 rjdistill_version/scripts/tourism.R

@@ -1,11 +1,15 @@
 ## ---- tourism-shared
 library(tsibble)
 library(tsibbletalk)
-tourism_shared <- tourism_monthly %>% 
+library(dplyr)
+tourism_shared <- tourism_monthly %>%
+  # Comment out the next line to run the full example
+  filter(State %in% c("Tasmania", "Western Australia")) %>%
+  mutate(Region = stringr::str_replace(Region, "Australia's ", "WA's ")) %>%
   as_shared_tsibble(spec = (State / Region) * Purpose)
 
 ## ---- plotly-key-tree
-p_l <- plotly_key_tree(tourism_shared, height = 1100, width = 800)
+p_l <- plotly_key_tree(tourism_shared, height = 800, width = 800)
 
 ## ---- tourism-series
 library(ggplot2)
@@ -27,8 +31,8 @@ p_br <- tourism_feat %>%
 library(plotly)
 subplot(p_l,
   subplot(
-    ggplotly(p_tr, tooltip = "Region", width = 1100),
-    ggplotly(p_br, tooltip = "Region", width = 1100),
+    ggplotly(p_tr, tooltip = "Region", width = 700),
+    ggplotly(p_br, tooltip = "Region", width = 700),
     nrows = 2),
   widths = c(.4, .6)) %>%
   highlight(dynamic = TRUE)

### 2021-08-09 rjdistill_version/tsibbletalk.Rmd
reduced size of main interactive plot and fixed label locations
updates to distill version

@@
 
-```{r tourism-multi, echo = TRUE, eval = knitr::is_html_output(), fig.cap = "Exploring an ensemble of linked plots of the Australian tourism data, built on a \\code{tourism\\_shared} object.
+```{r tourism-multi, layout="l-body-outset", echo = TRUE, eval = knitr::is_html_output(), fig.cap = "Exploring an ensemble of linked plots of the Australian tourism data, built on a \\code{tourism\\_shared} object.
 Click one of the nodes in the hierarchical tree to enable persistent linked brushing to compare two groups.
-Points and lines can also be selected in other plots."}
+Points and lines can also be selected in other plots.
+(Only Western Australia and Tasmania are included for the interactive plot in the html version, so size reasons.)"}
 ```

### 2021-08-09 rjdistill_version/tsibbletalk.R

@@ -19,29 +19,30 @@ read_chunk("scripts/pedestrian.R")
 library(feasts)
 library(tsibble)
 library(tsibbledata)
+library(patchwork)
 
 
 ## ---- print-retail------------------------------------------------------------
 print(aus_retail, n = 5)
 
 
-## ----highlight-retail, fig.height = 6, out.width = ".49\\linewidth", fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted. (a) An overlaid time series plot. (b) A scatter plot drawn from their time series features, where each dot represents a time series from (a).", fig.subcap = c("", ""), fig.ncol = 2----
+## ----highlight-retail, fig.height = 3.6, fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted. (a) An overlaid time series plot. (b) A scatter plot drawn from their time series features, where each dot represents a time series from (a)."----
 library(tidyverse)
 library(ggrepel)
 library(gghighlight)
 
-aus_retail %>%
+a <- aus_retail %>%
   as_tibble() %>% # gghighlight issue for group_by() + filter()
   mutate(group = paste(State, ":", Industry)) %>%
   ggplot(aes(x = Month, y = Turnover)) +
   geom_line(aes(group = group)) +
   gghighlight(
     State == "Queensland", Industry == "Department stores",
-    label_params = list(vjust = -2)
+    label_params = list(vjust = 0, nudge_y = 1, label.size = 0.15)
   ) +
   scale_x_yearmonth(breaks = yearmonth(c("1990 Jan", "2000 Jan", "2010 Jan")))
 
-aus_retail %>%
+b <- aus_retail %>%
   features(Turnover, feat_stl) %>%
   mutate(group = paste(State, ":", Industry)) %>%
   ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
@@ -51,13 +52,20 @@ aus_retail %>%
     State == "Queensland", Industry == "Department stores",
     use_direct_label = FALSE
   ) +
-  geom_label_repel(aes(label = group), vjust = 2, nudge_x = -0.1)
+  geom_label_repel(aes(label = group), vjust = 0.4,
+                   nudge_x = -0.1, label.size = 0.15)
+
+a+b
 
 
 ## ----tourism-shared, echo = TRUE----------------------------------------------
[…]
